### 2017-03-21 paper/frag_background.tex
Some background stuff

\section{Background}
% Potential additions:
% - netlink? generic netlink?

\subsection{Multipath TCP}
% Our discussion of Multipath TCP needs to be in enough depth that the reader
% understands how it works - essentially covering at least the first few pages
% of the RFC. In particular:
% - Subflows: what they are and how they work
% - How the initial subflow is created
% - How subsequent subflows are created and authenticated
% - Data sequencing and how data sequences are mapped onto subflow sequences
% - path management and data scheduling
% - ? how the linux kernel implementation is structured

In recent years, multi-homed devices have become increasingly common. The most
obvious example of these devices is the smartphone, which typically comes with
at least two network interfaces: one for cellular data, and one for Wi-Fi. While
these devices have become more common, most Internet protocols do not have ways
to take advantage of multiple interfaces.

Multipath TCP (MPTCP) is an extension to TCP which aims to mitigate this
problem. It is designed with several goals in mind:

\begin{itemize}
  \item To improve both the throughput and reliability of connections, relative
    to normal TCP.
  \item To remain compatible with applications that use TCP, so that they could
    use MPTCP without modification.
  \item To remain compatible with the Internet as a whole, especially with
    ``middleware'' such as firewalls, proxies, and NATs.
\end{itemize}
% https://tools.ietf.org/html/rfc6182#section-2

It achieves these goals by layering the protocol b

\subsection{Netfilter}
% Discuss what it is, and how it can be used to perform NAT

\subsection{OpenVPN}
% This will have to at least touch on TUN/TAP devices, and the variations of the
% OpenVPN protocol.

### 2017-03-22 paper/frag_implementation.tex
Start of implementation

\section{Implementation}

% smoother transition once we have everything

% Section intro:
% 1. Expectations (what hosts are there, and which ones run our code?)
% 2. Overview of the distinct pieces of software (client kernel, client daemon,
%    detour daemon)
% Subsection - Two Ways of Detours
% Subsection - Client kernel
% Subsection - Client daemon
% Subsection - Detour daemon

The implementation of this project is based on the following expectations.
First, the server must support Multipath TCP. However, beyond this, no other
modifications should be required of the server. Second, the client and detour
point are assumed to be ``active'' participants in the collective, and so they
may run whatever software is necessary to participate.

The client must run a modified kernel, which supports creating MPTCP subflows
across detours. In addition, the client must also run a daemon which discovers
and negotiates with detour points, finally reporting available ones to the
kernel. These concerns remain separate for several reasons. First, communication
with detour points requires communication over the Internet. While it is
possible to do user-level networking within the kernel, it is much simpler and
more future-proof to use the userspace socket API. Second, one type of detour
requires starting an OpenVPN process, an operation which should only be done
from userspace. Finally, userspace tools may be more easily configured and can
perform a more broad spectrum of computations, such as floating point math.

The detour need not support Multipath TCP. However, it must run one of two
possible daemons, depending on which type of detour it will be hosting.

\subsection{Detour Types}

In this project, we implement two different mechanisms for tunneling MPTCP
subflows across a detour.

\subsubsection{OpenVPN Tunneling}

The first mechanism involves the open-source tool OpenVPN. Detours run a
specially configured OpenVPN server, and clients connect. Unlike typical VPN
configurations, the client does not route all traffic through the VPN. Instead,
the client simply uses the VPN as a second network device. The client can then
use the VPN network device just as it would use any other physical device.

Traffic that goes through the VPN must undergo network address translation upon
reaching the detour point, since it functionally acts as a router.

\subsubsection{NAT Tunneling}

The second approach directly leverages NAT, without using OpenVPN for tunneling.
Since we only want to forward certain connections, we don't need the full power
of OpenVPN, and thus we can circumvent the overhead that comes with
encapsulation of datagrams.

The Linux kernel contains a robust packet filtering framework, known as
Netfilter. This framework includes utilities for performing network address
translation (NAT). This framework is typically what is used by consumer-grade
routers. Instead of encapsulation, the NAT tunneling approach directly addresses
the IP packets to the detour. The detour rewrites the source and destination
addresses using Netfilter for connection tracking.

### 2017-03-23 paper/frag_implementation.tex
Decent progress on detour stuff

@@
 \section{Implementation}
-
-% smoother transition once we have everything
-
 % Section intro:
@@
 %    detour daemon)
-% Subsection - Two Ways of Detours
+% Subsection - Detour daemon
 % Subsection - Client kernel
 % Subsection - Client daemon
-% Subsection - Detour daemon
+
+% smoother transition once we have everything
 
@@
 
-\subsection{Detour Types}
+\subsection{Detour Daemon}
 
 In this project, we implement two different mechanisms for tunneling MPTCP
-subflows across a detour.
+subflows across a detour. In both cases, the detour host must run a daemon that
+assists in tunneling traffic from the client to the server.
 
@@
 The first mechanism involves the open-source tool OpenVPN. Detours run a
-specially configured OpenVPN server, and clients connect. Unlike typical VPN
-configurations, the client does not route all traffic through the VPN. Instead,
-the client simply uses the VPN as a second network device. The client can then
-use the VPN network device just as it would use any other physical device.
+specially configured OpenVPN server. Clients connect to this VPN, creating a
+virtual network device. The client receives a routing rule from the detour
+advertising that it can reach all hosts on the Internet with a very high cost.
+As a result, the client operating system will prefer other routes to the VPN.
 
-Traffic that goes through the VPN must undergo network address translation upon
-reaching the detour point, since it functionally acts as a router.
+The VPN connection is configured as follows:
+
+\begin{itemize}
+\item The connection is over UDP, to avoid the so-called ``TCP Meltdown'' effect
+  caused by two congestion control algorithms interfering with each other.
+  % TODO citation
+\item The initial connection negotiation requires the client and server to
+  exchange certificates. OpenVPN allows the user to specify the PKI root. A
+  centralized detour management system could authorize clients and servers by
+  signing certificates, enabling tunneling for members and preventing it for
+  non-members.
+\item Subsequent messages are not protected by encryption or signatures, to
+  avoid the computational overhead.
+\end{itemize}
+
+One important aspect of VPN configuration is that the detour must create a
+private IP subnet. It will provide a DHCP service in order to assign IP
+addresses to clients as they connect to the VPN. Clients are expected to be able
+to connect to multiple detours simultaneously. In order to do this without
+conflict, each detour must use a distinct private IP subnet, so that there is no
+chance of the client addresses overlapping, and also so that the gateway address
+of the VPN is unique on the client.
+
+In a large-scale implementation of this technology, these subnet allocations
+could be handled by a centralized detour management server. The 10.0.0.0/8
+subnet is reserved for private networks, and if each detour were to establish
+its own /24 subnet, there would be capacity for 65,536 non-conflicting subnets.
+However, in our testing, subnets were manually allocated to avoid conflicts.
+
+Finally, in order to forward VPN traffic to the internet, the detour must be
+configured (via Netfilter) to perform network address translation (NAT) on
+outgoing packets. This process is typical for a home router and for VPNs which
+provide Internet access. Multipath TCP is designed with this behavior in mind,
+and so it is capable of functioning through NAT, so long as the middleboxes do
+not remove TCP options in transit.
+
+% TODO cite MPTCP design/architecture for middleboxes
+
+The OpenVPN implementation has several attractive qualities. First, it relies on
+a well-used protocol for forwarding traffic. Second, it need not be configured
+on a per-connection basis. Finally, it provides a built in mechanism for
+mutually authenticating the client and detour. However, there are some
+drawbacks. Since packets are tunneled, at least 28 bytes of overhead (IP and UDP
+headers) are required. To avoid fragmentation, the connection's maximum segment
+size (MSS) must be adjusted.
 
@@
 
-The second approach directly leverages NAT, without using OpenVPN for tunneling.
-Since we only want to forward certain connections, we don't need the full power
-of OpenVPN, and thus we can circumvent the overhead that comes with
-encapsulation of datagrams.
+The second approach directly modifies packets, without actually using a protocol
+(like OpenVPN) for tunneling. Since we only need to forward some connections to
+some destinations, the full power of OpenVPN is not necessary. Instead, packets
+are addressed directly to the detour. The detour can forward these to the
+final destination, and forward reply packets to the client. In order to do this,
+the detour must have advance knowledge of the final destination of the packet.
+So, clients use a simple signaling protocol to request tunnels via a detour.
 
-The Linux kernel contains a robust packet filtering framework, known as
-Netfilter. This framework includes utilities for performing network address
-translation (NAT). This framework is typically what is used by consumer-grade
-routers. Instead of encapsulation, the NAT tunneling approach directly addresses
-the IP packets to the detour. The detour rewrites the source and destination
-addresses using Netfilter for connection tracking.
+The Linux kernel's Netfilter framework allows for two types of NAT. The first
+kind is source NAT, or SNAT, rewrites the source address of the packet. This is
+the typical form of NAT used by home routers: the source address of an outgoing
+connection is rewritten to be the router's external IP address, but the
+destination is unchanged. The second kind of NAT supported by Netfilter is
+destination NAT, or DNAT. This rewrites the destination address of the packet.
+
+% TODO - netfilter docs citation
+
+The detour takes advantage of both forms of NAT offered by the kernel. First, it
+applies DNAT, setting the destination address to be the final server's IP
+address (arranged ahead of time). Then, it applies SNAT, setting the source
+address to be its own external IP. The Netfilter framework remembers these
+connections so that reply packets are properly forwarded back to the client.
+
+Thanks to Netfilter, this NAT mapping can be arranged with two \texttt{iptables}
+commands, with no custom kernel or user-space code required for packet
+forwarding. It is possible to implement this in userspace using raw sockets, but
+this approach has several drawbacks. It would have to be single-threaded, and it
+would require that packet data be copied from kernel space into user space and
+back. More critically, the connection tracking provided by the kernel is more
+robust than what would be implemented by user space.
+
+As mentioned earlier, the server address must be communicated to the detour
+ahead of time in order to use this scheme. To that end, we have designed a
+simple UDP-based for clients to ``request'' NAT mappings from a detour. The
+detour listens for requests on UDP port 45672. The request format is shown
+below.
+
+\begin{listing}
++-----------------------------------+
+| ver(1) | op(1)  | reserved (2)    |
++-----------------------------------+
+|           rip (4 bytes)           |
++-----------------------------------+
+| rpt (2 bytes)   | dpt (2 bytes)   |
++-----------------|-----------------+
+\end{listing}

### 2017-03-27 paper/frag_implementation.tex
More progress

@@
 
-\begin{listing}
+\begin{lstlisting}
 +-----------------------------------+
@@
 +-----------------|-----------------+
-\end{listing}
+\end{lstlisting}
+
+The \texttt{ver} field is a protocol version, currently set to 1. The
+\texttt{op} field contains the operation of this message: request (0) or
+response (1). The \texttt{reserved} field is unused, currently existing to pad
+the request. The field \texttt{rip} (short for ``remote IP'') is the address
+that the client wishes to create a detour to. \texttt{rpt} is the TCP port on
+the server to create a detour to. To understand the \texttt{dpt} field, it is
+useful to describe all of the addresses and ports in use during a tunneled
+connection:
+\begin{itemize}
+\item Client IP, port: the IP address of the client, and the client-side port,
+  which is usually ephemeral.
+\item Detour IP, port (client side): the IP and port that the client will
+  address its communication to.
+\item Detour IP, port (remote side): the IP and port which are used when the
+  detour opens its corresponding connection on the remote.
+\item Remote IP, port: the IP and port which the client is tunneling its
+  connection to.
+\end{itemize}
+
+The \texttt{dpt} field corresponds to the detour port, client side. In a
+request, the client proposes a port. In the response, the detour specifies the
+actual port which will be used. The server may wish to alter the \texttt{dpt}
+field for several reasons:
+\begin{itemize}
+\item A tunnel to the remote IP, port pair is already open on a different detour
+  port.
+\item The client already has a detour open (to a different remote) on the given
+  port.
+\item The detour may have a socket accepting connection on the proposed port,
+  and so it may wish to avoid creating a tunnel which would prevent the client
+  from accessing that service.
+\end{itemize}
+
+In order to create the tunnels, the detour runs two IPTables commands, which
+create NAT rules.

### 2017-03-30 paper/frag_implementation.tex
Some progress on paper

@@
 In order to create the tunnels, the detour runs two IPTables commands, which
-create NAT rules.
+create NAT rules. The first performs DNAT:
+
+\begin{lstlisting}
+iptables -t nat -A PREROUTING \
+         -s (*\emph{ClientAddress}*) \
+         -d (*\emph{DetourAddress}*) \
+         -p tcp --dport (*\emph{DetourPort}*) \
+         -j DNAT --to (*\emph{RemoteAddress}*):(*\emph{RemotePort}*)
+\end{lstlisting}
+
+In this command, we add to the \texttt{nat} table, on the \texttt{PREROUTING}
+chain. Incoming packets from the client, addressed to the agreed upon
+\textit{DetourPort} are sent to the \texttt{DNAT} chain, which rewrites the
+destination address to be the agreed upon remote address and port. This must
+occur before the kernel routes the packet. Once the kernel has routed the
+packet, the DNAT rule can be applied:
+
+\begin{lstlisting}
+iptables -t nat -A POSTROUTING \
+         -s (*\emph{ClientAddress}*) \
+         -d (*\emph{RemoteAddress}*) \
+         -p tcp --dport (*\emph{RemotePort}*) \
+         -j SNAT --to (*\emph{DetourAddress})
+\end{lstlisting}
+
+Since the destination address and port have already been modified, this rule
+looks for packets addressed to the remote address and port, but still using the
+client address as the source. It jumps these packets to the DNAT chain, where
+their source address will be rewritten as the detour's address.
+
+It is important to understand that these rules are generally only applied to the
+initial SYN segment of the TCP flows. Once these rules are applied, the
+connection tracking mechanism of the kernel takes over, ensuring that these
+rules are applied in reverse to response packets.
+
+The detour daemon is implemented as a Python script. It listens for tunnel
+requests, executes the correct commands to create them, and then sends
+responses. It ensures that detour ports do not collide with each other, or with
+ports that are open on the daemon machine. It also ensures that on termination,
+all tunnels are torn down.
+
+This mechanism has the advantage that it requires zero overhead. Since no
+headers are added, the full amount of data may be transmitted on these segments.
+However, its disadvantage is that tunnels must be configured each time a new
+connection is created on the client. While the client may simply send packets to
+a new remote through an OpenVPN tunnel, it must incur a RTT delay requesting a
+new tunnel from the NAT detour before it can begin establishing a subflow.
+
+\subsection{Client Daemon}
+
+In order for the client to use detours, it must either initiate OpenVPN
+connections with them, or dynamically request tunnels via the UDP protocol. In
+either case, a client program must manage these activities. In particular, a
+program must be able to notify the kernel when new detours are available for
+immediate use, and it must be able to respond to requests from the kernel for
+new detours. The mechanism for communication with the kernel will be discussed
+in the following section.
+
+\subsubsection{OpenVPN Client Management}
+
+The client daemon reads a configuration file which informs it of detours which
+use OpenVPN. On startup, it initiates an OpenVPN connection with each detour.
+Once the connections are established, it sends a message to the kernel with the
+virtual network device's name.
+
+\subsubsection{NAT Client Management}
+
+In order to use NAT tunnels, the client daemon must be much more complex. The
+client waits for kernel tunnel requests. Upon receiving these requests, it sends
+a UDP request to each NAT detour server specified in its configuration file. It
+then waits for responses from the detour servers. Upon receiving each response,
+it sends a reply back to the kernel notifying it of the new NAT tunnel.
+
+\subsection{Client Communication}
+
+In order to communicate between kernel and userspace, we use Generic Netlink
+sockets. Netlink is an address family for the socket interface
+(\texttt{AF\_NETLINK}, similar to \texttt{AF\_INET}), which allows just this
+sort of communication. There are several types of Netlink socket, but the
+Generic Netlink type allows any kernel component to define its own protocol,
+while still managing marshalling data.
+
+Our protocol is similar to the UDP protocol for communication between the client
+and detour. The kernel sends a message of type \texttt{DETOUR_C_REQ} to a
+``multicast group'' which the client daemon subscribes to. This message contains
+the necessary parameters (remote address, remote port). The client daemon may
+reply to the kernel with \texttt{DETOUR_C_ADD} commands, which add detour
+entries, and \texttt{DETOUR_C_DEL} commands, which remove them.
+
+The protocol allows the client to register both OpenVPN and NAT based detours,
+both of which are supported in the kernel implementation.
+
+\subsection{Client Kernel}
+
+Finally, the client kernel must be capable of requesting detours, receiving
+information from userspace, and using that information to establish subflows.
+Support for this functionality is implemented as a Linux kernel module. In
+particular, it is implemented as a MPTCP ``path manager.'' A path manager is a
+kernel module which makes decisions regarding when to advertise additional
+addresses and when to create new subflows. An instance of the path manager is
+created for each MPTCP connection, when the connection becomes fully established
+(that is, after the third packet in the handshake).
+
+Upon creation, our path manager enqueues a task to be executed in a long-running
+kernel thread. This task first sends a message to userspace on the Generic
+Netlink socket, requesting a detour. Then, the task looks through an internal
+list of detours. The kernel maintains two lists of detours. The first is a list
+of OpenVPN detours. These detours are not specific to the remote port and host.
+The second list contains NAT tunnels which were created by the client daemon and
+sent to the kernel via the \texttt{DETOUR_C_ADD} command.
+
+The task attempts to add up to $N$ subflows, where $N$ is a configurable limit.
+If it cannot (due to lack of subflows), it terminates. If OpenVPN detours are
+available, then they take precedence over the NAT detours, because they are
+available immediately.
+
+When the client receives a \texttt{DETOUR_C_ADD} command, it first adds the
+details of the tunnel to its internal list. Then, if the tunnel was a NAT
+tunnel, it wakes all the path managers so that they can attempt to create a
+subflow through that tunnel.

### 2017-04-06 reports/experiment.tex
Add experiment report.

\documentclass{article}

\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{color}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\lstset{%
  backgroundcolor=\color{white},     % background color
  basicstyle=\footnotesize\ttfamily, % monospace, small font
  breaklines=true,                   % nice line breaks
  captionpos=b,                      % put captions at bottom
  commentstyle=\color{mygreen},      % comments
  escapeinside={(*}{*)},             % if you want to add LaTeX in code
  keywordstyle=\color{blue},         % keyword
  stringstyle=\color{mymauve},       % string
}

\title{Proposed Experimentation}
\author{Stephen Brennan}

\begin{document}

\maketitle

\section{Background}

There are several options for deploying networking experiments. In this section,
I will describe several different options, the services they provide, and their
relationships. In particular, I will discuss PlanetLab, GENI, EmuLab, and M-Lab.

\subsection{M-Lab}

\href{https://measurementlab.net}{M-Lab} (more formally Measurement Lab) is a
large-scale, open source measurement partnership. Partners include Google Open
Source Research and PlanetLab itself. M-Lab allows researchers to create
Internet measurement tests that will run on their infrastructure. However, the
requirements for these tests are very restrictive. In particular:

\begin{itemize}
\item Tests must be actively initiated by the client, by visiting the website.
\item Code must be open source.
\item Data must be released into the public domain.
\end{itemize}

Users visit the website and manually run tests (such as Internet speed tests).
As a result, this framework is not likely to be useful for experimentation on
MPTCP, since most users do not run MPTCP capable machines, let alone those with
my customized kernel.

\subsection{PlanetLab}

\href{https://www.planet-lab.org/}{PlanetLab} appears to be the one of the
oldest research frameworks for Internet research. It allows researchers to
create ``slices'' which are containers for resources. Researchers may assign
``nodes'' to their slices, which will give them a virtual server on the slice.
This virtual server is not a true virtual machine - rather it is a lighter
weight containment solution. Although the User Guide suggests that this
mechanism is ``vserver'', more recent announcements suggest that the current
form of virtualization is ``LXC''. In either case, this means that users may not
directly modify the running kernel of their slice. It may be possible to create
a virtual machine on these virtual servers, but the performance impact may
suffer.

Much of the information on PlanetLab's website appears to be out of date or
unmaintained. It is difficult to be certain about what is and is not possible on
the service, when the User Guide itself was last updated in 2006.

\subsection{GENI}

GENI is a newer suite. It has a similar architecture to PlanetLab, organizing
research into slices, nodes, slivers, etc. However, it differs from PlanetLab in
a few key areas. First, its documentation and website are much newer and give a
good deal more information about the services available. Second, it describes
itself as a ``federation'' of computing resource ``aggregates'', including
PlanetLab. In other words, using GENI tools and APIs, researchers may request
and use PlanetLab's resources, along with those of EmuLab and others.

GENI's resources are therefore much more diverse than PlanetLab's. Of course,
PlanetLab slivers would have the same capabilities, regardless of whether you
accessed them through GENI or through PlanetLab directly. However, GENI also
allows access to aggregates which can run custom operating system images. This
would be much more suitable to an experiment which requires a modified kernel.

A downside of GENI is that, although it allows probably the broadest reach in
terms of resource diversity, it may not allow as much control over these
resources as the aggregates themselves might allow. An example is EmuLab, which
will be discussed next. EmuLab resources are likely to have the most flexibility
when requested through EmuLab itself, rather than through GENI. However, I
cannot really be certain about these things until I try them out.

\subsection{EmuLab}

EmuLab describes itself as a ``total network testbed'', giving researchers the
flexibility to emulate arbitrary network topologies. It aims to allow
controllable and repeatable experiments. One of the homepage's main claims is
that it allows you full root access and you may also run custom operating
systems. By all accounts, it seems like the best way to create reliable,
repeatable experiments for this project's use case.

\section{Experiment Description}

The experiments must be designed to answer the important questions about the
detour routing schemes. In particular:

\begin{itemize}
\item Does detour routing improve connection performance when the main route is
  unreliable? How much does it improve performance, and by what performance
  metrics?
\item Is there a measurable performance difference between the OpenVPN and NAT
  approaches to detour routing?
\item Is there a measurable performance difference between scheduling modules in
  this scenario?
\end{itemize}

To quantify performance, the following metrics will be useful:
\begin{itemize}
\item Throughput - total bytes transferred through a connection per unit time.
\end{itemize}

\subsection{Small Scale}

First, we should validate that the framework performs as expected on a
controlled network topology. A reasonable test framework would then be to set up
a network topology with multiple routes between a client and server, vary these
parameters, and compare the performance metrics to those of several controls:

\begin{itemize}
\item Vanilla TCP, using the route which has the ``lossy'' link.
\item Vanilla TCP, using the alternative route through the detour.
\item Multipath TCP, unmodified. In this scenario, the network topology would be
  altered such that the client has two interfaces and each has a different route
  to the server.
\end{itemize}

I would hypothesize that the detour routing solution will outperform the first
option, and it would be comparable to the rest.

% TODO: diagram of proposed topology

\subsection{Large Scale}

My ideas of what would make a suitable larger scale test are not as well-formed.
I believe we need a larger scale test to give some indication that this approach
could work ``in the wild''. For this sort of test, rather than setting up a
controlled, repeatable network topology, we would simply use nodes in separate
sites and perform a similar data transfer. The transfer would have to be
repeated (control and MPTCP) several times in a row to give real statistical
backbone to the experiment.

\end{document}

### 2017-04-06 reports/experiment.tex
Update experiment plan

@@
 
+\subsection{Mininet}
+
+Mininet is a single-computer network emulation platform. It allows users to
+define arbitrary network topologies, including hosts, links, and switches. The
+hosts are simply Linux processes (including potentially child processes) running
+in a separate kernel network namespace. All networking occurs on the same
+kernel, meaning that the machine running the experiment must have a custom
+kernel (rather than simply running a virtual machine with the custom kernel).
+
+It should be possible to run experiments on Mininet - the biggest concern would
+be ensuring that the required daemons and kernel modifications don't conflict
+with each other, since they would be running on the same kernel with only
+network namespaces separating them.
+
+As far as performance, making a Mininet experiment ``network-limited'' requires
+that you assign bandwidth to each link that is less than the total your CPU
+could handle. On my laptop (within a virtual machine), this maximum is around 24
+Gbps. A reasonable setup could assign each link to be 100Mbps, have several
+links and routers, and not stress the CPU for processing this.
+
 \section{Experiment Description}
 
-The experiments must be designed to answer the important questions about the
-detour routing schemes. In particular:
+The point of this thesis is to propose a mechanism for applying detour routing
+to any TCP connection. While it is out of scope to attempt to determine an
+optimal detour routing scheme, some experimentation is necessary, if for no
+other reason than to validate that this mechanism can provide the benefits of
+Multipath TCP to alternative ``detour'' routes. The main benefit of MPTCP is
+that it provides link aggregation, or in this case, path aggregation. That is,
+we are able to use the capacity of more than one path in our connection.
+
+Rather than design experiments that attempt to determine the best strategy for
+picking a detour, or the best strategy for scheduling data when you have
+multiple connections, we should simply show that this implementation provides
+these aggregation benefits. There are a couple types of benefits that we should
+attempt to quantify:
 
 \begin{itemize}
-\item Does detour routing improve connection performance when the main route is
-  unreliable? How much does it improve performance, and by what performance
-  metrics?
-\item Is there a measurable performance difference between the OpenVPN and NAT
-  approaches to detour routing?
-\item Is there a measurable performance difference between scheduling modules in
-  this scenario?
+\item Better throughput. In the case that our connection speed is bottlenecked
+  by a link that we can circumvent, we should be able to aggregate bandwidth so
+  that we improve throughput.
+\item Better response time. In the case that one path has high delay, we should
+  be able to improve response time by adding another path.
+\item Better fault tolerance. If a path has a high loss rate, we should be able
+  to improve our throughput by circumventing that lossy link.
 \end{itemize}
 
-To quantify performance, the following metrics will be useful:
+So, the experiments are designed around these three scenarios. We use the same
+network topologies for each scenario, but we alter the connection properties of
+a link along the ``Internet routed path'' (but not the detour routed path).
+Then, we compare the desired property (throughput or response time) to the
+following controls:
+
 \begin{itemize}
-\item Throughput - total bytes transferred through a connection per unit time.
+\item Vanilla TCP. No multipath TCP, just a single TCP flow from client to
+  server.
+\item Vanilla TCP, on the detour route.
 \end{itemize}
 
-\subsection{Small Scale}
+This way, we can determine whether the MPTCP detour mechanism is able to (at
+least) leverage the performance of the better path, if not aggregate their
+performance.
 
-First, we should validate that the framework performs as expected on a
-controlled network topology. A reasonable test framework would then be to set up
-a network topology with multiple routes between a client and server, vary these
-parameters, and compare the performance metrics to those of several controls:
-
-\begin{itemize}
-\item Vanilla TCP, using the route which has the ``lossy'' link.
-\item Vanilla TCP, using the alternative route through the detour.
-\item Multipath TCP, unmodified. In this scenario, the network topology would be
-  altered such that the client has two interfaces and each has a different route
-  to the server.
-\end{itemize}
-
-I would hypothesize that the detour routing solution will outperform the first
-option, and it would be comparable to the rest.
-
-% TODO: diagram of proposed topology
-
-\subsection{Large Scale}
-
-My ideas of what would make a suitable larger scale test are not as well-formed.
-I believe we need a larger scale test to give some indication that this approach
-could work ``in the wild''. For this sort of test, rather than setting up a
-controlled, repeatable network topology, we would simply use nodes in separate
-sites and perform a similar data transfer. The transfer would have to be
-repeated (control and MPTCP) several times in a row to give real statistical
-backbone to the experiment.
+These experiments could be run on Mininet or Emulab. In both cases, the results
+would be highly reproducible. If Mininet were used, we could run very many
+trials very quickly in order to get high statistical power. In addition, for
+publication we could provide a complete virtual machine image which could
+reproduce all experiments. Finally, if changes were required to the detour
+framework (as they will probably be), Mininet will allow for the quickest
+response time, so that we can run and re-run our tests quickly.
 

### 2017-04-26 paper/frag_implementation.tex

@@
 Our protocol is similar to the UDP protocol for communication between the client
-and detour. The kernel sends a message of type \texttt{DETOUR_C_REQ} to a
+and detour. The kernel sends a message of type \texttt{DETOUR\_C\_REQ} to a
 ``multicast group'' which the client daemon subscribes to. This message contains
 the necessary parameters (remote address, remote port). The client daemon may
-reply to the kernel with \texttt{DETOUR_C_ADD} commands, which add detour
-entries, and \texttt{DETOUR_C_DEL} commands, which remove them.
+reply to the kernel with \texttt{DETOUR\_C\_ADD} commands, which add detour
+entries, and \texttt{DETOUR\_C\_DEL} commands, which remove them.
 
@@
 The second list contains NAT tunnels which were created by the client daemon and
-sent to the kernel via the \texttt{DETOUR_C_ADD} command.
+sent to the kernel via the \texttt{DETOUR\_C\_ADD} command.
 
@@
 
-When the client receives a \texttt{DETOUR_C_ADD} command, it first adds the
+When the client receives a \texttt{DETOUR\_C\_ADD} command, it first adds the
 details of the tunnel to its internal list. Then, if the tunnel was a NAT

### 2017-04-26 paper/frag_implementation.tex
Additional implementation discussion

--- 2017-03-27 paper/frag_implementation.tex
@@
 
-The implementation of this project is based on the following expectations.
-First, the server must support Multipath TCP. However, beyond this, no other
-modifications should be required of the server. Second, the client and detour
-point are assumed to be ``active'' participants in the collective, and so they
-may run whatever software is necessary to participate.
+In this project, we implement a system which allows one host to dynamically add
+new overlay routes to a Multipath TCP connection. Each overlay route is a new
+subflow, which is tunneled across a third-party host. The system involves at
+least three hosts:
+\begin{itemize}
+\item The \emph{client} is the host which adds detour routes, attempting to
+  improve its connection quality. The client must be the host which initiates
+  the MPTCP connection.
+\item The \emph{server} is the host with which the client wishes to communicate.
+\item The \emph{detour} is a host which acts as a waypoint along the overlay
+  route from the client to the server. It tunnels the MPTCP between them. There
+  may be more than one detour at a time.
+\end{itemize}
 
-The client must run a modified kernel, which supports creating MPTCP subflows
-across detours. In addition, the client must also run a daemon which discovers
-and negotiates with detour points, finally reporting available ones to the
-kernel. These concerns remain separate for several reasons. First, communication
-with detour points requires communication over the Internet. While it is
-possible to do user-level networking within the kernel, it is much simpler and
-more future-proof to use the userspace socket API. Second, one type of detour
-requires starting an OpenVPN process, an operation which should only be done
-from userspace. Finally, userspace tools may be more easily configured and can
-perform a more broad spectrum of computations, such as floating point math.
+We assume only that the server supports Multipath TCP, and require no additional
+customization for our system to work. In our current implementation, the detour
+host must run some version of Linux, and the client must run a custom version of
+Linux.
 
-The detour need not support Multipath TCP. However, it must run one of two
-possible daemons, depending on which type of detour it will be hosting.
+The system is implemented via a number of interacting components. First, a path
+manager extension for the Linux MPTCP implementation, which allows the client's
+kernel to request new detour paths, receive them, and create subflows across
+them. Second, a user-space utility which receives messages from the kernel and
+creates the necessary detour paths. Finally, we implement two different types of
+detour daemons, which run on the detour host. These components are illustrated
+in Figure~\ref{f:MovingParts}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=\textwidth]{figures/MovingParts.pdf}
+  \caption{
+    The components and their communications. In \textbf{A}, the path manager
+    requests an overlay route. In \textbf{B}, the client daemon reports overlay
+    routes back to the path managers. In \textbf{C} and \textbf{D}, the client
+    and detour daemons negotiate an overlay route. In \textbf{E}, a subflow is
+    created through the detour.
+  }
+  \label{f:MovingParts}
+\end{figure}
+
+The typical interaction of these components is as follows. First, the client
+creates a MPTCP connection to the server. The first subflow uses the Internet
+routed path. Once this connection is fully established, the path manager
+requests overlay routes on which it may create subflows. The client and detour
+daemons negotiate a tunnel, which may be implemented in two different ways. The
+client reports this to the path-managers, which creates subflows for the MPTCP
+connection across every available detour, until a predefined limit is reached.
+
+In the remainder of this section, we will discuss each of these components at
+length.
 
@@
 
-In this project, we implement two different mechanisms for tunneling MPTCP
-subflows across a detour. In both cases, the detour host must run a daemon that
-assists in tunneling traffic from the client to the server.
+We have implemented two different mechanisms for tunneling MPTCP subflows across
+a detour. In both cases, the detour host must run a daemon that assists in
+tunneling traffic from the client to the server.
 
@@
 \item The initial connection negotiation requires the client and server to
-  exchange certificates. OpenVPN allows the user to specify the PKI root. A
-  centralized detour management system could authorize clients and servers by
-  signing certificates, enabling tunneling for members and preventing it for
-  non-members.
+  exchange certificates.
 \item Subsequent messages are not protected by encryption or signatures, to
-  avoid the computational overhead.
+  avoid overhead.
 \end{itemize}
@@
 headers) are required. To avoid fragmentation, the connection's maximum segment
-size (MSS) must be adjusted.
+size (MSS) may need to be adjusted, although in our testing, this was not
+required.
 
@@
 The Linux kernel's Netfilter framework allows for two types of NAT. The first
-kind is source NAT, or SNAT, rewrites the source address of the packet. This is
+kind, source NAT (SNAT), rewrites the source address of the packet. This is
 the typical form of NAT used by home routers: the source address of an outgoing
@@
 In order to create the tunnels, the detour runs two IPTables commands, which
-create NAT rules.
+create NAT rules. The first performs DNAT:
+
+\begin{lstlisting}
+iptables -t nat -A PREROUTING \
+         -s (*\emph{ClientAddress}*) \
+         -d (*\emph{DetourAddress}*) \
+         -p tcp --dport (*\emph{DetourPort}*) \
+         -j DNAT --to (*\emph{RemoteAddress}*):(*\emph{RemotePort}*)
+\end{lstlisting}
+
+In this command, we add to the \texttt{nat} table, on the \texttt{PREROUTING}
+chain. Incoming packets from the client, addressed to the agreed upon
+\textit{DetourPort} are sent to the \texttt{DNAT} chain, which rewrites the
+destination address to be the agreed upon remote address and port. This must
+occur before the kernel routes the packet. Once the kernel has routed the
+packet, the DNAT rule can be applied:
+
+\begin{lstlisting}
+iptables -t nat -A POSTROUTING \
+         -s (*\emph{ClientAddress}*) \
+         -d (*\emph{RemoteAddress}*) \
+         -p tcp --dport (*\emph{RemotePort}*) \
+         -j SNAT --to (*\emph{DetourAddress})
+\end{lstlisting}
+
+Since the destination address and port have already been modified, this rule
+looks for packets addressed to the remote address and port, but still using the
+client address as the source. It jumps these packets to the DNAT chain, where
+their source address will be rewritten as the detour's address.
+
+It is important to understand that these rules are generally only applied to the
+initial SYN segment of the TCP flows. Once these rules are applied, the
+connection tracking mechanism of the kernel takes over, ensuring that these
+rules are applied in reverse to response packets.
+
+The detour daemon is implemented as a Python script. It listens for tunnel
+requests, executes the correct commands to create them, and then sends
+responses. It ensures that detour ports do not collide with each other, or with
+ports that are open on the daemon machine. It also ensures that on termination,
+all tunnels are torn down.
+
+This mechanism has the advantage that it requires zero overhead. Since no
+headers are added, the full amount of data may be transmitted on these segments.
+However, its disadvantage is that tunnels must be configured each time a new
+connection is created on the client. While the client may simply send packets to
+a new remote through an OpenVPN tunnel, it must incur a RTT delay requesting a
+new tunnel from the NAT detour before it can begin establishing a subflow.
+
+\subsection{Path Manager}
+
+To enable subflows to use the detours created above, we implement a custom MPTCP
+path manager within the Linux kernel. The path manager is a module which
+dictates local policy for creating new subflows and announcing local addresses.
+For example, the \texttt{fullmesh} path manager creates a subflow between each
+address on the client and server.
+
+The path manager receives space within the MPTCP control buffer in order to
+track state, and it registers a set of callbacks with the protocol, so that it
+can be notified of events as they occur. Our path manager (named
+\texttt{detour}) runs a worker when a client-side MPTCP connection becomes fully
+established (i.e. after the three-way handshake completes). The first thing this
+is send a message to userspace, requesting a detour. Then, it looks through an
+internal list of detours. The kernel maintains two lists of detours (per network
+namespace). The first is a list of OpenVPN detours. These detour entries are not
+specific to a server, and they simply indicate the name of a network interface
+to use. The second list contains NAT detours. These consist of a detour address
+and port, and a server address and port. The worker may only select NAT entries
+which match the server address and port of the connection it is working on.
+
+The task attempts to add up to $N$ subflows, where $N$ is a configurable limit.
+It terminates in one of two ways. Either it has exhaused all of the available
+detours, but not yet created all $N$ subflows. Otherwise, if it creates $N$
+subflows successfully, it stops looking through the list, even if more detours
+are available. If OpenVPN detours are available, then they take precedence over
+the NAT detours, because they are available immediately. NAT detours created by
+the client daemon become available only once the daemon has processed the
+request from the kernel.
+
+The kernel listens for new detour routes. When it receives them, it adds them to
+the appropriate internal list. Then, it iterates over every MPTCP connection
+within the network namespace of the message, waking their workers to give them
+the opportunity to create a new subflow along this detour. If the detour was
+OpenVPN type, every worker notified. If the detour was NAT type, then only
+connections which match the server address and port are notified.
+
+\subsection{Client Daemon}
+
+The client daemon has several responsibilities. First, it must create OpenVPN
+connections report them to the kernel. Second, it must respond to detour
+requests from the kernel by creating NAT detours and reporting them back.
+
+The daemon has a configuration file which allows the user to specify the IP
+addresses of hosts to use for both types of detours. On startup, the daemon
+connects to each OpenVPN detour and reports the resulting network interfaces to
+the kernel. Then, it waits listening for detour requests from the kernel. For
+each request, it requests a detour from every NAT detour address. Upon receiving
+responses, it sends them the the kernel as well.
+
+\subsubsection{Client Communication}
+
+In order to communicate between kernel and userspace, we use Generic Netlink
+sockets. Netlink is an address family for the socket interface
+(\texttt{AF\_NETLINK}, similar to \texttt{AF\_INET}), which allows communication
+between kernel and userspace. Many existing Linux network configuration
+utilities communicate with the kernel over the Netlink address family, such as
+route and firewall configuration. Each of the subsystems being configured has a
+corresponding protocol family, such as \texttt{NETLINK\_NETFILTER} for Netfilter
+firewall configuration. These protocol families take the same place in the
+\texttt{socket()} system call as protocol families such as \texttt{IPPROTO\_TCP}
+in the \texttt{AF\_INET} and \texttt{AF\_INET6} address families.
+
+Creating a new Netlink protocol would involve allocating a new Netlink protocol
+family. To make this process easier, the Generic Netlink protocol family was
+created. It allows custom protocols (``Generic Netlink families'') to be
+created, dynamically allocated, and discovered. It also provides a common format
+for sending several data types, and creates a framework for defining message
+types. Further, Generic Netlink provides a multicast capability, allowing kernel
+and user sockets to broadcast messages to any socket which has joined a
+particular multicast group. Finally, Generic Netlink comes with a kernel and
+userspace library which allows for a consistent API and a level of abstraction
+when defining Generic Netlink families.
+
+To allow the kernel to request detours, and userspace to communicate the
+existence of these detours, we define a Generic Netlink family. This protocol is
+similar to the UDP protocol which the client daemon uses to request NAT tunnels.
+The kernel sends a message of type \texttt{DETOUR\_C\_REQ} to a ``multicast
+group'' which the client daemon subscribes to. This message contains the
+parameters necessary for the daemon to create a tunnel: the server address and
+port. The client daemon receives this message and creates tunnels, based on a
+configuration file. It then reports these tunnels to the kernel, using a
+\texttt{DETOUR\_C\_ADD} message type. The client daemon may also use the
+\texttt{DETOUR\_C\_DEL} message type to delete a tunnel entry it has previously
+added. Finally, the \texttt{DETOUR\_C\_ECHO} message type requests that the
+kernel log all tunnels for diagnostics.

### 2017-04-26 paper/paper.tex
Under-engineer the paper.

--- 2017-04-26 paper/frag_implementation.tex
@@
+\documentclass{article}
+
+\usepackage[margin=1.5in]{geometry}
+\usepackage{setspace}
+\usepackage{listings}
+\usepackage{graphicx}
+\usepackage[font={small,it}]{caption}
+\usepackage{color}
+\definecolor{mygreen}{rgb}{0,0.6,0}
+\definecolor{mygray}{rgb}{0.5,0.5,0.5}
+\definecolor{mymauve}{rgb}{0.58,0,0.82}
+\lstset{%
+  backgroundcolor=\color{white},     % background color
+  basicstyle=\footnotesize\ttfamily, % monospace, small font
+  breaklines=true,                   % nice line breaks
+  captionpos=b,                      % put captions at bottom
+  commentstyle=\color{mygreen},      % comments
+  escapeinside={(*}{*)},             % if you want to add LaTeX in code
+  keywordstyle=\color{blue},         % keyword
+  stringstyle=\color{mymauve},       % string
+}
+
+\doublespacing
+
+\title{Improving TCP connectivity via Multipath Overlays}
+\author{Stephen Brennan}
+\begin{document}
+\maketitle
+
+\section{Background}
+% Potential additions:
+% - netlink? generic netlink?
+
+\subsection{Multipath TCP}
+% Our discussion of Multipath TCP needs to be in enough depth that the reader
+% understands how it works - essentially covering at least the first few pages
+% of the RFC. In particular:
+% - Subflows: what they are and how they work
+% - How the initial subflow is created
+% - How subsequent subflows are created and authenticated
+% - Data sequencing and how data sequences are mapped onto subflow sequences
+% - path management and data scheduling
+% - ? how the linux kernel implementation is structured
+
+In recent years, multi-homed devices have become increasingly common. The most
+obvious example of these devices is the smartphone, which typically comes with
+at least two network interfaces: one for cellular data, and one for Wi-Fi. While
+these devices have become more common, most Internet protocols do not have ways
+to take advantage of multiple interfaces.
+
+Multipath TCP (MPTCP) is an extension to TCP which aims to mitigate this
+problem. It is designed with several goals in mind:
+
+\begin{itemize}
+  \item To improve both the throughput and reliability of connections, relative
+    to normal TCP.
+  \item To remain compatible with applications that use TCP, so that they could
+    use MPTCP without modification.
+  \item To remain compatible with the Internet as a whole, especially with
+    ``middleware'' such as firewalls, proxies, and NATs.
+\end{itemize}
+% https://tools.ietf.org/html/rfc6182#section-2
+
+It achieves these goals by layering the protocol b
+
+\subsection{Netfilter}
+% Discuss what it is, and how it can be used to perform NAT
+
+\subsection{OpenVPN}
+% This will have to at least touch on TUN/TAP devices, and the variations of the
+% OpenVPN protocol.
+
+\section{Related Work}
+% Binder
+% SCTP
+\subsection{Detour Routing}
+% this feels more like background than related work... this motivates the whole
+% project really.
+
 \section{Implementation}
@@
 kernel log all tunnels for diagnostics.
+
+\end{document}

### 2017-04-26 src/nat_detour.py
Correct hashbang and chmod u+x nat_detour.py

@@ -1,4 +1,4 @@
-#!/usr/bin/env python
+#!/usr/bin/env python3
 """
 Proxy daemon which configures NAT mappings within the kernel.
 

### 2017-04-26 setup.sh
Turns out you need openvpn to test openvpn.

@@ -3,7 +3,7 @@
 
 sudo dpkg -i tmp/*.deb
 
-sudo apt install -y python3-pip libnl-genl-3-dev
+sudo apt install -y python3-pip libnl-genl-3-dev openvpn
 sudo pip3 install psutil
 
 # Compile everything freshly on VM.

### 2017-04-26 src/experiment.py
Let's also let nat_detour.py specify its own python

@@ -147,7 +147,7 @@ def DetourNet(params):
 
 def setup_nat(client, detour):
     client.cmd('./client daemon ../etc/daemon-nat-vm.conf')
-    detour.cmd('python nat_detour.py &')
+    detour.cmd('./nat_detour.py &')
 
 
 def setup_vpn(client, detour):

### 2017-04-27 setup.sh
Use iperf3 -> causes some issues with openvpn, but we'll deal

@@ -3,7 +3,7 @@
 
 sudo dpkg -i tmp/*.deb
 
-sudo apt install -y python3-pip libnl-genl-3-dev openvpn
+sudo apt install -y python3-pip libnl-genl-3-dev openvpn iperf3
 sudo pip3 install psutil
 
 # Compile everything freshly on VM.

### 2017-04-27 src/experiment.py

@@ -171,8 +171,8 @@ def scenario(name, params, trials=30):
     client, detour, server = mn.get('client', 'detour', 'server')
     SETUP[params['scenario']](client, detour)
 
-    filename = '%s.%s.csv' % (name, params['scenario'])
-    server.sendCmd('iperf -s -y c | tee %s' % filename)
+    filename = '%s.%s.json' % (name, params['scenario'])
+    server.sendCmd('iperf3 -s -J | tee %s' % filename)
 
     # sleep synchronization is the worst, except for iperf
     import time; time.sleep(0.5)
@@ -180,8 +180,7 @@ def scenario(name, params, trials=30):
     for _ in range(trials):
         print('.', end='')
         sys.stdout.flush()
-        client.cmd('iperf -c ' + server.IP() + ' -y c')
-        server.monitor()     # ensure a line of server output was created
+        client.cmd('iperf3 -c ' + server.IP())
     print()
 
     mn.stop()

### 2017-04-27 src/experiment.py
I now hate martians, but this makes OpenVPN work

@@ -77,6 +77,17 @@ def clear_routes(*nodes):
         node.cmd('ip route flush table main')
 
 
+def disable_rp_filter(node):
+    # all -> used in computing the iface value. Be sure to set to 0
+    node.cmd('sysctl -w net.ipv4.conf.all.rp_filter=0')
+    # default -> used as value for new ifaces. Be sure to set to 0
+    node.cmd('sysctl -w net.ipv4.conf.default.rp_filter=0')
+    # lo -> jsut to be sure
+    node.cmd('sysctl -w net.ipv4.lo.default.rp_filter=0')
+    for name in node.nameToIntf:
+        node.cmd('sysctl -w net.ipv4.conf.%s.rp_filter=0' % name)
+
+
 def DetourNet(params):
     """A function that looks like a class... ಠ_ಠ.
 
@@ -141,6 +152,14 @@ def DetourNet(params):
     r1.cmd('sysctl -w net.ipv4.ip_forward=1')
     r2.cmd('sysctl -w net.ipv4.ip_forward=1')
     r3.cmd('sysctl -w net.ipv4.ip_forward=1')
+
+    ### CUT ME SOME ROUTING SLACK PLEASE
+    disable_rp_filter(r1)
+    disable_rp_filter(r2)
+    disable_rp_filter(r3)
+    disable_rp_filter(client)
+    disable_rp_filter(detour)
+    disable_rp_filter(server)
 
     return mn
 

### 2017-04-27 src/client.c
Create a thread for logging OpenVPN output and waiting

@@ -21,6 +21,7 @@
 #include <netinet/in.h>
 #include <arpa/inet.h>
 #include <regex.h>
+#include <pthread.h>
 
 #include <libconfig.h>
 #include <netlink/netlink.h>
@@ -36,6 +37,18 @@ void abort_perror(char *msg)
 {
 	perror(msg);
 	abort();
+}
+
+void pthread_mutex_lock_s(pthread_mutex_t *mutex)
+{
+	if (pthread_mutex_lock(mutex) != 0)
+		abort_perror("pthread_mutex_lock");
+}
+
+void pthread_mutex_unlock_s(pthread_mutex_t *mutex)
+{
+	if (pthread_mutex_unlock(mutex) != 0)
+		abort_perror("pthread_mutex_unlock");
 }
 
 /**
@@ -97,6 +110,8 @@ char *levels[] = {
 	"DEFAULT",
 };
 
+struct daemon_config;
+
 /**
  * A struct for tracking running VPN processes.
  * @next: linked list
@@ -109,6 +124,10 @@ struct vpn_manager {
 	pid_t openvpn_pid; /* only valid in the parent, obviously */
 	int fildes[2];
 	char ifname[IFNAMSIZ];
+	struct daemon_config *dc; // so we can log
+	pthread_t thread;
+	int id;
+	FILE *stdout;
 };
 
 /**
@@ -132,6 +151,10 @@ struct detour_manager {
  * @vpns: list of vpns
  * @detours: list of detours
  * @sk: netlink socket we listen on
+ * @logfile: where to write log entries
+ * @daemonize: whether or not to fork a new process first
+ * @loglevel: filters log entries
+ * @log_lock: held by _dc_log function to avoid concurrency. problems caused by
  */
 struct daemon_config {
 	struct vpn_manager *vpns;
@@ -140,6 +163,7 @@ struct daemon_config {
 	FILE *logfile;
 	int daemonize;
 	int loglevel;
+	pthread_mutex_t log_lock;
 };
 
 void _dc_log(struct daemon_config *dc, int level, const char *format, ...)
@@ -152,6 +176,7 @@ void _dc_log(struct daemon_config *dc, int level, const char *format, ...)
 	if (level > dc->loglevel)
 		return;
 
+	pthread_mutex_lock_s(&dc->log_lock);
 	t = time(NULL);
 	tm = localtime(&t);
[…]

### 2017-04-27 pkg.sh
Use JSON output on client and a little sleep sync to make things work

@@ -4,7 +4,6 @@ rm -f vm.tar
 tar cf vm.tar \
 	src/* \
 	etc/* \
-	tmp/linux-firmware-image.deb \
-	tmp/linux-image.deb \
+	tmp/* \
 	setup.sh
 echo Package created!

### 2017-04-27 src/experiment.py

@@ -191,7 +191,7 @@ def scenario(name, params, trials=30):
     SETUP[params['scenario']](client, detour)
 
     filename = '%s.%s.json' % (name, params['scenario'])
-    server.sendCmd('iperf3 -s -J | tee %s' % filename)
+    server.sendCmd('iperf3 -s -J > %s' % filename)
 
     # sleep synchronization is the worst, except for iperf
     import time; time.sleep(0.5)
@@ -199,8 +199,10 @@ def scenario(name, params, trials=30):
     for _ in range(trials):
         print('.', end='')
         sys.stdout.flush()
-        client.cmd('iperf3 -c ' + server.IP())
+        client.cmd('iperf3 -c ' + server.IP() + ' -J')
+        time.sleep(0.5)
     print()
+    time.sleep(0.5)
 
     mn.stop()
 

### 2017-04-27 src/experiment.py
Allow experiments to run without custom kernel

@@ -22,8 +22,11 @@ containing only headers is created for future inspection.
 """
 from __future__ import print_function
 
+import argparse
 import os
+import subprocess
 import sys
+import time
 
 from mininet.cli import CLI
 from mininet.log import lg, LEVELS
@@ -164,17 +167,36 @@ def DetourNet(params):
     return mn
 
 
-def setup_nat(client, detour):
-    client.cmd('./client daemon ../etc/daemon-nat-vm.conf')
+def is_custom_kernel():
+    output = subprocess.check_output(['dmesg'])
+    return 'mptcp' in output and 'detour' in output
+
+
+def setup_nat(net):
+    client, detour, server = net.get('client', 'detour', 'server')
     detour.cmd('./nat_detour.py &')
+    if is_custom_kernel():
+        client.cmd('./client daemon ../etc/daemon-nat-vm.conf')
+    else:
+        time.sleep(0.5)
+        client.cmd('./request.py %s %s %d %d' % (detour.IP(), server.IP(),
+                                                 5201, 5201))
+        return 'iperf3 -c ' + detour.IP() + ' -J'
 
 
-def setup_vpn(client, detour):
-    client.cmd('./client daemon ../etc/daemon-vpn-vm.conf')
+def setup_vpn(net):
+    client, detour, server = net.get('client', 'detour', 'server')
     detour.cmd('./vpn_detour.sh &')
+    if is_custom_kernel():
+        client.cmd('./client daemon ../etc/daemon-vpn-vm.conf')
+    else:
+        client.cmd('openvpn --remote %s --client --dev tun --ca ../tmp/ca.crt '
+                   '--cert ../tmp/client1.crt --key ../tmp/client1.key '
+                   '--topology p2p --pull --nobind' % detour.IP())
+        return 'iperf3 -c ' + server.IP() + ' -J -B 10.8.0.4'
 
 
-def setup_ctrl(client, detour):
+def setup_ctrl(net):
     pass
 
 
@@ -185,57 +207,76 @@ SETUP = {
 }
 
 
-def scenario(name, params, trials=30):
-    mn = DetourNet(params)
-    client, detour, server = mn.get('client', 'detour', 'server')
-    SETUP[params['scenario']](client, detour)
+def params_easy():
+    return BASIC_PARAMS.copy()
 
-    filename = '%s.%s.json' % (name, params['scenario'])
+
+def params_lossy():
[…]

### 2017-04-27 src/request.py

@@ -1,4 +1,4 @@
-#!/usr/bin/env python
+#!/usr/bin/env python3
 """
 Send requests to the mproxy daemon. Must run from client.
 

### 2017-04-27 data/output.py
JSON Parsing utilities

@@ -1,0 +1,47 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""Parse JSON output of experiments."""
+
+import json
+
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def json_parse(filename):
+    """Parse a file that contains a bunch of JSON objects.
+
+    This assumes that each object starts and ends
+    """
+    with open(filename) as f:
+        string = ''
+        for line in f:
+            string += line
+            try:
+                yield json.loads(string)
+                string = ''
+            except:
+                pass
+
+
+def objects_to_bps(obj_list):
+    """Converts a sequence of JSON objects to the bits per second."""
+    for obj in obj_list:
+        bps = obj.get('end', {}).get('sum_received', {}).get('bits_per_second')
+        if bps:
+            yield bps
+
+
+def get_numpy_array(filename):
+    return np.array(list(objects_to_bps(json_parse(filename))))
+
+
+def plot_comparison(name):
+    ctrl = get_numpy_array('%s.control.json' % name) / 1000000
+    nat = get_numpy_array('%s.nat.json' % name) / 1000000
+    vpn = get_numpy_array('%s.vpn.json' % name) / 1000000
+    print(len(ctrl), len(nat), len(vpn))
+    fig, ax = plt.subplots()
+    ax.boxplot([ctrl, nat, vpn], labels=['Control', 'NAT', 'VPN'])
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison (%s)' % name)

### 2017-04-27 src/experiment.py
oops print

@@ -242,7 +242,7 @@ def scenario(setup_name, params_name, trials=30):
         print('.', end='')
         sys.stdout.flush()
         if iperf_cmd:
-            print(client.cmd(iperf_cmd))
+            client.cmd(iperf_cmd)
         else:
             client.cmd('iperf3 -c ' + server.IP() + ' -J')
         time.sleep(0.5)

### 2017-04-27 src/experiment.py
Makes openvpn work on vanilla kernel.

@@ -24,6 +24,7 @@ from __future__ import print_function
 
 import argparse
 import os
+import re
 import subprocess
 import sys
 import time
@@ -190,10 +191,27 @@ def setup_vpn(net):
     if is_custom_kernel():
         client.cmd('./client daemon ../etc/daemon-vpn-vm.conf')
     else:
-        client.cmd('openvpn --remote %s --client --dev tun --ca ../tmp/ca.crt '
-                   '--cert ../tmp/client1.crt --key ../tmp/client1.key '
-                   '--topology p2p --pull --nobind' % detour.IP())
-        return 'iperf3 -c ' + server.IP() + ' -J -B 10.8.0.4'
+        client.cmd('openvpn --remote %s 1194 udp --client --dev tun '
+                   '--ca ../tmp/ca.crt --cert ../tmp/client1.crt '
+                   '--key ../tmp/client1.key ' '--topology p2p --pull '
+                   '--nobind &' % detour.IP())
+
+        # We need to wait until the openvpn interface is up. This will also
+        # give us the peer IP address, which we already configured, but it's
+        # always better to dynamically determine.
+        output = client.monitor()
+        regexp = re.compile(r'ip addr add dev tun0 local ([0-9.]+) peer ([0-9.]+)')
+        while not regexp.search(output):
+            output += client.monitor()
+        match = regexp.search(output)
+        local, peer = match.groups()
+        # For some reason even after openvpn says the interface is up, we need
+        # to keep waiting a bit longer before we can use it as a route gateway.
+        time.sleep(0.5)
+        # We have to add a very specific routing rule here so that the iperf
+        # traffic goes through OpenVPN. This is, unfortunately, necessary since
+        # the iperf3 -B option does not seem to work in this case.
+        client.cmd('route add -host %s gw %s' % (server.IP(), peer))
 
 
 def setup_ctrl(net):

### 2017-04-27 src/experiment.py
This should have been how it was the whole time :/

@@ -41,8 +41,8 @@ BASIC_PARAMS = {
     'client_r1': {'delay': '5ms', 'bw': 20},
     'r1_r2': {'delay': '5ms', 'bw': 10},
     'r2_server': {'delay': '5ms', 'bw': 20},
-    'r1_r3': {'delay': '5ms', 'bw': 20},
-    'r2_r3': {'delay': '5ms', 'bw': 20},
+    'r1_r3': {'delay': '5ms', 'bw': 10},
+    'r2_r3': {'delay': '5ms', 'bw': 10},
     'r3_detour': {'delay': '5ms', 'bw': 20},
 }
 

### 2017-04-28 data/output.py
Fix output code for more vanilla scenarios

@@ -37,11 +37,17 @@ def get_numpy_array(filename):
 
 
 def plot_comparison(name):
-    ctrl = get_numpy_array('%s.control.json' % name) / 1000000
-    nat = get_numpy_array('%s.nat.json' % name) / 1000000
-    vpn = get_numpy_array('%s.vpn.json' % name) / 1000000
-    print(len(ctrl), len(nat), len(vpn))
+    mptcp_ctrl = get_numpy_array('%s.control.mptcp.json' % name) / 1000000
+    mptcp_nat = get_numpy_array('%s.nat.mptcp.json' % name) / 1000000
+    mptcp_vpn = get_numpy_array('%s.vpn.mptcp.json' % name) / 1000000
+    vanilla_ctrl = get_numpy_array('%s.control.vanilla.json' % name) / 1000000
+    vanilla_nat = get_numpy_array('%s.nat.vanilla.json' % name) / 1000000
+    vanilla_vpn = get_numpy_array('%s.vpn.vanilla.json' % name) / 1000000
     fig, ax = plt.subplots()
-    ax.boxplot([ctrl, nat, vpn], labels=['Control', 'NAT', 'VPN'])
+    data = [mptcp_ctrl, mptcp_nat, mptcp_vpn, vanilla_ctrl, vanilla_nat,
+            vanilla_vpn]
+    labels = ['1 Subflow', 'NAT', 'VPN', 'TCP', 'TCP(NAT)', 'TCP(VPN)']
+    print(map(len, data))
+    ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
     ax.set_title('Throughput Comparison (%s)' % name)

### 2017-04-28 src/experiment.py
Unused imports

@@ -30,11 +30,9 @@ import sys
 import time
 
 from mininet.cli import CLI
-from mininet.log import lg, LEVELS
 from mininet.net import Mininet
 from mininet.topo import Topo
 from mininet.link import TCLink
-from mininet.util import waitListening
 
 
 BASIC_PARAMS = {

### 2017-04-28 src/vido_init.sh
I swear, I came up with this fix myself

@@ -1,0 +1,21 @@
+#!/bin/sh
+#
+# Fix for using iperf3 within virtual machines run with vido!
+#
+# Due to some funky business regarding the v9fs (used to access the host fs) and
+# the unlink + ftruncate system calls used in iperf3, iperf3 will fail to run on
+# the client, citing that it could not create a stream (No such file or
+# directory). The reason is almost certainly very interesting, but ultimately
+# not relevant. We can work around it by mounting our own ramfs at /tmp, so that
+# when iperf3 tries to make its own ramfs, it will get our tmpfs, rather than
+# talking to v9fs and in turn talking to the host OS.
+#
+# This email references the unlink followed by X problem in V9, and it seems
+# that the fix is not ready:
+# https://lkml.org/lkml/2016/6/22/358
+#
+# Additionally, this GitHub issue proposes the same fix:
+# https://github.com/01org/cc-oci-runtime/issues/152
+# However, I came up with the same fix independently :D
+#
+mount -t ramfs -o size=256M ramfs /tmp

### 2017-05-04 pkg.sh
compress pkg

@@ -1,7 +1,7 @@
 #!/usr/bin/sh
 make -C src clean
 rm -f vm.tar
-tar cf vm.tar \
+tar cJf vm.tar.xz \
 	src/* \
 	etc/* \
 	tmp/* \

### 2017-05-04 etc/aws_exp_setup.sh
aws setup script

@@ -1,0 +1,39 @@
+# Run as root.
+cd
+
+export DEBIAN_FRONTEND=noninteractive
+
+# We'll use the same tarball as the VMs but we won't use the bundled setup.sh
+echo Downloading vm tarball...
+wget -q https://brenns10.keybase.pub/vm.tar.xz
+chmod a+rx vm.tar.xz
+chmod a+x .
+echo Extracting vm tarball...
+sudo -u ubuntu tar xf vm.tar.xz -C ~ubuntu
+
+echo Updating package database...
+apt update -q
+echo Upgrading packages...
+apt upgrade -y -q
+echo Installing required packages...
+apt install -y -q build-essential python3-pip libnl-genl-3-dev openvpn iperf3 tmux libconfig-dev
+echo Installing Python packages...
+pip3 install psutil
+echo Installing custom kernel...
+dpkg -i ~ubuntu/tmp/*.deb
+
+echo Setting default kernel...
+SUBMENU=$(egrep -o -m 1 "gnulinux-advanced[0-9a-zA-Z-]+" /boot/grub/grub.cfg)
+BOOT_OPTION=$(egrep -o -m 1 "gnulinux-4.1.38-custom-advanced[0-9a-zA-Z-]+" /boot/grub/grub.cfg)
+sed -i.bak "s/#\?GRUB_DEFAULT.*/GRUB_DEFAULT=\"$SUBMENU>$BOOT_OPTION\"/g" /etc/default/grub
+update-grub
+
+echo Building tools...
+sudo -u ubuntu make -C ~ubuntu/src clean
+sudo -u ubuntu make -C ~ubuntu/src all
+
+echo Adding ssh keys...
[…]

### 2017-05-05 etc/aws_exp_setup.sh
AWS Experiment updates - lets go EC2

@@ -1,3 +1,4 @@
+#!/bin/bash
 # Run as root.
 cd
 
@@ -35,5 +36,5 @@ sudo -u ubuntu make -C ~ubuntu/src all
 echo Adding ssh keys...
 echo "ssh-rsa AAAAB3NzaC1yc2EAAAADAQABAAABAQDEupr15Lbq8jFBca/tXhydvD2nKQ/vWnxXzOndG7TqKGy53mtlgR6VzaV4WR3blrLdDG223a64VlQaj00xm+8e/eAn+IRH3RpV2ZvFT+BdZmej8E4y0sSlkiCw2sjtyKiGu0Pyk2HKbA8njG+aVsJ5YSQVwBBP1nEqKoMItvXzZwQJQ09SCeklT5zugcJz+frGFifjz1I6bs7Hb9vNa7hNi79JjJ1rNoeANkBmunW20+M7hMOuNjGWl47XGDlew7tn3vc2Eiiohf0cyrueBPq63Y7tMEfhmyCEPfakkjg5idt3soRGKCBqyoVuYyEIOJSTnuGBkThZteGHLlpKvMVz stephen@greed_2013-09-26" >> ~ubuntu/.ssh/authorized_keys
 
-echo Completed successfully! Rebooting...
-reboot
+echo Completed successfully! Shutting down...
+shutdown -h now

### 2017-05-05 src/aws_experiment.py

@@ -1,0 +1,172 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""AWS Detour Experiment
+
+This script creates several EC2 instances in different datacenters. It does the
+necessary initialization to each of the hosts, waits for their initialization
+to complete, and then runs some experiments on them.
+
+Prerequisites:
+1. You must have an SSH key pair. You must import it into EC2 on *each* region
+   the experiment will run. While this could be automated, I would prefer not
+   to bother with APIs that center around one-off configuration.
+2. The default security group of the default VPC ought to be configured to
+   to allow ingress on any port. By default it allows ingress from other
+   members of the security group, egress to anywhere. If nothing else, add SSH
+   ingress. It's just much more convenient. And I don't want to do it
+   programmatically.
+
+"""
+import collections
+from os.path import abspath
+from os.path import dirname
+from os.path import join
+from pprint import pprint
+
+import boto3
+
+
+# Key Pair Name. Hopefully you named it the same thing in every region.
+KEYPAIR = 'stephen@greed'
+
+# The filename of the AWS setup script.
+SETUP_SCRIPT = join(dirname(abspath(__file__)), '../etc/aws_exp_setup.sh')
+
+# These tags are applied to everything we create, so it's easier to clean up
+# via the Console if things go wrong.
+TAGS = [
+    {'Key': 'Project', 'Value': 'Thesis'},
+]
+
+
+# These become arguments to boto3.resource('ec2')
+RESOURCE_PARAMS = {}
+RESOURCE_PARAMS_EACH = {
+    'client': {
+        'region_name': 'us-west-1', # N. California
+    },
+    'detour': {
+        'region_name': 'us-east-2', # Ohio
+    },
+    'server': {
+        'region_name': 'us-east-1', # N. Virginia
+    },
+}
+
+# These become arguments to ec2.create_instances()
+CREATE_INSTANCES_PARAMS = {
+    'MinCount': 1,
+    'MaxCount': 1,
+    'UserData': open(SETUP_SCRIPT).read(),
+    'InstanceType': 't2.micro',
+    'TagSpecifications': [
+        {
+            'ResourceType': 'instance',
+            'Tags': TAGS,
+        },
+    ],
+    'KeyName': KEYPAIR,
+}
+CREATE_INSTANCES_PARAMS_EACH = {
+    'client': {
+        'ImageId': 'ami-30476250',
+    },
+    'detour': {
+        'ImageId': 'ami-33ab8f56',
+    },
+    'server': {
[…]

### 2017-05-05 src/aws_experiment.py
Add and run experiment. :/

@@ -18,16 +18,30 @@ Prerequisites:
 
 """
 import collections
+import sys
+from io import BytesIO
 from os.path import abspath
 from os.path import dirname
 from os.path import join
 from pprint import pprint
 
 import boto3
+import paramiko
 
+
+NAT_CFG = '''
+client: {
+  detours = ["%s"];
+  vpns = [];
+  daemonize = false;
+  logfile = "daemon-nat.log";
+  loglevel = "DEFAULT";
+}
+'''
 
 # Key Pair Name. Hopefully you named it the same thing in every region.
 KEYPAIR = 'stephen@greed'
+KEYFILE = '/home/stephen/.ssh/id_rsa'
 
 # The filename of the AWS setup script.
 SETUP_SCRIPT = join(dirname(abspath(__file__)), '../etc/aws_exp_setup.sh')
@@ -36,6 +50,9 @@ SETUP_SCRIPT = join(dirname(abspath(__file__)), '../etc/aws_exp_setup.sh')
 # via the Console if things go wrong.
 TAGS = [
     {'Key': 'Project', 'Value': 'Thesis'},
+]
+TAG_FILTER = [
+    {'Name': 'tag:Project', 'Values': ['Thesis']}
 ]
 
 
@@ -150,23 +167,130 @@ def first_time_wait(instances):
         inst.wait_until_running()
 
 
+def ssh_connect(instances):
+    print('Connecting to all instances via ssh...')
+    clients = {}
+    for k, inst in instances.items():
+        print('Connecting to %s' % k)
+        while True:
+            c = paramiko.SSHClient()
+            c.load_system_host_keys()
+            c.set_missing_host_key_policy(paramiko.AutoAddPolicy())
+            try:
+                c.connect(inst.public_ip_address, username='ubuntu',
+                          key_filename=KEYFILE)
+                clients[k] = c
+                print('Connected to %s (%s)' % (k, inst.public_ip_address))
+                break
+            except:
+                print('Failed to connect. Retrying...')
+                pass
+    return clients
+
+
+def many_iperf(n, client_transport, server_chan, server_output, instances):
+    print('IPERF: ', end='')
+    sys.stdout.flush()
+    for _ in range(n):
+        client_chan = client_transport.open_session()
+        client_chan.set_combine_stderr(True)
+        client_chan.exec_command(command='iperf3 -c %s -J' %
+                                 instances['server'].public_ip_address)
+        print('<', end='')
+        sys.stdout.flush()
+        while not client_chan.exit_status_ready():
[…]

### 2017-05-16 src/client.c
client: Add threading for receiving responses.
I'm not sure if I would feel okay defending when the client program I wrote
could have been crippled simply by a single detour not responding to a request.

@@ -3,7 +3,21 @@
  * build: make client
  * run: ./client CMD [ARGS ...]
  *
- * TODO: if scalability is required/desired, we need threading
+ * This has grown into a large, complex program. It has commands for sending
+ * one-off messages to the kernel. However its main function is the `daemon`
+ * command, which runs in the background, managing OpenVPN and handling mproxy
+ * detour requests and responses.
+ *
+ * The daemon mode is multithreaded:
+ * - The main thread parses configuration and spawns the other threads. It then
+ *   loops waiting for netlink detour requests. It reads on netlink socket and
+ *   writes on the mproxy sockets.
+ * - The recv_thread listens for mproxy responses and reports them to the
+ *   kernel. This is only thread safe because the detour_manager list is not
+ *   modified during normal operation, and send/recv operations are thread safe
+ *   in POSIX. It reads on mproxy sockets and writes on the netlink socket.
+ * - For each OpenVPN instance there is a managing thread. It captures output
+ *   and logs it, ensuring we don't miss important information.
  */
 
 #include <stdio.h>
@@ -18,6 +32,7 @@
 #include <signal.h>
 #include <wait.h>
 #include <sys/socket.h>
+#include <sys/select.h>
 #include <netinet/in.h>
 #include <arpa/inet.h>
 #include <regex.h>
@@ -82,6 +97,16 @@ ssize_t indexof(char **arr, size_t nelem, const char *str)
  * Macro for getting the size of a static array.
  */
 #define nelem(arr) (sizeof(arr) / sizeof(arr[0]))
+
+/**
+ * Macro for getting the minimum of two things
+ */
+#define min(x, y) ((x) <= (y) ? (x) : (y))
+
+/**
+ * Macro for getting the maximum of two things
+ */
+#define max(x, y) ((x) > (y) ? (x) : (y))
 
 /**
  * Log levels - messages with level <= the set level are logged. Note similarity
@@ -155,6 +180,7 @@ struct detour_manager {
  * @daemonize: whether or not to fork a new process first
  * @loglevel: filters log entries
[…]

### 2017-05-16 pkg.sh
update pkg back to .tar and add "manifest" support

@@ -1,9 +1,10 @@
 #!/usr/bin/sh
 make -C src clean
 rm -f vm.tar
-tar cJf vm.tar.xz \
-	src/* \
-	etc/* \
-	tmp/* \
-	setup.sh
+
+FILES="src/** etc/** tmp/** setup.sh"
+tar cf vm.tar $FILES
+sha256sum $FILES vm.tar > sha256sums.txt
+
 echo Package created!
+echo Manifest in sha256sums.txt

### 2017-05-17 pkg.sh
more pkg changes

@@ -3,8 +3,9 @@ make -C src clean
 rm -f vm.tar
 
 FILES="src/** etc/** tmp/** setup.sh"
-tar cf vm.tar $FILES
+tar cvf vm.tar $FILES
 sha256sum $FILES vm.tar > sha256sums.txt
 
+echo
 echo Package created!
 echo Manifest in sha256sums.txt

### 2017-05-17 pkg.sh
final pkg change

@@ -4,7 +4,8 @@ rm -f vm.tar
 
 FILES="src/** etc/** tmp/** setup.sh"
 tar cvf vm.tar $FILES
-sha256sum $FILES vm.tar > sha256sums.txt
+sha256sum $FILES > sha256sums.txt
+tar -xOf vm.tar | sha256sum >> sha256sums.txt
 
 echo
 echo Package created!

### 2017-05-17 etc/aws_exp_setup.sh
fix aws_exp_setup.sh to use vm.tar

@@ -3,14 +3,15 @@
 cd
 
 export DEBIAN_FRONTEND=noninteractive
+TARNAME=vm.tar
 
 # We'll use the same tarball as the VMs but we won't use the bundled setup.sh
 echo Downloading vm tarball...
-wget -q https://brenns10.keybase.pub/vm.tar.xz
-chmod a+rx vm.tar.xz
+wget -q https://brenns10.keybase.pub/$TARNAME
+chmod a+rx $TARNAME
 chmod a+x .
 echo Extracting vm tarball...
-sudo -u ubuntu tar xf vm.tar.xz -C ~ubuntu
+sudo -u ubuntu tar xf $TARNAME -C ~ubuntu
 
 echo Updating package database...
 apt update -q

### 2017-05-17 src/experiment.py
experiment: use lia congestion control rather than cubic in MPTCP

@@ -176,6 +176,8 @@ def setup_nat(net):
     detour.cmd('./nat_detour.py &')
     if is_custom_kernel():
         client.cmd('./client daemon ../etc/daemon-nat-vm.conf')
+        client.cmd('sysctl -w net.ipv4.tcp_congestion_control=lia')
+        server.cmd('sysctl -w net.ipv4.tcp_congestion_control=lia')
     else:
         time.sleep(0.5)
         client.cmd('./request.py %s %s %d %d' % (detour.IP(), server.IP(),
@@ -188,6 +190,8 @@ def setup_vpn(net):
     detour.cmd('./vpn_detour.sh &')
     if is_custom_kernel():
         client.cmd('./client daemon ../etc/daemon-vpn-vm.conf')
+        client.cmd('sysctl -w net.ipv4.tcp_congestion_control=lia')
+        server.cmd('sysctl -w net.ipv4.tcp_congestion_control=lia')
     else:
         client.cmd('openvpn --remote %s 1194 udp --client --dev tun '
                    '--ca ../tmp/ca.crt --cert ../tmp/client1.crt '

### 2017-05-17 src/experiment.py
experiment: add several more params

@@ -228,18 +228,73 @@ SETUP = {
 
 
 def params_easy():
+    """
+    In this scenario, the core links are 10 Mbit while the access links are 20.
+    This means that detour routing could actually aggregate more throughput
+    than a single path could.
+    """
     return BASIC_PARAMS.copy()
 
 
 def params_lossy():
+    """
+    This scenario is similar to 'easy', but there is loss along the default
+    route.
+    """
     d = BASIC_PARAMS.copy()
     d['r1_r2']['loss'] = 1
+    return d
+
+
+def params_delayed():
+    """
+    This scenario is similar to 'easy', but there is a high latency link across
+    the default route.
+    """
+    d = BASIC_PARAMS.copy()
+    d['r1_r2']['delay'] = '100ms'
+    return d
+
+
+def params_sym():
+    """
+    In this scenario, each link has 10 Mbit bandwidth. As a result, there may
+    not be any gain by aggregating paths.
+    """
+    d = BASIC_PARAMS.copy()
+    d['client_r1']['bw'] = 10
+    d['r2_server']['bw'] = 10
+    d['r3_detour']['bw'] = 10
+    return d
+
+
+def params_sym_lossy():
+    """
+    This scenario, based on symmetric, has each link with the same bandwidth.
+    However, the default route is lossy. Thus, the two flows are competing for
+    bandwidth and the alternative subflow needs to make up the total bandwidth.
+    """
+    d = params_symmetric()
+    d['r1_r2']['loss'] = 1
+    return d
+
+
+def params_sym_delayed():
+    """
+    This scenario is based on symmetric, but the default route has high delay.
+    """
+    d = params_symmetric()
+    d['r1_r2']['delay'] = '100ms'
     return d
 
 
 PARAMS = {
     'easy': params_easy,
     'lossy': params_lossy,
+    'delayed': params_delayed,
+    'sym': params_sym,
+    'sym_lossy': params_sym_lossy,
+    'sym_delayed': params_sym_delayed,
 }
 
 

### 2017-05-17 src/experiment.py
experiment: more trials and more params

@@ -331,13 +331,14 @@ def scenario(setup_name, params_name, trials=30):
 
 def main():
     parser = argparse.ArgumentParser(description='Mininet experiment harness')
-    parser.add_argument('--params', type=str, default='easy,lossy',
+    parser.add_argument('--params', type=str,
+                        default='easy,lossy,delayed,sym,sym_lossy,sym_delayed',
                         help='comma separated list of params')
     parser.add_argument('--setups', type=str, default='control,nat,vpn',
                         help='comma separated list of setups')
     parser.add_argument('--cli', action='store_true',
                         help='start a cli with the first param and setup')
-    parser.add_argument('--trials', '-t', type=int, default=30)
+    parser.add_argument('--trials', '-t', type=int, default=60)
 
     args = parser.parse_args()
     params = args.params.split(',')

### 2017-05-17 src/experiment.py
experiment: who has two thumbs and writes bugs?
https://media.giphy.com/media/pU7ZQcEQ8Loxq/giphy.gif

@@ -274,7 +274,7 @@ def params_sym_lossy():
     However, the default route is lossy. Thus, the two flows are competing for
     bandwidth and the alternative subflow needs to make up the total bandwidth.
     """
-    d = params_symmetric()
+    d = params_sym()
     d['r1_r2']['loss'] = 1
     return d
 
@@ -283,7 +283,7 @@ def params_sym_delayed():
     """
     This scenario is based on symmetric, but the default route has high delay.
     """
-    d = params_symmetric()
+    d = params_sym()
     d['r1_r2']['delay'] = '100ms'
     return d
 

### 2017-05-17 src/client.c
client: improve daemonize comment with additional research

@@ -770,6 +770,12 @@ void destroy_daemon_config(struct daemon_config *dc)
  * case, the parent is going to exit immediately after forking anyway, so the
  * second fork (whose purpose is to avoid creating a zombie process when the
  * parent continues running) is unnecessary.
+ *
+ * The other purpose of a second fork is to create a new session with setsid(),
+ * (in the first child), and then fork again, killing the first child. This way
+ * the second child is not the leader of its own session, and it cannot
+ * accidentally acquire a terminal and get terminated by it.
+ * http://stackoverflow.com/questions/881388/what-is-the-reason-for-performing-a-double-fork-when-creating-a-daemon
  */
 void daemonize(void)
 {

### 2017-05-17 src/aws_experiment.py
aws_experiment: lets hope OpenVPN is kind to us

@@ -39,6 +39,16 @@ client: {
 }
 '''
 
+VPN_CFG = '''
+client: {
+  detours = [];
+  vpns = ["%s"];
+  daemonize = false;
+  logfile = "daemon-nat.log";
+  loglevel = "DEFAULT";
+}
+'''
+
 # Key Pair Name. Hopefully you named it the same thing in every region.
 KEYPAIR = 'stephen@greed'
 KEYFILE = '/home/stephen/.ssh/id_rsa'
@@ -235,20 +245,24 @@ def run_exp_nat(instances, clients):
     many_iperf(30, client_transport, server_chan, server_output, instances)
 
     # save control server output
-    with open('server1.json', 'wb') as f:
+    with open('server.ctrl.json', 'wb') as f:
         f.write(server_output.getvalue())
 
     # send a client daemon config
     print('Sending over the daemon config...')
     sftp = clients['client'].open_sftp()
     nat_cfg = NAT_CFG % instances['detour'].public_ip_address
+    vpn_cfg = VPN_CFG % instances['detour'].public_ip_address
     sftp.putfo(BytesIO(nat_cfg.encode('utf8')),
                '/home/ubuntu/etc/daemon-nat-aws.conf')
+    sftp.putfo(BytesIO(vpn_cfg.encode('utf8')),
+               '/home/ubuntu/etc/daemon-vpn-aws.conf')
     sftp.close()
 
     # run the client daemon
     print('Starting up the client daemon...')
     client_daemon_chan = client_transport.open_session()
+    client_daemon_chan.get_pty() # so we can SIGHUP it
     client_daemon_chan.set_combine_stderr(True)
     client_daemon_chan.exec_command(
         command='cd src; sudo ./client daemon ../etc/daemon-nat-aws.conf'
@@ -258,7 +272,22 @@ def run_exp_nat(instances, clients):
     many_iperf(30, client_transport, server_chan, server_output, instances)
 
     # save nat server output
-    with open('server2.json', 'wb') as f:
+    with open('server.nat.json', 'wb') as f:
+        f.write(server_output.getvalue())
+
+    # kill the client daemon and start a vpn one
+    client_daemon_chan.close()
+    client_daemon_chan = client_transport.open_session()
+    client_daemon_chan.set_combine_stderr(True)
+    client_daemon_chan.exec_command(
+        command='cd src; sudo ./client daemon ../etc/daemon-vpn-aws.conf'
+    )
+
+    server_output = BytesIO()
[…]

### 2017-05-17 src/aws_experiment.py
aws_experiment: fixes and data for aws openvpn

@@ -44,7 +44,7 @@ client: {
   detours = [];
   vpns = ["%s"];
   daemonize = false;
-  logfile = "daemon-nat.log";
+  logfile = "daemon-vpn.log";
   loglevel = "DEFAULT";
 }
 '''
@@ -222,17 +222,29 @@ def many_iperf(n, client_transport, server_chan, server_output, instances):
     print()
 
 
+def blocking_cmd(ssh, cmd):
+    chan = ssh.open_session()
+    chan.set_combine_stderr(True)
+    chan.exec_command(command=cmd)
+    chan.recv_exit_status()
+
+
 def run_exp_nat(instances, clients):
     detour_transport = clients['detour'].get_transport()
     client_transport = clients['client'].get_transport()
     server_transport = clients['server'].get_transport()
 
     # start detour daemon
-    print('Starting detour daemon...')
+    print('Starting detour nat daemon...')
     detour_chan = detour_transport.open_session()
     detour_chan.set_combine_stderr(True)
     detour_chan.exec_command(command='cd src; sudo ./nat_detour.py')
     print(detour_chan.recv(4096))
+    print('Starting detour vpn daemon...')
+    detour_vpn_chan = detour_transport.open_session()
+    detour_vpn_chan.set_combine_stderr(True)
+    detour_vpn_chan.exec_command(command='cd src; sudo ./vpn_detour.sh')
+    print(detour_vpn_chan.recv(4096))
 
     # start server
     print('Starting iperf server...')
@@ -249,7 +261,7 @@ def run_exp_nat(instances, clients):
         f.write(server_output.getvalue())
 
     # send a client daemon config
-    print('Sending over the daemon config...')
+    print('Sending over the daemon configs...')
     sftp = clients['client'].open_sftp()
     nat_cfg = NAT_CFG % instances['detour'].public_ip_address
     vpn_cfg = VPN_CFG % instances['detour'].public_ip_address
@@ -260,7 +272,7 @@ def run_exp_nat(instances, clients):
     sftp.close()
 
     # run the client daemon
-    print('Starting up the client daemon...')
+    print('Starting up the nat daemon...')
     client_daemon_chan = client_transport.open_session()
     client_daemon_chan.get_pty() # so we can SIGHUP it
     client_daemon_chan.set_combine_stderr(True)
@@ -275,8 +287,18 @@ def run_exp_nat(instances, clients):
     with open('server.nat.json', 'wb') as f:
         f.write(server_output.getvalue())
 
[…]

### 2017-05-18 data/2017-05-17-extended/plots.py
plots for the data

@@ -1,0 +1,57 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved (thanks python 3.6...?)
+META = {
+    'easy': 'Core-limited',
+    'lossy': 'Core-limited with Loss',
+    'delayed': 'Core-limited with High Latency',
+    'sym': 'Symmetric',
+    'sym_lossy': 'Symmetric with Loss',
+    'sym_delayed': 'Symmetric with High Latency',
+}
+
+# order preserved
+SCEN = {
+    'control.mptcp': '1 Subflow',
+    'nat.mptcp': 'NAT',
+    'vpn.mptcp': 'VPN',
+    'control.vanilla': 'TCP',
+    'nat.vanilla': 'TCP (NAT)',
+    'vpn.vanilla': 'TCP (VPN)',
+}
+
+
+def make_single_plot(base):
+    print(base)
+    data = []
+    labels = []
+    for s, label in SCEN.items():
+        data.append(o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000)
+        labels.append(label)
+
+    fig, ax = plt.subplots()
+    print(list(map(len, data)))
+    ax.boxplot(data, 0, '', labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: ' + META[base])
+    fig.savefig('%s.pdf' % base)
+
+
+def make_all_plots():
+    for base in META.keys():
+        make_single_plot(base)
+
+
+if __name__ == '__main__':
+    make_all_plots()

### 2017-05-18 data/2017-05-17-extended/plots.py
A new type of plot, showing throughput vs time. Looking good!
I'm committing this now for safety, but also because I want to make sure I write
down my concern. I'm looking at gridded comparisons of every time series I
produced, and sym.control.* ones look funky. The corresponding easy.control.*
time series are not nearly as noisy, and they're nearly straight line. I'm
investigating the possibility that the noise is due to something in my
experiment harness. I'm doing this by rebooting the VM and running
sym.control.mptcp again. If it looks good as the first one, I'm probably okay.

@@ -53,5 +53,31 @@ def make_all_plots():
         make_single_plot(base)
 
 
+def make_time_series_grid(base):
+    print(base)
+    rows = ['mptcp', 'vanilla']
+    cols = ['control', 'nat', 'vpn']
+    fig, axarr = plt.subplots(len(rows), len(cols), sharex=True, sharey=True)
+    for ridx, row in enumerate(rows):
+        for cidx, col in enumerate(cols):
+            data = o.get_time_series_arrays('%s.%s.%s.json' % (base, col, row))
+            for series in data:
+                axarr[ridx][cidx].plot(series[:10])
+    axarr[0][0].set_ylabel('MPTCP')
+    axarr[1][0].set_ylabel('TCP')
+    axarr[1][0].set_xlabel('Control')
+    axarr[1][1].set_xlabel('NAT')
+    axarr[1][2].set_xlabel('VPN')
+    fig.savefig('timegrid.%s.pdf' % base)
+
+
+def make_all_time_series_grid():
+    for base in META.keys():
+        make_time_series_grid(base)
+
+
 if __name__ == '__main__':
-    make_all_plots()
+    if len(sys.argv) > 1 and sys.argv[1] == 'time':
+        make_all_time_series_grid()
+    else:
+        make_all_plots()

### 2017-05-18 data/output.py

@@ -36,6 +36,14 @@ def get_numpy_array(filename):
     return np.array(list(objects_to_bps(json_parse(filename))))
 
 
+def time_series(obj):
+    return np.array([o['sum']['bits_per_second'] for o in obj['intervals']]) / 1000000
+
+
+def get_time_series_arrays(filename):
+    return list(map(time_series, json_parse(filename)))
+
+
 def plot_comparison(name):
     mptcp_ctrl = get_numpy_array('%s.control.mptcp.json' % name) / 1000000
     mptcp_nat = get_numpy_array('%s.nat.mptcp.json' % name) / 1000000

### 2017-05-18 src/mn_experiment.sh
Fix odd stupidity with noisy data in previous runs.

@@ -1,0 +1,26 @@
+#!/bin/sh
+
+PARAMS=sym
+./experiment.py -t 20 --params $PARAMS --setup control
+./experiment.py -t 20 --params $PARAMS --setup nat
+./experiment.py -t 20 --params $PARAMS --setup vpn
+PARAMS=sym_lossy
+./experiment.py -t 20 --params $PARAMS --setup control
+./experiment.py -t 20 --params $PARAMS --setup nat
+./experiment.py -t 20 --params $PARAMS --setup vpn
+PARAMS=sym_delayed
+./experiment.py -t 20 --params $PARAMS --setup control
+./experiment.py -t 20 --params $PARAMS --setup nat
+./experiment.py -t 20 --params $PARAMS --setup vpn
+PARAMS=easy
+./experiment.py -t 20 --params $PARAMS --setup control
+./experiment.py -t 20 --params $PARAMS --setup nat
+./experiment.py -t 20 --params $PARAMS --setup vpn
+PARAMS=lossy
+./experiment.py -t 20 --params $PARAMS --setup control
+./experiment.py -t 20 --params $PARAMS --setup nat
+./experiment.py -t 20 --params $PARAMS --setup vpn
+PARAMS=delayed
+./experiment.py -t 20 --params $PARAMS --setup control
+./experiment.py -t 20 --params $PARAMS --setup nat
+./experiment.py -t 20 --params $PARAMS --setup vpn

### 2017-05-18 src/mn_experiment.sh
MOAR TRIALS

@@ -1,26 +1,27 @@
 #!/bin/sh
 
+TRIALS=60
 PARAMS=sym
-./experiment.py -t 20 --params $PARAMS --setup control
-./experiment.py -t 20 --params $PARAMS --setup nat
-./experiment.py -t 20 --params $PARAMS --setup vpn
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
 PARAMS=sym_lossy
-./experiment.py -t 20 --params $PARAMS --setup control
-./experiment.py -t 20 --params $PARAMS --setup nat
-./experiment.py -t 20 --params $PARAMS --setup vpn
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
 PARAMS=sym_delayed
-./experiment.py -t 20 --params $PARAMS --setup control
-./experiment.py -t 20 --params $PARAMS --setup nat
-./experiment.py -t 20 --params $PARAMS --setup vpn
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
 PARAMS=easy
-./experiment.py -t 20 --params $PARAMS --setup control
-./experiment.py -t 20 --params $PARAMS --setup nat
-./experiment.py -t 20 --params $PARAMS --setup vpn
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
 PARAMS=lossy
-./experiment.py -t 20 --params $PARAMS --setup control
-./experiment.py -t 20 --params $PARAMS --setup nat
-./experiment.py -t 20 --params $PARAMS --setup vpn
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
 PARAMS=delayed
-./experiment.py -t 20 --params $PARAMS --setup control
[…]

### 2017-05-18 data/aws2/plots.py
AWS plots yo... I think we need more trials

@@ -1,0 +1,54 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved
+SCEN = {
+    'server.ctrl': '1 Subflow',
+    'server.nat': 'NAT',
+    'server.vpn': 'VPN',
+}
+
+
+def make_throughput_plot():
+    data = []
+    labels = []
+    for s, label in SCEN.items():
+        data.append(o.get_numpy_array('%s.json' % s) / 1000000)
+        labels.append(label)
+
+    fig, ax = plt.subplots()
+    print(list(map(len, data)))
+    ax.boxplot(data, 0, '', labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: AWS')
+    fig.savefig('aws.pdf')
+
+
+def make_time_series_grid():
+    fig, axarr = plt.subplots(1, len(SCEN), sharex=True, sharey=True)
+    for cidx, col in enumerate(SCEN.keys()):
+        data = o.get_time_series_arrays('%s.json' % col)
+        for series in data:
+            axarr[cidx].plot(series[:10])
+    axarr[0].set_ylabel('Throughput (Mbps')
+    axarr[0].set_xlabel('1 Subflow')
+    axarr[1].set_xlabel('NAT')
+    axarr[2].set_xlabel('VPN')
+    fig.savefig('timegrid.aws.pdf')
+
+
+if __name__ == '__main__':
+    if len(sys.argv) > 1 and sys.argv[1] == 'time':
+        make_time_series_grid()
+    else:
+        make_throughput_plot()

### 2017-05-18 src/aws_experiment.py
trials 30 -> 100!!

@@ -29,6 +29,8 @@ import boto3
 import paramiko
 
 
+TRIALS = 100
+
 NAT_CFG = '''
 client: {
   detours = ["%s"];
@@ -254,7 +256,7 @@ def run_exp_nat(instances, clients):
 
     # run iperf several times, recv()ing server output
     server_output = BytesIO()
-    many_iperf(30, client_transport, server_chan, server_output, instances)
+    many_iperf(TRIALS, client_transport, server_chan, server_output, instances)
 
     # save control server output
     with open('server.ctrl.json', 'wb') as f:
@@ -281,7 +283,7 @@ def run_exp_nat(instances, clients):
     )
 
     server_output = BytesIO()
-    many_iperf(30, client_transport, server_chan, server_output, instances)
+    many_iperf(TRIALS, client_transport, server_chan, server_output, instances)
 
     # save nat server output
     with open('server.nat.json', 'wb') as f:
@@ -306,7 +308,7 @@ def run_exp_nat(instances, clients):
     )
 
     server_output = BytesIO()
-    many_iperf(30, client_transport, server_chan, server_output, instances)
+    many_iperf(TRIALS, client_transport, server_chan, server_output, instances)
 
     # save vpn output
     with open('server.vpn.json', 'wb') as f:

### 2017-05-18 src/aws_experiment.py
aws_experiment: i don't wanna go home yet

@@ -18,6 +18,7 @@ Prerequisites:
 
 """
 import collections
+import os
 import sys
 from io import BytesIO
 from os.path import abspath
@@ -52,8 +53,8 @@ client: {
 '''
 
 # Key Pair Name. Hopefully you named it the same thing in every region.
-KEYPAIR = 'stephen@greed'
-KEYFILE = '/home/stephen/.ssh/id_rsa'
+KEYPAIR = os.getenv('KEYPAIR') or 'stephen@greed'
+KEYFILE = os.getenv('KEYFILE') or '/home/stephen/.ssh/id_rsa'
 
 # The filename of the AWS setup script.
 SETUP_SCRIPT = join(dirname(abspath(__file__)), '../etc/aws_exp_setup.sh')

### 2017-05-18 data/aws3/plots.py
AWS please don't charge me lots of money for your bandwidth!

@@ -1,0 +1,54 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved
+SCEN = {
+    'server.ctrl': '1 Subflow',
+    'server.nat': 'NAT',
+    'server.vpn': 'VPN',
+}
+
+
+def make_throughput_plot():
+    data = []
+    labels = []
+    for s, label in SCEN.items():
+        data.append(o.get_numpy_array('%s.json' % s) / 1000000)
+        labels.append(label)
+
+    fig, ax = plt.subplots()
+    print(list(map(len, data)))
+    ax.boxplot(data, 0, '', labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: AWS')
+    fig.savefig('aws.pdf')
+
+
+def make_time_series_grid():
+    fig, axarr = plt.subplots(1, len(SCEN), sharex=True, sharey=True)
+    for cidx, col in enumerate(SCEN.keys()):
+        data = o.get_time_series_arrays('%s.json' % col)
+        for series in data:
+            axarr[cidx].plot(series[:10])
+    axarr[0].set_ylabel('Throughput (Mbps)')
+    axarr[0].set_xlabel('1 Subflow')
+    axarr[1].set_xlabel('NAT')
+    axarr[2].set_xlabel('VPN')
+    fig.savefig('timegrid.aws.pdf')
+
+
+if __name__ == '__main__':
+    if len(sys.argv) > 1 and sys.argv[1] == 'time':
+        make_time_series_grid()
+    else:
+        make_throughput_plot()

### 2017-05-20 src/experiment.py
experiment: add CPU limiting for extra safety
this will not affect the results at all... JSON output indicates CPU utilization
is in the single digit percents for everything

@@ -55,12 +55,13 @@ class DetourTopo(Topo):
         Topo.__init__(self)
 
         # Add hosts and switches
-        client = self.addHost('client')
-        r1 = self.addHost('r1')
-        r2 = self.addHost('r2')
-        server = self.addHost('server')
-        r3 = self.addHost('r3')
-        detour = self.addHost('detour')
+        cpu = 0.1
+        client = self.addHost('client', cpu=cpu)
+        r1 = self.addHost('r1', cpu=cpu)
+        r2 = self.addHost('r2', cpu=cpu)
+        server = self.addHost('server', cpu=cpu)
+        r3 = self.addHost('r3', cpu=cpu)
+        detour = self.addHost('detour', cpu=cpu)
 
         # Add links
         self.addLink(client, r1, **params.get('client_r1', {}))

### 2017-05-20 data/output.py
data: code for computing max CPU utilization. never went above 5%!

@@ -2,6 +2,7 @@
 # -*- coding: utf-8 -*-
 """Parse JSON output of experiments."""
 
+import itertools
 import json
 
 import numpy as np
@@ -32,6 +33,23 @@ def objects_to_bps(obj_list):
             yield bps
 
 
+def objects_to_utilization(obj_list):
+    """Converts a sequence of JSON objects to their CPU utulization average."""
+    for obj in obj_list:
+        # remote_total tends to be a percent or two higher, report it to be
+        # conservative
+        u = obj.get('end', {}).get('cpu_utilization_percent', {}).get('remote_total')
+        if u:
+            yield u
+
+
+def show_max_utilization(filenames):
+    """Report the max CPU utilization of all the JSON files."""
+    all_utilization = map(objects_to_utilization, map(json_parse, filenames))
+    chained = itertools.chain.from_iterable(all_utilization)
+    print(max(chained))
+
+
 def get_numpy_array(filename):
     return np.array(list(objects_to_bps(json_parse(filename))))
 
@@ -54,8 +72,16 @@ def plot_comparison(name):
     fig, ax = plt.subplots()
     data = [mptcp_ctrl, mptcp_nat, mptcp_vpn, vanilla_ctrl, vanilla_nat,
             vanilla_vpn]
-    labels = ['1 Subflow', 'NAT', 'VPN', 'TCP', 'TCP(NAT)', 'TCP(VPN)']
+    labels = ['1 Sub flow', 'NAT', 'VPN', 'TCP', 'TCP(NAT)', 'TCP(VPN)']
     print(list(map(len, data)))
     ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
     ax.set_title('Throughput Comparison (%s)' % name)
+
+
+if __name__ == '__main__':
+    import sys
+    if sys.argv[1] == 'cpu':
+        print('Computing max CPU utilization.')
+        filenames = sys.argv[2:]
+        show_max_utilization(filenames)

### 2017-05-22 data/2017-05-21-cpu-limit/plots.py
swap those figures like they're hot

@@ -1,0 +1,83 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved (thanks python 3.6...?)
+META = {
+    'easy': 'Core-limited',
+    'lossy': 'Core-limited with Loss',
+    'delayed': 'Core-limited with High Latency',
+    'sym': 'Symmetric',
+    'sym_lossy': 'Symmetric with Loss',
+    'sym_delayed': 'Symmetric with High Latency',
+}
+
+# order preserved
+SCEN = {
+    'control.mptcp': '1 Subflow',
+    'nat.mptcp': 'NAT',
+    'vpn.mptcp': 'VPN',
+    'control.vanilla': 'TCP',
+    'nat.vanilla': 'TCP (NAT)',
+    'vpn.vanilla': 'TCP (VPN)',
+}
+
+
+def make_single_plot(base):
+    print(base)
+    data = []
+    labels = []
+    for s, label in SCEN.items():
+        data.append(o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000)
+        labels.append(label)
+
+    fig, ax = plt.subplots()
+    print(list(map(len, data)))
+    ax.boxplot(data, 0, '', labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: ' + META[base])
+    fig.savefig('%s.pdf' % base)
+
+
+def make_all_plots():
+    for base in META.keys():
+        make_single_plot(base)
+
+
+def make_time_series_grid(base):
+    print(base)
+    rows = ['mptcp', 'vanilla']
+    cols = ['control', 'nat', 'vpn']
+    fig, axarr = plt.subplots(len(rows), len(cols), sharex=True, sharey=True)
+    for ridx, row in enumerate(rows):
+        for cidx, col in enumerate(cols):
+            data = o.get_time_series_arrays('%s.%s.%s.json' % (base, col, row))
+            for series in data:
+                axarr[ridx][cidx].plot(series[:10])
+    axarr[0][0].set_ylabel('MPTCP')
+    axarr[1][0].set_ylabel('TCP')
+    axarr[1][0].set_xlabel('Control')
[…]

### 2017-05-23 paper/presentation.tex
presentation: initial commit

\documentclass{beamer}
\usetheme{CWRU}

\title{Exploring Alternative Routes Using Multipath TCP}
\author{Stephen Brennan}
\date{June 5, 2017}

\begin{document}
\frame{\titlepage}
\end{document}

### 2017-05-26 paper/presentation.tex
A start on presentation introduction
presentation: I swear this shouldn't be this hard

@@
 
+\usepackage{appendixnumberbeamer}
+\usepackage{pgfpages}
+\usepackage{wrapfig}
+\usepackage{graphicx}
+\setbeamertemplate{note page}[plain]
+\setbeameroption{show notes on second screen=right}
+
 \begin{document}
 \frame{\titlepage}
+
+\section{Introduction \& Background}
+
+\begin{frame}
+  \frametitle{Internet Routing Inefficiencies}
+  \begin{columns}
+  \begin{column}{0.7\textwidth}
+  \begin{itemize}
+  \item The default route is not always the best, in terms of latency or
+    reliability
+  \item Peering agreements and policy based routing can result in suboptimal
+    routing decisions \cite{detour}
+  \end{itemize}
+  \end{column}
+  \begin{column}{0.3\textwidth}
+    \includegraphics[width=\textwidth]{figures/detour.png} \\
+    \textit{Figure from \cite{detour}}
+  \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Internet Routing Inefficiencies}
+
+  \includegraphics[height=0.85\textheight]{figures/detour-packetloss.png} \\
+  \textit{Figure from \cite{detour}}
+  \note{hello, world}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Access Link Underutilization}
+
+  \begin{itemize}
+  \item Fiber-to-the-home is coming!
+  \item However, residential bandwidth is not fully utilized
+    \cite{fibertothehome}
+    \begin{itemize}
+    \item Short-lived TCP sessions?
+    \item Small TCP initial window?
+    \item Network core can't support bandwidth?
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Contributions}
+
+  \begin{itemize}
+  \item
+  \end{itemize}
+\end{frame}
+
+\section{Related Work}
+\section{Implementation}
+\section{Evaluation}
+\section{Conclusion}
+
+\appendix % slide numbering includes nothing after here
+
+\section{References}
+\begin{frame}[allowframebreaks]
+  \frametitle{References}
+  \bibliographystyle{ieeetr}
+  \bibliography{paper.bib}
+\end{frame}
+
 \end{document}

### 2017-05-29 paper/presentation.tex
presentation: intro is done?
presentation: Add real notes, better intro

@@
 \author{Stephen Brennan}
+\institute{Case Western Reserve University}
 \date{June 5, 2017}
@@
 \usepackage{wrapfig}
+\usepackage[export]{adjustbox}
 \usepackage{graphicx}
@@
 \setbeameroption{show notes on second screen=right}
+\setbeamertemplate{navigation symbols}{}%remove navigation symbols
+
+\newcommand{\FigureSlide}[2]{
+  \begin{frame}[plain]
+    \includegraphics[max width=\textwidth,max height=\textheight]{#1}
+    \note{#2}
+  \end{frame}
+}
 
 \begin{document}
-\frame{\titlepage}
+\frame{
+  \titlepage
+  \note{
+    Hello everyone. I'm Stephen, and today I'll be discussing my thesis,
+    entitled ``Exploring Alternative Routes Using Multipath TCP''.
+  }
+}
 
-\section{Introduction \& Background}
+\begin{frame}{Overview}
+  \tableofcontents
+  \note{
+    Here is how this talk is going to be structured. First, I'll be giving a bit
+    of motivation for this project, along with a problem statement and our
+    background information. Next, I'll discuss related work in this field, and
+    show how this thesis fits into the current research. After that, I'll give
+    an overview of the implementation of the components of the thesis. Next,
+    I'll discuss the experiments I used to validate the mechanism, and finally
+    we'll conclude with a brief summary and discussion of future work.
+  }
+\end{frame}
+
+\section{Introduction}
+
+\FigureSlide{figures/InternetSlide.pdf}{
+  Okay, so let's get started! I want to start off with a brief and simplified
+  look at the Internet. As we know, the Internet is made up of routers, the blue
+  dots, and Autonomous Systems, the tan groups. The task of each router is to
+  forward small packets of information toward their final destination. The route
+  that packets take is determined by a lot of factors.
+
+  While this project isn't directly about routing, let's discuss some of these
+  factors. Routing is typically performed, in some way or another, by an
+  application of Dijkstra's algorithm for finding a shortest path. While this
+  might be optimal in a theoretical sense, the practical application is not, and
+  for several reasons.
+
+  For one, routing algorithms are going to work by ``hop count'',
+  meaning that a shorter path should be preferred, even if the longer path might
+  actually exhibit better properties. For another, the Autonomous Systems are
+  frequently businesses or ISPs, and the connections they make are determined by
+  business agreements. They may create policies that are less than optimal in
+  order to fulfill business objectives.
+}
 
@@
     routing decisions \cite{detour}
+  \item A route that passes through a ``detour'' may be better
   \end{itemize}
@@
   \end{columns}
+  \note{
+    Research in the field has been able to fairly concretely demonstrate the
+    problem. Just because your traffic takes a particular route, doesn't mean it
+    is going to be the best path for you, in terms of latency or reliability.
+
+    Instead, if you were to route your packets to a ``detour'' and then to the
+    final destination, you might actually get path which has better round trip
+    time or loss rate. On the right we have an example of a situation found by a
+    study, where the default route from Los Altos to Walnut Creek had a round
+    trip time of 137 ms, while the combined round trip time of a path via Santa
+    Clara was only 30 ms. The reason was that the default route actually
+    traveled through Chicago!
+  }
 \end{frame}
 
-\begin{frame}
-  \frametitle{Internet Routing Inefficiencies}
+\FigureSlide{figures/detour-packetloss.png}{
+  This effect is not limited to isolated examples, either. This same study
+  looked at a large group of computers on university networks. They computed
+  latency and packet loss rates between each pair of computers. Then, they
+  searched for pairs in which the ``triangle inequality'' was violated.
 
-  \includegraphics[height=0.85\textheight]{figures/detour-packetloss.png} \\
-  \textit{Figure from \cite{detour}}
-  \note{hello, world}
-\end{frame}
+  This graph shows packet loss rates between pairs of computers. The blue is the
+  default Internet routed path, and the yellow is the best alternative path that
+  uses a single detour. Most of the yellow dots are actually on the zero line.
+
+  You can see here that for a large majority of pairs, there is an alternative
+  route that is more reliable. For a small fraction here, the difference in
+  reliability is actually huge.
+}
 
@@
     \end{itemize}
+  \item Using alternative routes can improve performance
+  \item Aggregating multiple routes can perform even better
   \end{itemize}
+
+  \note{
+    Now let's take a brief detour and discuss something else. Recently, we've
+    been seeing some exciting trends in residential bandwidth. New access links
+    are always being created, and the latest improvement is fiber to the home.
+    These can provide up to Gigabit speeds to a home network. However, we've
+    had studies right here at CWRU showing that these residential links don't
+    use most of that capacity.
+
+    There are several good explanations for this. Short lived TCP sessions may
+    never leave the slow-start stage. A small TCP initial window will limit the
+    throughput that can be achieved in a reasonable amount of time. But there's
+    also the possibility that the network core is becoming the bottleneck for
+    some of these connections.
+
+    Two things seem to follow. First, there are good chances that a detour route
+    could provide better throughput for a TCP connection, as studies have
+    already demonstrated. Second, if we have extra access link capacity, why not
+    use these paths in combination with each other, rather than choosing only
+    one?
+  }
 \end{frame}
@@
 
+  \textbf{Problem:} Internet routing inefficiencies can prevent TCP connections
+  from achieving their full potential throughput.
+
+  \vspace{0.5cm}
+
+  \textbf{Solution:} A system which can use and aggregate alternative routes for
+  TCP connections.
+
+  \vspace{0.5cm}
+
+  \textbf{Contributions:}
+
   \begin{itemize}
-  \item
+  \item A mechanism for adding alternative routes to Multipath TCP (MPTCP)
+    sessions among single-homed hosts.
+  \item An evaluation of this mechanism on emulated and real-world networks.
   \end{itemize}
+  \note{
+    Putting this together, we have a problem: routing inefficiencies can prevent
+    TCP connections from achieving their full potential throughput. Access links
+    may be capable of much better bandwidth than a user actually gets.
+
+    Our solution is to create a system which can use and aggregate alternative
+    paths for TCP connections. Using these paths means that it can route traffic
+    along them. Aggregation means combining the potential throughput of both
+    paths, to get better throughput than either.
+
+    The contributions this thesis makes are twofold: first, a mechanism that
+    allows single-homed devices to add subflows along alternative routes via
+    MPTCP. Second, an evaluation of this system on emulated and real world
+    networks.
+  }
+\end{frame}
+
+\section{Background}
+
+\begin{frame}
+  \frametitle{Multipath TCP}
+
+  \begin{itemize}
+  \item Multi-homed devices are becoming more common
+    \begin{itemize}
+    \item Smartphones
+    \item Datacenters
+    \item Laptops
+    \end{itemize}
+  \item TCP still views a connection as a five-tuple: (TCP, Source IP, Source
+    port, Destination IP, Destination Port)
+  \item Multi-homed devices are forced to choose a network interface
+  \item Multipath TCP is an extension to TCP, allowing hosts to use multiple
+    addresses in the same connection
+  \end{itemize}
+    \note{
+      Recently, we've seen a rise in devices that have multiple ways to connect
+      to the Internet, network interfaces. For instance, smartphones include
+      cellular and WiFi radios, and can be connected to the Internet on both.
+      Recent data centers have been taking advantage of multiple interfaces and
+      highly redundant network topologies. Laptops can also contain WiFi,
+      Ethernet, and even cellular radios.
+
+      Unfortunately, TCP was designed for a simpler time, when computers had a
+      single interface and their addresses never changed. This forces
+      multi-homed devices to decide which interface they'll use and stick to
+      that choice.
+
+      Multipath TCP aims to solve this problem. It is an extension to TCP (not a
+      replacement) that allows devices to make use of all of their network
+      interfaces and addresses in a single connection.
+    }
 \end{frame}

### 2017-05-30 paper/presentation.tex
Done with background

@@
 \usepackage{graphicx}
+\usepackage{fancyvrb}
 \setbeamertemplate{note page}[plain]
@@
   \begin{frame}[plain]
+    \begin{centering}
+      \vfill
       \includegraphics[max width=\textwidth,max height=\textheight]{#1}
+      \vfill
+      \end{centering}
     \note{#2}
+  \end{frame}
+}
+
+\newcommand{\SectionHead}[2]{
+  \begin{frame}{#1}
+    \tableofcontents[currentsection,hideallsubsections]
+    \note{#1}
   \end{frame}
@@
 
+\section{Introduction}
+
 \begin{frame}{Overview}
-  \tableofcontents
+  \tableofcontents[hideallsubsections]
   \note{
@@
 \end{frame}
-
-\section{Introduction}
 
@@
 \section{Background}
+\SectionHead{}{}
+\subsection{Multipath TCP}
 
@@
     \note{
+
+      This project is based on MPTCP. While it's not my intent to give a
+      comprehensive protocol description in this presentation, I do want to give
+      enough background to understand how my project uses it.
+
       Recently, we've seen a rise in devices that have multiple ways to connect
@@
 
+\begin{frame}{Design Goals}
+
+  \begin{itemize}
+  \item Improve performance and reliability over current TCP, by aggregating
+    paths created by multiple interfaces.
+  \item Do no harm to single-path TCP, by taking no more bandwidth over shared
+    bottlenecks than standard TCP would
+  \item Remain compatible with TCP applications and the Internet
+    \begin{itemize}
+    \item Present the same socket API to applications
+    \item Remain similar to TCP on the wire, to remain compatible with Internet
+      middleboxes
+    \end{itemize}
+  \end{itemize}
+  \note{
+    As a protocol, MPTCP has a few key design goals. First, it is designed to
+    improve performance and reliability compared to standard TCP. Second, it
+    should not do any harm to existing single-path TCP. Third, it should remain
+    compatible to applications and to the Internet as a whole.
+  }
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{Architecture}
+  \begin{center}
+\begin{BVerbatim}
+                     +-------------------------------+
+                     |           Application         |
++---------------+    +-------------------------------+
+|  Application  |    |             MPTCP             |
++---------------+    + - - - - - - - + - - - - - - - +
+|      TCP      |    | Subflow (TCP) | Subflow (TCP) |
++---------------+    +-------------------------------+
+|      IP       |    |       IP      |      IP       |
++---------------+    +-------------------------------+
+\end{BVerbatim}
+  \end{center}
+  \note{
+    The way that MPTCP achieves this is by layering the protocol extension in
+    between TCP and the application layer. An MPTCP connection consists of one
+    or more (well, potentially zero or more) TCP connections, which we refer to
+    as ``subflows''. These subflows bear some extra TCP options which provide
+    signaling information, but other than that they behave very similarly to
+    standard TCP. The only major difference is that their receive window is
+    shared, rather than per-subflow. The MPTCP connection has its own sequence
+    numbers, and each subflow has independent sequence numbers as well.
+  }
+\end{frame}
+
+\begin{frame}
+  \frametitle{Path Management}
+
+  \begin{itemize}
+  \item Subflows are established with a three way handshake
+  \item First subflow uses \texttt{MP\_CAPABLE} option
+  \item Subsequent subflows use \texttt{MP\_JOIN} option
+  \item Additional addresses may be advertised using \texttt{ADD\_ADDR} at any
+    time
+  \item Either side may create new subflows at any time
+  \end{itemize}
+  \note{
+
+    When we create an MPTCP connection, a three-way handshake occurs, just like
+    standard TCP. However, the three-way handshake includes the MP CAPABLE
+    option, signalling to the other host that it supports MPTCP. A SYNACK with
+    MP CAPABLE signals that the passive opener supports MPTCP as well, and the
+    final ACK with MP CAPABLE confirms the use of MPTCP. The contents of these
+    options are key material, which is used to authenticate creation of new
+    subflows.
+
+    Subsequent subflows are created with a similar procedure, but using the MP
+    JOIN option. The contents of the option this time are HMAC signatures based
+    on the key material. If the HMAC can be verified, the hosts can be certain
+    that they are communicating with the same host they initially established a
+    connection with.
+
+    An MPTCP implementation may advertise that it has additional addresses (and
+    therefore additional interfaces) via an ADD ADDR option. There is also
+    support for deletion.
+
+    Finally, the MPTCP specification says nothing about when subflows should be
+    created or destroyed. This is entirely left up to local policy, and of
+    course research. In the Linux kernel implementation of MPTCP, a module
+    interface exists for this, called the path manager.
+
+  }
+\end{frame}
+
+\FigureSlide{figures/DataScheduling.pdf}{
+
+  So, now that we have subflows, what do we do with them? How do we get data
+  through them to the other application? Well, we do this with two important
+  concepts: a data scheduler and the data sequence mapping.
+
+  Imagine one host wants to send some data. It passes it through the send system
+  call, and eventually the MPTCP stack has a large piece of data. A component
+  called the scheduler gets to make the executive decision of how to divide up
+  the data. It does this however it pleases, and the chunks make their way down
+  to the subflows.
+
+  In order to be able to put them back together, a data sequence signal is sent
+  on the subflow.
+
+}
+
 \section{Related Work}

### 2017-05-30 paper/presentation.tex
presentation: related work

@@
 \setbeamertemplate{navigation symbols}{}%remove navigation symbols
+\AtBeginNote{%
+    \let\enumerate\itemize%
+    \let\endenumerate\enditemize%
+}
 
@@
     \tableofcontents[currentsection,hideallsubsections]
-    \note{#1}
+    \note{#2}
   \end{frame}
@@
 \section{Related Work}
+\SectionHead{}{
+  Would love to discuss more, but limited time
+}
+
+\subsection{Overlay Networks}
+
+\begin{frame}
+  \frametitle{Resilient Overlay Networks}
+
+  \begin{itemize}
+  \item Rather than use only one detour, create an overlay network
+  \item Overlay nodes use the Internet as their ``link layer''
+    \note[item]{In Intro, discussed Detour study. RON is something of a
+      spiritual successor. Rather than simple detours, use overlay networks.}
+  \item Routing performed at each node using measured link characteristics
+    \note[item]{Each node is an Internet host. Constantly monitoring link
+      characteristics of network, updating routing tables with performance
+      metrics.}
+  \item Several studies based on RON:
+    \note[item]{RON achieved good results, was testbed for additional studies.}
+    \begin{itemize}
+    \item Redundant multipath routing
+      \note[item]{One study compared two routing strategies: adaptively choosing
+      the best path, or redundantly using multiple paths. No aggregation.}
+  \item ``Biologically inspired'' multipath routing
+    \note[item]{Another - biologically inspired. Choose strongest path as
+      primary, have several backups, involves differential equations and
+      randomness.}
+  \item mTCP \note[item]{mTCP: most important. A modification of TCP before
+      MPTCP existed. Built on RON + userspace TCP stack. Did not have MPTCP
+      compatibility concerns - routed segments across different paths. Worked on
+      single-homed devices due to RON}
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Other Related Items}
+
+  \note[item]{Outside of overlay networks, several related pieces of work, both
+    in research and in the standards bodies.}
+
+  \begin{itemize}
+  \item Source routing (IPv4) and segment routing (IPv6)
+    \note[item]{Notable - it is possible to specify intermediaries in both IPv4
+      and IPv6. For security reasons these aren't respected. Standards exist
+      mainly for use in internal networks.}
+  \item Binder: aggregating bandwidth of several Internet gateways
+    \note[item]{One use of MPTCP that is very similar was called Binder. A rural
+      town was away from any ISP. Had several potential gateways far away, but
+      low bandwidth and not good reliability. Established connections across
+      each gateway, and created an OpenVPN connection to an Internet-side
+      server, with subflow across each.}
+  \item SOSR: Scalable One-hop Source Routing \note[item]{Work inspired by RON,
+      but trying to simplify. This work simply tunneled connections across a
+      single hop, using a NAT approach similar to ours. Found they got most of
+      RON's benefit without the full overlay system.}
+  \item IETF MPTCP proxying draft
+    \note[item]{Similar because allows single-homed devices to use MPTCP.
+      However this approach expects non MPTCP server, and is ``transparent''
+      proxy.}
+  \end{itemize}
+\end{frame}
+
 \section{Implementation}

### 2017-05-30 paper/presentation.tex
presentation: Implementation

@@
 \section{Implementation}
+\SectionHead{}{
+  Time to get into the actual implementation of this project.
+}
+\subsection{Overview}
+
+\FigureSlide{figures/Concept.pdf}{
+  \begin{itemize}
+  \item Some terminology.
+  \item Active opener - ``client''
+  \item Passive opener - ``server''
+  \item Intermediary - ``detour''
+  \item This mechanism allows subflows to be created across (potentially
+    multiple) detours
+  \item (in addition to the main subflow directly from client to server)
+  \end{itemize}
+}
+
+\begin{frame}
+  \frametitle{Ingredients}
+
+  \note{
+    There are several components that go together to make this thesis work.
+  }
+
+  \begin{itemize}
+  \item Multipath TCP Linux Implementation v0.91
+  \item Custom path manager
+  \item OpenVPN
+  \item Netfilter / IPTables frameworks
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/MovingParts.pdf}{
+  \begin{itemize}
+  \item There are three components that make this work.
+  \item Assume server runs any MPTCP stack, unmodified
+  \item First component: Client runs Linux MPTCP with custom path manager
+  \item Second component: Client runs userspace daemon which communicates with
+    detour.
+  \item Third component: detour runs userspace daemon. Currently detour is
+    restricted to any relatively modern linux (IPTables, Netfilter). Need not
+    support MPTCP.
+  \end{itemize}
+}
+
+\subsection{Detour Daemon}
+
+\begin{frame}
+  \frametitle{Strategies for Detours}
+
+  \begin{itemize}
+  \item OpenVPN Approach
+    \begin{itemize}
+    \item Establish an OpenVPN connection with detour
+    \item Send packets as normal through the virtual interface
+    \item Packets encapsulated via OpenVPN protocol
+    \item Detour still performs NAT
+    \end{itemize}
+  \item NAT Approach
+    \begin{itemize}
+    \item Address packets directly to detour
+    \item Detour alters source and destination address, forwards packet
+    \item Address information must be arranged ahead of time
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{OpenVPN Approach}
+
+  \begin{itemize}
+  \item OpenVPN typically provides encryption and authentication
+    \note[item]{OpenVPN is well known for providing an encrypted VPN service.
+      While this is normally a good thing, in our use case, it is just
+      overhead.}
+  \item Configure to only provide authentication on startup, no encryption or
+    message signatures
+    \note[item]{To address this, we configure OpenVPN so that it will still
+      require X.509 keys in order to negotiate the connection. But no encryption
+      signatures will occur for data.}
+  \item Use UDP as transport, to avoid ``TCP Meltdown''
+    \note[item]{When TCP is encapsulated within itself, we can encounter the TCP
+      Meltdown effect. This is due to multiple timers in the different layers.
+      To address this, we use UDP as the transport layer.}
+  \item VPN appears as network device to the kernel
+  \item No per-MPTCP-connection signalling, but has per-packet overhead
+    \note[item]{Some advantages here again, the VPN is just a device to the
+      kernel, so this is the case MPTCP is designed to handle. There is no
+      signalling, so connections can be created immediately, but there is a
+      per-packet overhead.}
+  \end{itemize}
+
+  \note[item]{
+    \textbf{If asked about TCP Meltdown:} Packet loss causes retransmission on
+    outer TCP connection. Its retransmission timer is likely longer than the
+    inner TCP. The inner timer fires a retransmission, and exponential backoff
+    begins. Since the outer TCP is blocked until the retransmission, the inner
+    adds up more retransmissions and exponentially backs off much more, and so
+    the connection stalls quickly.
+  }
+\end{frame}
+
+\begin{frame}[fragile]
+  \frametitle{NAT Approach}
+
+  \begin{center}
+    \vfill
+\begin{BVerbatim}
++-----------------------------------+
+| ver(1) | op(1)  | reserved (2)    |
++-----------------------------------+
+|           rip (4 bytes)           |
++-----------------------------------+
+| rpt (2 bytes)   | dpt (2 bytes)   |
++-----------------|-----------------+
+\end{BVerbatim}
+    \vfill
+
+    Custom protocol for arranging NAT detours
+  \end{center}
+  \note[item]{The NAT approach is perhaps the most straightforward.}
+  \note[item]{The biggest issue is that a client must inform the detour of the
+    server address and port. They must also agree what port on the detour to
+    use. This protocol does this.}
+  \note[item]{The ver field is a version, currently set to 1.}
+  \note[item]{The op field is the operation, which can be either ``request'' or
+    ``response''.}
+  \note[item]{rip is the remote address to connect to}
+  \note[item]{rpt is remote port}
+  \note[item]{dpt is detour port}
+  \note[item]{In the request, dpt is simply a ``proposal'', typically the same
+    port as rpt. The detour replies back with the final choice of dpt.}
+  \note[item]{Detour may want to choose a different one due to several types of
+    port conflicts which could occur in detours, or because the proposed port
+    (e.g. 80) is open and listening, and the detour doesn't want to mask it.}
+\end{frame}
+
+\subsection{Path Manager}
+
+\begin{frame}
+  \frametitle{Path Manager}
+
+  \begin{itemize}
+  \item Once a MPTCP connection is established, path manager is informed
+  \item Path manager runs in a background thread
+  \item Requests detours from client daemon
+  \item Adds up to $N$ additional subflows, where $N$ is configurable. By
+    default $N=2$
+  \item Whenever a new detour becomes available, runs again
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/KernDataStruct.pdf}{
+  \begin{itemize}
+  \item The path manager holds onto some internal data
+  \item Tracks each NAT and VPN detour in lists, with timestamps
+  \item Timestamps ensure that when the worker runs again, entries in the list
+    are not reconsidered.
+  \item The internal data is not global, it exists per network-namespace.
+  \item Connections belong to network namespace, so do network interfaces.
+  \end{itemize}
+}
+
+\subsection{Client Daemon}
+
+\begin{frame}
+  \frametitle{Client Daemon}
+
+  \begin{itemize}
+  \item Userspace daemon required for tasks which are not well-suited for the
+    kernel:
+    \begin{itemize}
+    \item Starting processes
+    \item Using UDP sockets
+    \end{itemize}
+  \item Daemon reads configuration file containing NAT and VPN detours.
+  \item VPN instances are started up first and reported to kernel
+  \item Wait for detour requests from kernel, send UDP requests, report replies
+    to kernel
+  \item All communication over Generic Netlink
+  \end{itemize}
+\end{frame}
+
+\subsection{Putting it Together}
+
+\begin{frame}[t]
+  \frametitle{Putting it Together (NAT)}
+  \begin{enumerate}
+  \invisible<1>{\item[(0)] At startup, client daemon connects to VPN and reports VPN to kernel}
+  \item Application creates MPTCP connection to MPTCP supporting server
+  \item Once 3WHS completes, path manager requests a detour from client daemon
+  \item Client daemon receives request and sends UDP request to every client
+    daemon listed in configuration file
+  \item Detour daemon sets up detour, sends reply
+  \item Client daemon forwards reply to kernel
+  \item The path manager restarts the MPTCP connection's thread, which creates
+    a new subflow via this detour
+  \end{enumerate}
+  \note[item]{When you do it with a VPN, the main difference is that the VPN is
+    already created. So the path manager can immediately create a subflow
+    through the VPN, rather than waiting for a detour entry}
+\end{frame}
+
+\begin{frame}[t]
+  \frametitle{Putting it Together (VPN)}
+
+  \begin{enumerate}
+  \item[(0)] At startup, client daemon connects to VPN and reports VPN to kernel
+  \item Application creates MPTCP connection to MPTCP supporting server
+  \item Once 3WHS completes, path manager requests a detour from client daemon.
+  \item Meanwhile, it uses the VPN already available and establishes a subflow.
+  \end{enumerate}
+\end{frame}
+
 \section{Evaluation}

### 2017-05-30 paper/presentation.tex
Rename figures because LaTeX is stupid

@@
     \note{#2}
+  \end{frame}
+}
+
+\newcommand{\TitledFigureSlide}[2]{
+  \begin{frame}{#1}
+    \begin{centering}
+      \vfill
+      \includegraphics[max width=\textwidth,max height=\textheight]{#1}
+      \vfill
+      \end{centering}
   \end{frame}
@@
 \section{Evaluation}
+\SectionHead{}{}
+
+\begin{frame}
+  \frametitle{Types of Experiment}
+
+  \begin{itemize}
+  \item Previous work has established that there do exist common scenarios where
+    detour routing can improve path characteristics
+  \item These experiments do not duplicate results
+  \item Answer the following:
+    \begin{itemize}
+    \item Given a network where an alternative route has better properties, can
+      this mechanism effectively use it?
+    \item Given a network where there is an alternative route and the potential
+      for aggregation, can this mechanism aggregate throughput?
+    \item Can this mechanism run without modification on the Internet?
+    \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\subsection{Mininet Experiments}
+
+\begin{frame}
+  \frametitle{Mininet Experiments}
+
+  \begin{itemize}
+  \item Mininet allows you to create arbitrary network topologies
+  \item Uses host networking stack rather than alternative or simulation
+  \item Uses namespacing (foundation of containerization) rather than
+    virtualization
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/Topology.pdf}{
+  \begin{itemize}
+  \item This is network topology we us
+  \item Has path diversity, a client, a server, a detour
+  \item Each link has 5ms delay.
+  \end{itemize}
+}
+
+\begin{frame}
+  \frametitle{Scenarios}
+
+  \begin{itemize}
+  \item Two types of network:
+    \begin{itemize}
+    \item \textbf{Symmetric:} every link has 10Mbps bandwidth
+    \item \textbf{Core-limited:} core links have 10Mbps, access links have 20Mbps
+    \end{itemize}
+  \item Three variations:
+    \begin{itemize}
+    \item \textbf{Normal:} no loss
+    \item \textbf{Lossy:} 1\% packet loss on Link 2
+    \item \textbf{Delayed:} 100ms delay on Link 2
+    \end{itemize}
+  \item Total of 6 scenarios
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Mechanisms}
+
+  \begin{itemize}
+  \item \textbf{1-Subflow:} MPTCP with no available detours
+  \item \textbf{NAT:} Using NAT detour
+  \item \textbf{VPN:} Using VPN detour
+  \item \textbf{TCP:} TCP over default route
+  \item \textbf{TCP(NAT):} TCP via the NAT tunnel
+  \item \textbf{TCP(VPN):} TCP via the VPN tunnel
+  \end{itemize}
+  \note[item]{Don't forget to say we use iperf to measure throughput}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Results}
+\end{frame}
+
+\FigureSlide{figures/sym.pdf}{
+  \begin{itemize}
+  \item This is
+  \end{itemize}
+}
+
 \section{Conclusion}
@@
 \appendix % slide numbering includes nothing after here
+
+\section{All Figures}
+\TitledFigureSlide{figures/sym.pdf}{Throughput Comparison, Symmetric}
+\TitledFigureSlide{figures/timegrid-sym.pdf}{Timelapse, Symmetric}
+
+\TitledFigureSlide{figures/sym-lossy.pdf}{Throughput Comparison, Symmetric with Loss}
+\TitledFigureSlide{figures/timegrid-sym-lossy.pdf}{Timelapse, Symmetric with Loss}
+
+\TitledFigureSlide{figures/sym-delayed.pdf}{Throughput Comparison, Symmetric with Delay}
+\TitledFigureSlide{figures/timegrid-sym-delayed.pdf}{Timelapse, Symmetric with Delay}
+
+\TitledFigureSlide{figures/easy.pdf}{Throughput Comparison, Core-limited}
+\TitledFigureSlide{figures/timegrid-easy.pdf}{Timelapse, Core-limited}
+
+\TitledFigureSlide{figures/easy-lossy.pdf}{Throughput Comparison, Core-limited with Loss}
+\TitledFigureSlide{figures/timegrid-easy-lossy.pdf}{Timelapse, Core-limited with Loss}
+
+\TitledFigureSlide{figures/easy-delayed.pdf}{Throughput Comparison, Core-limited with Delay}
+\TitledFigureSlide{figures/timegrid-easy-delayed.pdf}{Timelapse, Core-limited with Delay}
 

### 2017-05-30 paper/presentation.tex
presentation: evaluation

@@
 \newcommand{\TitledFigureSlide}[2]{
-  \begin{frame}{#1}
+  \begin{frame}{#2}
     \begin{centering}
@@
   \begin{itemize}
-  \item This is
+  \item This is a comparison of throughput on Symmetric, Normal. takeaways:
+  \item (1) MPTCP is more variable than TCP
+  \item (2) The VPN and NAT approaches don't perform as well when there is
+    nothing to gain. This seems to be a result of their overhead and maybe
+    congestion control.
   \end{itemize}
 }
+
+\begin{frame}
+  \frametitle{Results}
+
+  \begin{itemize}
+  \item MPTCP trials are more variable
+  \item Mechanism introduces overhead when there is no potential for gain
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/sym-lossy.pdf}{
+  \begin{itemize}
+  \item When a small loss rate is added, the story begins to change.
+  \item Suddenly, MPTCP NAT and VPN are performing very comparably to vanilla
+    TCP over those same paths
+  \item MPTCP 1 subflow takes a huge performance hit
+  \end{itemize}
+}
+
+\begin{frame}
+  \frametitle{Results}
+
+  \begin{itemize}
+  \item MPTCP trials are more variable
+  \item Mechanism introduces overhead when there is no potential for gain
+  \item Mechanism performs similarly to TCP over best path when the default path
+    has loss
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/easy.pdf}{
+  \begin{itemize}
+  \item When there is the potential to aggregate throughput, the story changes
+    again.
+  \item The mechanism is capable of aggregating nearly all bandwidth
+  \end{itemize}
+}
+
+\begin{frame}
+  \frametitle{Results}
+
+  \begin{itemize}
+  \item MPTCP trials are more variable
+  \item Mechanism introduces overhead when there is no potential for gain
+  \item Mechanism performs similarly to TCP over best path when the default path
+    has loss
+  \item Mechanism can effectively aggregate bandwidth when the network has the
+    potential
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/lossy.pdf}{
+  \begin{itemize}
+  \item With the lossy link, it still works
+  \end{itemize}
+}
+
+\begin{frame}
+  \frametitle{Results}
+
+  \begin{itemize}
+  \item MPTCP trials are more variable
+  \item Mechanism introduces overhead when there is no potential for gain
+  \item Mechanism performs similarly to TCP over best path when the default path
+    has loss
+  \item Mechanism can effectively aggregate bandwidth when the network has the
+    potential, even in the presence of loss
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/delayed.pdf}{
+  \begin{itemize}
+  \item With delayed link, there are some funky interactions with slow slow
+    start, but MPTCP still outperforms the best paths
+  \end{itemize}
+}
+
+\begin{frame}
+  \frametitle{Results}
+
+  \begin{itemize}
+  \item MPTCP trials are more variable
+  \item Mechanism introduces overhead when there is no potential for gain
+  \item Mechanism performs similarly to TCP over best path when the default path
+    has loss
+  \item Mechanism can effectively aggregate bandwidth when the network has the
+    potential, even in the presence of loss or high latency
+  \item<2> NAT consistently outperforms VPN both in MPTCP and TCP, but by a small
+    amount.
+  \end{itemize}
+\end{frame}
+
+\subsection{AWS Experiments}
+\begin{frame}
+  \frametitle{AWS Experiments}
+
+  \begin{itemize}
+  \item Deployed client, server, and detour implementations to different AWS
+    regions
+  \item Ran similar throughput measurements for MPTCP
+  \item Performed at much higer level, but didn't show similar improvements
+  \end{itemize}
+\end{frame}
+
+\FigureSlide{figures/aws.pdf}{}
 
@@
 
-\TitledFigureSlide{figures/easy-lossy.pdf}{Throughput Comparison, Core-limited with Loss}
-\TitledFigureSlide{figures/timegrid-easy-lossy.pdf}{Timelapse, Core-limited with Loss}
+\TitledFigureSlide{figures/lossy.pdf}{Throughput Comparison, Core-limited with Loss}
+\TitledFigureSlide{figures/timegrid-lossy.pdf}{Timelapse, Core-limited with Loss}
 
-\TitledFigureSlide{figures/easy-delayed.pdf}{Throughput Comparison, Core-limited with Delay}
-\TitledFigureSlide{figures/timegrid-easy-delayed.pdf}{Timelapse, Core-limited with Delay}
+\TitledFigureSlide{figures/delayed.pdf}{Throughput Comparison, Core-limited with Delay}
+\TitledFigureSlide{figures/timegrid-delayed.pdf}{Timelapse, Core-limited with Delay}
+
+\TitledFigureSlide{figures/aws.pdf}{Throughput Comparison, AWS}
+\TitledFigureSlide{figures/timegrid-aws.pdf}{Timelapse, AWS}
 

### 2017-05-30 paper/presentation.tex
presentation: Conclusion

@@
 \section{Conclusion}
+\SectionHead{}{}
+
+\begin{frame}
+  \frametitle{Future Work}
+
+  \begin{itemize}
+  \item Detour Collective
+  \item Dynamic subflow addition and removal
+  \item Data scheduling
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Summary}
+
+  \begin{itemize}
+  \item Created a system for adding detour routes to MPTCP connections between
+    single-homed devices.
+  \item Like MPTCP, this system works outside of experimental networks
+  \item System is capable of achieving similar performance to the best available
+    path when no aggregation is possible
+  \item System is capable of aggregating throughput when possible
+  \end{itemize}
+\end{frame}
 

### 2017-05-31 paper/presentation.tex
presentation: correction
presentation: improve some presenter comments, use inline citations

@@
 \usepackage{fancyvrb}
+\usepackage[style=authoryear,backend=bibtex]{biblatex}
 \setbeamertemplate{note page}[plain]
@@
 }
+
+\bibliography{paper}
 
@@
   \note{
-    Here is how this talk is going to be structured. First, I'll be giving a bit
-    of motivation for this project, along with a problem statement and our
-    background information. Next, I'll discuss related work in this field, and
-    show how this thesis fits into the current research. After that, I'll give
-    an overview of the implementation of the components of the thesis. Next,
-    I'll discuss the experiments I used to validate the mechanism, and finally
-    we'll conclude with a brief summary and discussion of future work.
+
+    \textbf{Transition:} Here is an overview of how this talk is organized.
+
+    \begin{itemize}
+    \item Start with an Introduction motivating this thesis, ending with a
+      problem statement
+    \item Background on Multipath TCP
+    \item Related work such as overlays
+    \item Implementation of the project
+    \item Experimental evaluation
+    \item Conclude with discussion of related work and a summary
+    \end{itemize}
+
   }
@@
 \FigureSlide{figures/InternetSlide.pdf}{
-  Okay, so let's get started! I want to start off with a brief and simplified
-  look at the Internet. As we know, the Internet is made up of routers, the blue
-  dots, and Autonomous Systems, the tan groups. The task of each router is to
-  forward small packets of information toward their final destination. The route
-  that packets take is determined by a lot of factors.
 
-  While this project isn't directly about routing, let's discuss some of these
-  factors. Routing is typically performed, in some way or another, by an
-  application of Dijkstra's algorithm for finding a shortest path. While this
-  might be optimal in a theoretical sense, the practical application is not, and
-  for several reasons.
+  \textbf{Transition:} I would like to start this talk with a brief, simplified
+  look at routing on the Internet.
 
-  For one, routing algorithms are going to work by ``hop count'',
-  meaning that a shorter path should be preferred, even if the longer path might
-  actually exhibit better properties. For another, the Autonomous Systems are
-  frequently businesses or ISPs, and the connections they make are determined by
-  business agreements. They may create policies that are less than optimal in
-  order to fulfill business objectives.
+  \begin{itemize}
+  \item As we know, routers, ASes, in blue and tan \textbf{respectively}
+  \item Routers forward packets based on routing tables
+  \item Tables built via ``distributed shortest path algorithms'', optimal in
+    terms of AS-hops
+  \item Inefficiencies result for particular paths
+    \begin{itemize}
+    \item AS hops doesn't take into account path bandwidth, congestion, etc
+    \item ASes are businesses, make policies and route based on financial bottom
+      line
+    \end{itemize}
+  \end{itemize}
+
 }
@@
   \item Peering agreements and policy based routing can result in suboptimal
-    routing decisions \cite{detour}
+    routing decisions \footnotemark
   \item A route that passes through a ``detour'' may be better
@@
     \includegraphics[width=\textwidth]{figures/detour.png} \\
-    \textit{Figure from \cite{detour}}
+    \textit{Example of an inefficient default route \footnotemark[1]}
   \end{column}
   \end{columns}
+  \footcitetext{detour}
   \note{
-    Research in the field has been able to fairly concretely demonstrate the
-    problem. Just because your traffic takes a particular route, doesn't mean it
-    is going to be the best path for you, in terms of latency or reliability.
 
-    Instead, if you were to route your packets to a ``detour'' and then to the
-    final destination, you might actually get path which has better round trip
-    time or loss rate. On the right we have an example of a situation found by a
-    study, where the default route from Los Altos to Walnut Creek had a round
-    trip time of 137 ms, while the combined round trip time of a path via Santa
-    Clara was only 30 ms. The reason was that the default route actually
-    traveled through Chicago!
+    \begin{itemize}
+    \item Not simply something we can hypothesize about.
+    \item Real world examples show that default routes aren't the best paths for
+      users.
+    \item In fact, many times you can add a ``detour'' to your route and end
+      up with an improved path.
+    \item At right, the default path from Los Altos to Walnut Creek has round
+      trip time of 137 milliseconds, while the constituent paths of a detour
+      route through Santa Clara add up to 40 milliseconds
+    \end{itemize}
+
   }
@@
 \FigureSlide{figures/detour-packetloss.png}{
-  This effect is not limited to isolated examples, either. This same study
-  looked at a large group of computers on university networks. They computed
-  latency and packet loss rates between each pair of computers. Then, they
-  searched for pairs in which the ``triangle inequality'' was violated.
 
-  This graph shows packet loss rates between pairs of computers. The blue is the
-  default Internet routed path, and the yellow is the best alternative path that
-  uses a single detour. Most of the yellow dots are actually on the zero line.
+  \begin{itemize}
+  \item Not just isolated examples.
+  \item Same study looked at a group of computers across universities
+  \item Observed pairwise RTT, loss rate.
+  \item Searched for pairs where ``triangle inequality violated'', i.e. lower
+    RTT or loss rate via detour.
+  \item In graph, blue is default route, yellow is best alternative route, and
+    we compare loss rate
+  \item For majority of pairs, alternative route is more reliable.
+  \item For some pairs, reliability difference is staggering
+  \end{itemize}
 
-  You can see here that for a large majority of pairs, there is an alternative
-  route that is more reliable. For a small fraction here, the difference in
-  reliability is actually huge.
 }
@@
   \frametitle{Access Link Underutilization}
+  % this slide reads as very stupid - at least the first bullet point does.
+  % let's think about making a better case for improving links, network core
+  % bottlenecks, etc
 
@@
   \item However, residential bandwidth is not fully utilized
-    \cite{fibertothehome}
+    \footcite{fibertothehome}
     \begin{itemize}
@@
   \item Once 3WHS completes, path manager requests a detour from client daemon
-  \item Client daemon receives request and sends UDP request to every client
-    daemon listed in configuration file
+  \item Client daemon receives request and sends UDP request to every detour
+    listed in configuration file
   \item Detour daemon sets up detour, sends reply
@@
 
-\section{References}
-\begin{frame}[allowframebreaks]
-  \frametitle{References}
-  \bibliographystyle{ieeetr}
-  \bibliography{paper.bib}
-\end{frame}
-
 \end{document}

### 2017-06-01 paper/presentation.tex
presentation: the rest of the paper refs
presentation: new citation style (cough cough)
presentation: improve notes dramatically, implementing some suggestions

@@
   \end{frame}
+}
+
+% fuckit
+\newcommand{\mycitetext}[1]{
+  \citeauthor{#1}. \citetitle{#1}. \citeyear{#1}
+}
+\newcommand{\mycite}[1]{
+  \footnote{\mycitetext{#1}}
 }
@@
   \end{columns}
-  \footcitetext{detour}
+  \footnotetext{\mycitetext{detour}}
   \note{
@@
   \item However, residential bandwidth is not fully utilized
-    \footcite{fibertothehome}
+    \mycite{fibertothehome}
     \begin{itemize}
@@
   \note{
-    Now let's take a brief detour and discuss something else. Recently, we've
-    been seeing some exciting trends in residential bandwidth. New access links
-    are always being created, and the latest improvement is fiber to the home.
-    These can provide up to Gigabit speeds to a home network. However, we've
-    had studies right here at CWRU showing that these residential links don't
-    use most of that capacity.
 
-    There are several good explanations for this. Short lived TCP sessions may
-    never leave the slow-start stage. A small TCP initial window will limit the
-    throughput that can be achieved in a reasonable amount of time. But there's
-    also the possibility that the network core is becoming the bottleneck for
-    some of these connections.
+    \textbf{Transition:} Take a brief detour and discuss residential access
+    speeds. (lol detour)
 
-    Two things seem to follow. First, there are good chances that a detour route
-    could provide better throughput for a TCP connection, as studies have
-    already demonstrated. Second, if we have extra access link capacity, why not
-    use these paths in combination with each other, rather than choosing only
-    one?
+    \begin{itemize}
+    \item Residential access links always increasing
+    \item Dial-up, broadband, cable, now fiber-to-the-home
+    \item Local study on fiber connections shows bandwidth not fully utilized
+      (go over each possibility, incl send buffer ``anemia'')
+    \item If network core can't support bandwidth, the detour routing approach
+      can help
+    \item If we have plenty of access bandwidth, why not aggregate paths we have
+      available to us?
+    \end{itemize}
+
   }
@@
   \note{
-    Putting this together, we have a problem: routing inefficiencies can prevent
-    TCP connections from achieving their full potential throughput. Access links
-    may be capable of much better bandwidth than a user actually gets.
 
-    Our solution is to create a system which can use and aggregate alternative
-    paths for TCP connections. Using these paths means that it can route traffic
-    along them. Aggregation means combining the potential throughput of both
-    paths, to get better throughput than either.
+    \begin{itemize}
+    \item Putting together, problem: routing inefficiencies can prevent TCP
+      connections from achieving throughput.
+    \item Furthermore, while detour approaches are plentiful, and even a few
+      which use multiple paths, as well as application-level strategies, there
+      are no strategies which are usable by unmodified applications
+    \item A solution could be a system which uses and aggregates alternative
+      routes for MPTCP connections.
+    \item Contribution 1 - a mechanism for adding new routes to MPTCP
+      connections which are single-homed, with no modification to applications.
+    \item Contribution 2 - experimental validation that the mechanism works,
+      even at higher throughputs
+    \end{itemize}
 
-    The contributions this thesis makes are twofold: first, a mechanism that
-    allows single-homed devices to add subflows along alternative routes via
-    MPTCP. Second, an evaluation of this system on emulated and real world
-    networks.
   }
@@
 \section{Background}
-\SectionHead{}{}
+\SectionHead{}{
+  \textbf{Transition:} Before explaining my solution in more detail, we'll need
+  some background on Multipath TCP.
+}
 \subsection{Multipath TCP}
@@
 
-      This project is based on MPTCP. While it's not my intent to give a
-      comprehensive protocol description in this presentation, I do want to give
-      enough background to understand how my project uses it.
+      \begin{itemize}
+      \item Project is based on MPTCP, which enables multi-path transport for
+        unmodified applications in a way not possible before
+      \item Purpose is not to give complete description of MPTCP, but an
+        overview
+      \item Canonical motivation for MPTCP: smartphone with multiple interfaces.
+      \item Also servers in datacenters (redundant topo) or laptops (wifi,
+        ethernet, cell?)
+      \item \textit{(See remaining slide bullets, they're good)}
+      \end{itemize}
 
-      Recently, we've seen a rise in devices that have multiple ways to connect
-      to the Internet, network interfaces. For instance, smartphones include
-      cellular and WiFi radios, and can be connected to the Internet on both.
-      Recent data centers have been taking advantage of multiple interfaces and
-      highly redundant network topologies. Laptops can also contain WiFi,
-      Ethernet, and even cellular radios.
-
-      Unfortunately, TCP was designed for a simpler time, when computers had a
-      single interface and their addresses never changed. This forces
-      multi-homed devices to decide which interface they'll use and stick to
-      that choice.
-
-      Multipath TCP aims to solve this problem. It is an extension to TCP (not a
-      replacement) that allows devices to make use of all of their network
-      interfaces and addresses in a single connection.
     }
@@
   \begin{itemize}
+  \item Remain compatible with TCP applications and the Internet
+    \begin{itemize}
+    \item \textit{Present the same socket API to applications}
+    \item Remain similar to TCP on the wire, to remain compatible with Internet
+      middleboxes
+    \end{itemize}
   \item Improve performance and reliability over current TCP, by aggregating
@@
     bottlenecks than standard TCP would
-  \item Remain compatible with TCP applications and the Internet
-    \begin{itemize}
-    \item Present the same socket API to applications
-    \item Remain similar to TCP on the wire, to remain compatible with Internet
-      middleboxes
-    \end{itemize}
   \end{itemize}
   \note{
-    As a protocol, MPTCP has a few key design goals. First, it is designed to
-    improve performance and reliability compared to standard TCP. Second, it
-    should not do any harm to existing single-path TCP. Third, it should remain
-    compatible to applications and to the Internet as a whole.
+
+    \textbf{Transition:} Multipath TCP's design is motivated by several goals.
+
+    \begin{itemize}
+    \item Primary goal: compatibility
+    \item Particularly API compatibility. Runs with unmodified applications
+    \item Internet compat too. Middleboxes (NAT, Firewall, Normalizers, etc)
+      should not be confused by, or corrupt, MPTCP
+    \item Remaining two: performance, but not at the expense of existing TCP
+    \end{itemize}
+
   }
@@
   \note{
-    The way that MPTCP achieves this is by layering the protocol extension in
-    between TCP and the application layer. An MPTCP connection consists of one
-    or more (well, potentially zero or more) TCP connections, which we refer to
-    as ``subflows''. These subflows bear some extra TCP options which provide
-    signaling information, but other than that they behave very similarly to
-    standard TCP. The only major difference is that their receive window is
-    shared, rather than per-subflow. The MPTCP connection has its own sequence
-    numbers, and each subflow has independent sequence numbers as well.
+
+    \begin{itemize}
+    \item Achieved by layering between TCP and application
+    \item \textbf{Presents exact same API}
+    \item \textbf{Unmodified applications can use it}
+    \item Rather than creating single TCP flow, create MPTCP ``connection''
+      which may contain multiple ``subflows''
+    \item Subflows look very similar to regular TCP, except some TCP options
+    \item Share receive window
+    \item Subflows have independent sequence numbers
+    \item Data has sequence numbers too
+    \end{itemize}
+
   }
@@
   \item Either side may create new subflows at any time
+  \item Spec says nothing about when they should be created, left up to policy
+  \item Kernel uses modular path manager interface for this
   \end{itemize}
   \note{
-
-    When we create an MPTCP connection, a three-way handshake occurs, just like
-    standard TCP. However, the three-way handshake includes the MP CAPABLE
-    option, signalling to the other host that it supports MPTCP. A SYNACK with
-    MP CAPABLE signals that the passive opener supports MPTCP as well, and the
-    final ACK with MP CAPABLE confirms the use of MPTCP. The contents of these
-    options are key material, which is used to authenticate creation of new
-    subflows.
-
-    Subsequent subflows are created with a similar procedure, but using the MP
-    JOIN option. The contents of the option this time are HMAC signatures based
-    on the key material. If the HMAC can be verified, the hosts can be certain
-    that they are communicating with the same host they initially established a
-    connection with.
-
-    An MPTCP implementation may advertise that it has additional addresses (and
-    therefore additional interfaces) via an ADD ADDR option. There is also
-    support for deletion.
-
-    Finally, the MPTCP specification says nothing about when subflows should be
-    created or destroyed. This is entirely left up to local policy, and of
-    course research. In the Linux kernel implementation of MPTCP, a module
-    interface exists for this, called the path manager.
 
@@
 
-  So, now that we have subflows, what do we do with them? How do we get data
-  through them to the other application? Well, we do this with two important
-  concepts: a data scheduler and the data sequence mapping.
+  \textbf{Transition:} Now that we have subflows, how do we split up data across
+  them?
 
-  Imagine one host wants to send some data. It passes it through the send system
-  call, and eventually the MPTCP stack has a large piece of data. A component
-  called the scheduler gets to make the executive decision of how to divide up
-  the data. It does this however it pleases, and the chunks make their way down
-  to the subflows.
-
-  In order to be able to put them back together, a data sequence signal is sent
-  on the subflow.
+  \begin{itemize}
+  \item Again, MPTCP specifies mechanism but not policy
+  \item Diagram top shows sender
+  \item Data goes into scheduler, which divides it across subflows (will come
+    back to that)
+  \item To send it across the wire, MPTCP adds DSS option, which maps subflow
+    sequence number X to data sequence number Y, for Z bytes
+  \item At receiver, use options to reassemble data
+  \item Data may be redundantly transferred, can't break subflow TCP semantics
+  \item Scheduler has considerable freedom in dividing data
+  \item Lowest RTT first ... ack-clocking
+  \end{itemize}
 
@@
 \SectionHead{}{
-  Would love to discuss more, but limited time
+  \textbf{Transition:} Now in order to appreciate the place of this thesis,
+  let's examine some related work. The topic includes detour routing, overlay
+  routing, source routing, and multipath transport.
 }
@@
 \begin{frame}
-  \frametitle{Resilient Overlay Networks}
+  \frametitle{Resilient Overlay Networks\mycite{ron}}
 
@@
     \begin{itemize}
-    \item Redundant multipath routing
+    \item Redundant multipath routing\mycite{andersen2003best}
       \note[item]{One study compared two routing strategies: adaptively choosing
@@
   \item ``Biologically inspired'' multipath routing
+    \mycite{leibnitz2006biologically}
     \note[item]{Another - biologically inspired. Choose strongest path as
@@
       randomness.}
-  \item mTCP \note[item]{mTCP: most important. A modification of TCP before
+  \item mTCP \mycite{zhang2004transport}
+    \note[item]{mTCP: most important. A modification of TCP before
       MPTCP existed. Built on RON + userspace TCP stack. Did not have MPTCP
@@
   \end{itemize}
+
+  \note[item]{\textbf{Remember:} \textit{No ``last piece'' of related work.}}
 \end{frame}
@@
   \item Binder: aggregating bandwidth of several Internet gateways
+    \mycite{binder}
     \note[item]{One use of MPTCP that is very similar was called Binder. A rural
@@
       server, with subflow across each.}
-  \item SOSR: Scalable One-hop Source Routing \note[item]{Work inspired by RON,
+  \item SOSR: Scalable One-hop Source Routing \mycite{gummadi2004improving}
+    \note[item]{Work inspired by RON,
       but trying to simplify. This work simply tunneled connections across a
@@
   \end{itemize}
+
+  \note[item]{\textbf{Remember:} \textit{No ``last piece'' of related work.}}
+
+  % TODO: may want to swap out some less interesting things with BitTorrent /
+  % gnutella / other peer-to-peer apps
 \end{frame}
@@
 \SectionHead{}{
-  Time to get into the actual implementation of this project.
+
+  \textbf{Transition:} Now that we have seen work related to these
+  topics, let's discuss how I was able to implement the system described -
+  one-hop detour routing, path aggregation, and transparency to the application.
+
 }
@@
 \section{Evaluation}
-\SectionHead{}{}
+\SectionHead{}{
+  \textbf{Transition:} With that implementation in mind, we will now discuss the
+  experiments I designed to validate this system.
+}
 

### 2017-06-01 paper/presentation.tex
presentation: add forgotten presenter notes
presentation: title some figure slides

@@
 
-\newcommand{\TitledFigureSlide}[2]{
+\newcommand{\TitledFigureSlide}[3]{
   \begin{frame}{#2}
@@
       \end{centering}
+      \note{#3}
   \end{frame}
@@
 
-\FigureSlide{figures/InternetSlide.pdf}{
+\TitledFigureSlide{figures/InternetSlide.pdf}{Internet Architecture}{
 
@@
   \item Either side may create new subflows at any time
+  \end{itemize}
+  \note{
+
+    \begin{itemize}
+    \item Normal 3WHS for subflow initiation
+    \item First subflow contains MP CAPABLE, exchanges key material
+    \item Subsequent subflows contain MP JOIN, plus HMAC to verify identities
+    \item Either side may passively announce the existence of an address
     \item Spec says nothing about when they should be created, left up to policy
@@
     \end{itemize}
-  \note{
 
@@
 
-\FigureSlide{figures/Concept.pdf}{
+\TitledFigureSlide{figures/Concept.pdf}{Concept Overview}{
   \begin{itemize}
@@
 \section{All Figures}
-\TitledFigureSlide{figures/sym.pdf}{Throughput Comparison, Symmetric}
-\TitledFigureSlide{figures/timegrid-sym.pdf}{Timelapse, Symmetric}
+\TitledFigureSlide{figures/sym.pdf}{Throughput Comparison, Symmetric}{}
+\TitledFigureSlide{figures/timegrid-sym.pdf}{Timelapse, Symmetric}{}
 
-\TitledFigureSlide{figures/sym-lossy.pdf}{Throughput Comparison, Symmetric with Loss}
-\TitledFigureSlide{figures/timegrid-sym-lossy.pdf}{Timelapse, Symmetric with Loss}
+\TitledFigureSlide{figures/sym-lossy.pdf}{Throughput Comparison, Symmetric with Loss}{}
+\TitledFigureSlide{figures/timegrid-sym-lossy.pdf}{Timelapse, Symmetric with Loss}{}
 
-\TitledFigureSlide{figures/sym-delayed.pdf}{Throughput Comparison, Symmetric with Delay}
-\TitledFigureSlide{figures/timegrid-sym-delayed.pdf}{Timelapse, Symmetric with Delay}
+\TitledFigureSlide{figures/sym-delayed.pdf}{Throughput Comparison, Symmetric with Delay}{}
+\TitledFigureSlide{figures/timegrid-sym-delayed.pdf}{Timelapse, Symmetric with Delay}{}
 
-\TitledFigureSlide{figures/easy.pdf}{Throughput Comparison, Core-limited}
-\TitledFigureSlide{figures/timegrid-easy.pdf}{Timelapse, Core-limited}
+\TitledFigureSlide{figures/easy.pdf}{Throughput Comparison, Core-limited}{}
+\TitledFigureSlide{figures/timegrid-easy.pdf}{Timelapse, Core-limited}{}
 
-\TitledFigureSlide{figures/lossy.pdf}{Throughput Comparison, Core-limited with Loss}
-\TitledFigureSlide{figures/timegrid-lossy.pdf}{Timelapse, Core-limited with Loss}
+\TitledFigureSlide{figures/lossy.pdf}{Throughput Comparison, Core-limited with Loss}{}
+\TitledFigureSlide{figures/timegrid-lossy.pdf}{Timelapse, Core-limited with Loss}{}
 
-\TitledFigureSlide{figures/delayed.pdf}{Throughput Comparison, Core-limited with Delay}
-\TitledFigureSlide{figures/timegrid-delayed.pdf}{Timelapse, Core-limited with Delay}
+\TitledFigureSlide{figures/delayed.pdf}{Throughput Comparison, Core-limited with Delay}{}
+\TitledFigureSlide{figures/timegrid-delayed.pdf}{Timelapse, Core-limited with Delay}{}
 
-\TitledFigureSlide{figures/aws.pdf}{Throughput Comparison, AWS}
-\TitledFigureSlide{figures/timegrid-aws.pdf}{Timelapse, AWS}
+\TitledFigureSlide{figures/aws.pdf}{Throughput Comparison, AWS}{}
+\TitledFigureSlide{figures/timegrid-aws.pdf}{Timelapse, AWS}{}
 

### 2017-06-01 src/experiment.py
experiment: add tcpdump option and create a capture script

@@ -299,7 +299,7 @@ PARAMS = {
 }
 
 
-def scenario(setup_name, params_name, trials=30):
+def scenario(setup_name, params_name, trials=30, trace=False):
     params = PARAMS[params_name]()
     mn = DetourNet(params)
     iperf_cmd = SETUP[setup_name](mn)
@@ -307,7 +307,15 @@ def scenario(setup_name, params_name, trials=30):
 
     client, detour, server = mn.get('client', 'detour', 'server')
 
-    filename = '%s.%s.%s.json' % (params_name, setup_name, name)
+    fn_base = '%s.%s.%s' % (params_name, setup_name, name)
+    filename = fn_base + '.json'
+    tcpdump_fn = fn_base + '.pcap'
+    # 48 byte ethernet
+    # 20 byte ip (hope there's no options)
+    # 20 byte tcp (not including options)
+    # 40 byte maximum TCP option space
+    # = 128 bytes captured
+    server.cmd('tcpdump -U -s 128 -w %s tcp port 5201 &' % tcpdump_fn)
     server.sendCmd('iperf3 -s -J > %s' % filename)
     print(filename + ': ', end='')
 
@@ -339,6 +347,8 @@ def main():
                         help='comma separated list of setups')
     parser.add_argument('--cli', action='store_true',
                         help='start a cli with the first param and setup')
+    parser.add_argument('--trace', action='store_true',
+                        help='create tcpdump trace at server')
     parser.add_argument('--trials', '-t', type=int, default=60)
 
     args = parser.parse_args()
@@ -353,7 +363,7 @@ def main():
     else:
         for param in params:
             for setup in setups:
-                scenario(setup, param, args.trials)
+                scenario(setup, param, trials=args.trials, trace=args.trace)
 
 
 if __name__ == '__main__':

### 2017-06-01 src/make_captures.sh

@@ -1,0 +1,26 @@
+#!/bin/sh
+
+TRIALS=1
+PARAMS=sym
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+
+PARAMS=sym_lossy
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+
+PARAMS=sym_delayed
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+
+PARAMS=easy
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+
+PARAMS=lossy
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+
+PARAMS=delayed
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn

### 2017-06-02 data/capplot.sh
data: add capplot.sh for automatically creating capture plots

@@ -1,0 +1,70 @@
+#!/usr/bin/env bash
+#
+# Create tcpdump capture plots for experiment captures.
+#
+# This script is very, very particular and very, very hacky. It also depends on
+# some exotic tooling, at least one of which is not publicly available. Sorry.
+# The script will output a gnuplot script, along with the supporting files. The
+# script will in fact be suitable to run in batch, producing a PDF. You can edit
+# it after the fact for better titles, etc.
+#
+# usage: ./capplot.sh capturename.pcap               # must have pcap extension
+# output:
+#  capturename.pdf
+#  capturename.gpl
+#  capturename.labels
+#  capturename.datasets
+
+# Fail on first error
+set -e
+# Fail on unset var
+set -u
+# Print commands
+#set -x
+
+# Requires mptcpparser, a tool which is not publicly available.
+MPTCPPARSER=$HOME/code/mptcptools/mptcpparser
+
+# Requires xpl2gpl - http://www.tcptrace.org/xpl2gpl/
+XPL2GPL=$HOME/code/xpl2gpl
+
+FILENAME=$(basename "$1")
+BASE="${FILENAME%.*}"
+DIR=`mktemp -d`
+
+cp $1 $DIR/input.pcap
+pushd $DIR
+
+# packet trace -> xplot file
+$MPTCPPARSER input.pcap
+
+# xplot file -> gnuplot file
+$XPL2GPL connection_1-ORIGIN.xpl
+
+# Make PDF output!
+sed -i 's/.*set term.*/set term pdf/' connection_1-ORIGIN.gpl
+sed -i 's/.*set output.*/set output "'$BASE.pdf'"/' connection_1-ORIGIN.gpl
+
+# Replace names with the original trace name.
+sed -i "s/connection_1-ORIGIN/$BASE/g" connection_1-ORIGIN.gpl
+
+# Move the plot command to the end.
+grep '^plot.*' connection_1-ORIGIN.gpl >> tmp
+sed -i '/.*plot.*/d' connection_1-ORIGIN.gpl
+sed -i '/.*replot.*/d' connection_1-ORIGIN.gpl
+sed -i '/.*pause.*/d' connection_1-ORIGIN.gpl
+cat connection_1-ORIGIN.gpl tmp > $BASE.gpl
+# now our finished product is $BASE.gpl
+
+popd
+
+# Put output files into place.
[…]

### 2017-06-03 data/2017-05-21-cpu-limit/plots.py
data: a little filtration goes a long way

@@ -45,7 +45,7 @@ def make_single_plot(base):
     ax.boxplot(data, 0, '', labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
     ax.set_title('Throughput Comparison: ' + META[base])
-    fig.savefig('%s.pdf' % base)
+    fig.savefig('%s.pdf' % base.replace('.', '-').replace('_', '-'))
 
 
 def make_all_plots():
@@ -68,7 +68,7 @@ def make_time_series_grid(base):
     axarr[1][0].set_xlabel('Control')
     axarr[1][1].set_xlabel('NAT')
     axarr[1][2].set_xlabel('VPN')
-    fig.savefig('timegrid.%s.pdf' % base)
+    fig.savefig('timegrid-%s.pdf' % base.replace('.', '-').replace('_', '-'))
 
 
 def make_all_time_series_grid():

### 2017-06-03 data/output.py

@@ -28,9 +28,20 @@ def json_parse(filename):
 def objects_to_bps(obj_list):
     """Converts a sequence of JSON objects to the bits per second."""
     for obj in obj_list:
-        bps = obj.get('end', {}).get('sum_received', {}).get('bits_per_second')
-        if bps:
-            yield bps
+        if 'error' in obj.keys():
+            return
+        # First, filter out the intervals that are after 1 second (slow start)
+        # and before the 10 second mark (when the connection winds down).
+        # Then compute bits per second and return it.
+        intervals = []
+        for interval in obj['intervals']:
+            if interval['sum']['start'] >= 0.9 and interval['sum']['end'] <= 10.1:
+                intervals.append(interval)
+
+        seconds = intervals[-1]['sum']['end'] - intervals[0]['sum']['start']
+        bytes_ = sum(x['sum']['bytes'] for x in intervals)
+
+        yield bytes_ * 8 / seconds
 
 
 def objects_to_utilization(obj_list):

### 2017-06-03 data/capplot.sh
data: include xplots and mappings and data and more

@@ -10,10 +10,7 @@
 #
 # usage: ./capplot.sh capturename.pcap               # must have pcap extension
 # output:
-#  capturename.pdf
-#  capturename.gpl
-#  capturename.labels
-#  capturename.datasets
+#   many files, all starting with capturename. Just read the script
 
 # Fail on first error
 set -e
@@ -25,9 +22,6 @@ set -u
 # Requires mptcpparser, a tool which is not publicly available.
 MPTCPPARSER=$HOME/code/mptcptools/mptcpparser
 
-# Requires xpl2gpl - http://www.tcptrace.org/xpl2gpl/
-XPL2GPL=$HOME/code/xpl2gpl
-
 FILENAME=$(basename "$1")
 BASE="${FILENAME%.*}"
 DIR=`mktemp -d`
@@ -38,33 +32,15 @@ pushd $DIR
 # packet trace -> xplot file
 $MPTCPPARSER input.pcap
 
-# xplot file -> gnuplot file
-$XPL2GPL connection_1-ORIGIN.xpl
-
-# Make PDF output!
-sed -i 's/.*set term.*/set term pdf/' connection_1-ORIGIN.gpl
-sed -i 's/.*set output.*/set output "'$BASE.pdf'"/' connection_1-ORIGIN.gpl
-
-# Replace names with the original trace name.
-sed -i "s/connection_1-ORIGIN/$BASE/g" connection_1-ORIGIN.gpl
-
-# Move the plot command to the end.
-grep '^plot.*' connection_1-ORIGIN.gpl >> tmp
-sed -i '/.*plot.*/d' connection_1-ORIGIN.gpl
-sed -i '/.*replot.*/d' connection_1-ORIGIN.gpl
-sed -i '/.*pause.*/d' connection_1-ORIGIN.gpl
-cat connection_1-ORIGIN.gpl tmp > $BASE.gpl
-# now our finished product is $BASE.gpl
-
 popd
 
 # Put output files into place.
-mv $DIR/$BASE.gpl .
-mv $DIR/connection_1-ORIGIN.labels $BASE.labels
-mv $DIR/connection_1-ORIGIN.datasets $BASE.datasets
+mv $DIR/connection_1-ORIGIN.xpl $BASE.xpl
+mv $DIR/connection_1-MAPPING.txt $BASE.mapping.txt
 
-# Generate the pdf
-gnuplot $BASE.gpl
[…]

### 2017-06-03 src/experiment.py
experiment: don't trace unless asked to

@@ -315,7 +315,8 @@ def scenario(setup_name, params_name, trials=30, trace=False):
     # 20 byte tcp (not including options)
     # 40 byte maximum TCP option space
     # = 128 bytes captured
-    server.cmd('tcpdump -U -s 128 -w %s tcp port 5201 &' % tcpdump_fn)
+    if trace:
+        server.cmd('tcpdump -U -s 128 -w %s tcp port 5201 &' % tcpdump_fn)
     server.sendCmd('iperf3 -s -J > %s' % filename)
     print(filename + ': ', end='')
 

### 2017-06-03 data/2017-05-21-cpu-limit/plots.py
data: more honest plots

@@ -42,7 +42,7 @@ def make_single_plot(base):
 
     fig, ax = plt.subplots()
     print(list(map(len, data)))
-    ax.boxplot(data, 0, '', labels=labels)
+    ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
     ax.set_title('Throughput Comparison: ' + META[base])
     fig.savefig('%s.pdf' % base.replace('.', '-').replace('_', '-'))

### 2017-06-03 src/experiment.py
experiment: a wider range of parameters

@@ -237,24 +237,35 @@ def params_easy():
     return BASIC_PARAMS.copy()
 
 
-def params_lossy():
+def params_loss_rate(rate):
     """
     This scenario is similar to 'easy', but there is loss along the default
     route.
     """
-    d = BASIC_PARAMS.copy()
-    d['r1_r2']['loss'] = 1
-    return d
+    def inner():
+        d = BASIC_PARAMS.copy()
+        d['r1_r2']['loss'] = rate
+        return d
+    return inner
 
 
-def params_delayed():
+def params_lossy():
+    params_loss_rate(1)()
+
+
+def params_delay_ms(ms):
     """
     This scenario is similar to 'easy', but there is a high latency link across
     the default route.
     """
-    d = BASIC_PARAMS.copy()
-    d['r1_r2']['delay'] = '100ms'
-    return d
+    def inner():
+        d = BASIC_PARAMS.copy()
+        d['r1_r2']['delay'] = str(ms) + 'ms'
+        return d
+    return inner
+
+def params_delayed():
+    params_delay_ms(100)()
 
 
 def params_sym():
@@ -269,33 +280,47 @@ def params_sym():
     return d
 
 
-def params_sym_lossy():
+def params_sym_loss_rate(rate):
     """
     This scenario, based on symmetric, has each link with the same bandwidth.
     However, the default route is lossy. Thus, the two flows are competing for
     bandwidth and the alternative subflow needs to make up the total bandwidth.
     """
-    d = params_sym()
-    d['r1_r2']['loss'] = 1
-    return d
+    def inner():
+        d = params_sym()
+        d['r1_r2']['loss'] = rate
+        return d
+    return inner
 
 
-def params_sym_delayed():
+def params_sym_delay_ms(ms):
     """
     This scenario is based on symmetric, but the default route has high delay.
     """
-    d = params_sym()
-    d['r1_r2']['delay'] = '100ms'
-    return d
+    def inner():
+        d = params_sym()
+        d['r1_r2']['delay'] = str(ms) + 'ms'
+        return d
+    return inner
 
 
 PARAMS = {
     'easy': params_easy,
-    'lossy': params_lossy,
-    'delayed': params_delayed,
+    'lossy05': params_loss_rate(0.5),
+    'lossy': params_loss_rate(1),
+    'lossy2': params_loss_rate(2),
[…]

### 2017-06-03 src/mn_param_range.sh

@@ -1,0 +1,44 @@
+#!/bin/sh
+
+TRIALS=20
+PARAMS=sym_lossy05
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=sym_lossy2
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=sym_lossy5
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=lossy05
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=lossy2
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=lossy5
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+
+PARAMS=sym_delayed20
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=sym_delayed50
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=delayed20
+./experiment.py -t $TRIALS --params $PARAMS --setup control
+./experiment.py -t $TRIALS --params $PARAMS --setup nat
+./experiment.py -t $TRIALS --params $PARAMS --setup vpn
+PARAMS=delayed50
[…]

### 2017-06-03 data/param_validation/plots.py
Add parameter validation data.

@@ -1,0 +1,87 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved (thanks python 3.6...?)
+META = {
+    'lossy05': 'Core-limited, 0.5% Loss',
+    'lossy2': 'Core-limited, 2% Loss',
+    'lossy5': 'Core-limited, 5% Loss',
+    'delayed20': 'Core-limited, 20ms Delay',
+    'delayed50': 'Core-limited, 50ms Delay',
+    'sym_lossy05': 'Symmetric, 0.5% Loss',
+    'sym_lossy2': 'Symmetric, 2% Loss',
+    'sym_lossy5': 'Symmetric, 5% Loss',
+    'sym_delayed20': 'Symmetric, 20ms Delay',
+    'sym_delayed50': 'Symmetric, 50ms Delay',
+}
+
+# order preserved
+SCEN = {
+    'control.mptcp': '1 Subflow',
+    'nat.mptcp': 'NAT',
+    'vpn.mptcp': 'VPN',
+    'control.vanilla': 'TCP',
+    'nat.vanilla': 'TCP (NAT)',
+    'vpn.vanilla': 'TCP (VPN)',
+}
+
+
+def make_single_plot(base):
+    print(base)
+    data = []
+    labels = []
+    for s, label in SCEN.items():
+        data.append(o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000)
+        labels.append(label)
+
+    fig, ax = plt.subplots()
+    print(list(map(len, data)))
+    ax.boxplot(data, labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: ' + META[base])
+    fig.savefig('%s.pdf' % base.replace('.', '-').replace('_', '-'))
+
+
+def make_all_plots():
+    for base in META.keys():
+        make_single_plot(base)
+
+
+def make_time_series_grid(base):
+    print(base)
+    rows = ['mptcp', 'vanilla']
+    cols = ['control', 'nat', 'vpn']
+    fig, axarr = plt.subplots(len(rows), len(cols), sharex=True, sharey=True)
+    for ridx, row in enumerate(rows):
+        for cidx, col in enumerate(cols):
[…]

### 2017-06-04 etc/aws_exp_setup_vanilla.sh
aws: let's try to do this with vanilla TCP too

@@ -1,5 +1,5 @@
 #!/bin/bash
-# Run as root. Activates MPTCP kernel.
+# Run as root. Does not activate MPTCP kernel.
 cd
 
 export DEBIAN_FRONTEND=noninteractive
@@ -21,14 +21,8 @@ echo Installing required packages...
 apt install -y -q build-essential python3-pip libnl-genl-3-dev openvpn iperf3 tmux libconfig-dev
 echo Installing Python packages...
 pip3 install psutil
-echo Installing custom kernel...
-dpkg -i ~ubuntu/tmp/*.deb
 
-echo Setting default kernel...
-SUBMENU=$(egrep -o -m 1 "gnulinux-advanced[0-9a-zA-Z-]+" /boot/grub/grub.cfg)
-BOOT_OPTION=$(egrep -o -m 1 "gnulinux-4.1.38-custom-advanced[0-9a-zA-Z-]+" /boot/grub/grub.cfg)
-sed -i.bak "s/#\?GRUB_DEFAULT.*/GRUB_DEFAULT=\"$SUBMENU>$BOOT_OPTION\"/g" /etc/default/grub
-update-grub
+echo Not bothering with the MPTCP kernel, have fun with vanilla TCP
 
 echo Building tools...
 sudo -u ubuntu make -C ~ubuntu/src clean
@@ -36,9 +30,6 @@ sudo -u ubuntu make -C ~ubuntu/src all
 
 echo Adding ssh keys...
 echo "ssh-rsa AAAAB3NzaC1yc2EAAAADAQABAAABAQDEupr15Lbq8jFBca/tXhydvD2nKQ/vWnxXzOndG7TqKGy53mtlgR6VzaV4WR3blrLdDG223a64VlQaj00xm+8e/eAn+IRH3RpV2ZvFT+BdZmej8E4y0sSlkiCw2sjtyKiGu0Pyk2HKbA8njG+aVsJ5YSQVwBBP1nEqKoMItvXzZwQJQ09SCeklT5zugcJz+frGFifjz1I6bs7Hb9vNa7hNi79JjJ1rNoeANkBmunW20+M7hMOuNjGWl47XGDlew7tn3vc2Eiiohf0cyrueBPq63Y7tMEfhmyCEPfakkjg5idt3soRGKCBqyoVuYyEIOJSTnuGBkThZteGHLlpKvMVz stephen@greed_2013-09-26" >> ~ubuntu/.ssh/authorized_keys
-
-echo Setting congestion control to lia
[…]

### 2017-06-04 src/aws_experiment.py

@@ -17,14 +17,13 @@ Prerequisites:
    programmatically.
 
 """
-import collections
 import os
+import re
 import sys
 from io import BytesIO
 from os.path import abspath
 from os.path import dirname
 from os.path import join
-from pprint import pprint
 
 import boto3
 import paramiko
@@ -58,6 +57,7 @@ KEYFILE = os.getenv('KEYFILE') or '/home/stephen/.ssh/id_rsa'
 
 # The filename of the AWS setup script.
 SETUP_SCRIPT = join(dirname(abspath(__file__)), '../etc/aws_exp_setup.sh')
+SETUP_SCRIPT_VANILLA = join(dirname(abspath(__file__)), '../etc/aws_exp_setup_vanilla.sh')
 
 # These tags are applied to everything we create, so it's easier to clean up
 # via the Console if things go wrong.
@@ -133,22 +133,23 @@ def merge(d1, d2):
 
 def resource_params(k):
     v = merge(RESOURCE_PARAMS, RESOURCE_PARAMS_EACH[k])
-    #pprint(v)
     return v
 
 
 def create_instances_params(k):
     v = merge(CREATE_INSTANCES_PARAMS, CREATE_INSTANCES_PARAMS_EACH[k])
-    #pprint(v)
     return v
 
 
-def create_instances():
+def create_instances(vanilla=False):
     instances = {}
     for k in RESOURCE_PARAMS_EACH.keys():
         print('Creating instance %s...' % k)
         resource = boto3.resource('ec2', **resource_params(k))
-        instance = resource.create_instances(**create_instances_params(k))
+        params = create_instances_params(k)
+        if vanilla:
+            params['UserData'] = open(SETUP_SCRIPT_VANILLA).read()
+        instance = resource.create_instances(**params)
         assert len(instance) == 1
         instances[k] = instance[0]
     return instances
@@ -201,14 +202,15 @@ def ssh_connect(instances):
     return clients
 
 
-def many_iperf(n, client_transport, server_chan, server_output, instances):
+def many_iperf(n, client_transport, server_chan, server_output, instances, addr=None):
     print('IPERF: ', end='')
     sys.stdout.flush()
+    if addr is None:
+        addr = instances['server'].public_ip_address
     for _ in range(n):
         client_chan = client_transport.open_session()
         client_chan.set_combine_stderr(True)
-        client_chan.exec_command(command='iperf3 -c %s -J' %
[…]

### 2017-06-04 src/aws_experiment.py
aws: decode output and print exceptions before ipdb

@@ -402,10 +402,10 @@ def run_exp_vanilla(instances, clients):
     # We need to wait until the openvpn interface is up. This will also
     # give us the peer IP address, which we already configured, but it's
     # always better to dynamically determine.
-    output = client_daemon_chan.recv(4096)
+    output = client_daemon_chan.recv(4096).decode('utf8')
     regexp = re.compile(r'ip addr add dev tun0 local ([0-9.]+) peer ([0-9.]+)')
     while not regexp.search(output):
-        output += client_daemon_chan.recv(4096)
+        output += client_daemon_chan.recv(4096).decode('utf8')
     match = regexp.search(output)
     local, peer = match.groups()
 
@@ -448,6 +448,11 @@ def main(vanilla=False):
             run_exp_vanilla(instances, clients)
         else:
             run_exp_mptcp(instances, clients)
+
+    except:
+        # report it so that debugging is better
+        import traceback
+        print(traceback.format_exc())
 
     finally:
         print('Opening ipdb... When you exit, we\'re done here.')

### 2017-06-04 paper/presentation.tex
presentation: switch to bytefield where I can
presentation: supplement Intro w/ application-layer arguments

@@
 \usepackage{fancyvrb}
+\usepackage{bytefield}
 \usepackage[style=authoryear,backend=bibtex]{biblatex}
@@
       trip time of 137 milliseconds, while the constituent paths of a detour
-      route through Santa Clara add up to 40 milliseconds
+      route through Santa Clara add up to 30 milliseconds
     \end{itemize}
@@
     we compare loss rate
-  \item For majority of pairs, alternative route is more reliable.
-  \item For some pairs, reliability difference is staggering
+  \item For 80\% of pairs, alternative route is more reliable. For 25\% of
+    pairs, alternative route improves loss rate by at least a factor of 6
+  \item For some pairs, reliability difference is staggering!
   \end{itemize}
@@
   \frametitle{Access Link Underutilization}
-  % this slide reads as very stupid - at least the first bullet point does.
-  % let's think about making a better case for improving links, network core
-  % bottlenecks, etc
-
   \begin{itemize}
-  \item Fiber-to-the-home is coming!
+  \item Residential bandwidth constantly improves
   \item However, residential bandwidth is not fully utilized
@@
     \item Short-lived TCP sessions?
-    \item Small TCP initial window?
+    \item Anemic send buffers?
     \item Network core can't support bandwidth?
@@
 
+\TitledFigureSlide{figures/Concept.pdf}{Concept}{
+  \begin{itemize}
+  \item Detour routing has shown promise for a long time
+  \item Some applications have been able to leverage it (e.g. Bittorrent)
+  \item However, few unmodified applications may use the detour routing strategy
+  \item Our concept is to use MPTCP to achieve detour routing in unmodified
+    applications
+  \item We use tunneling mechanisms to put a subflow across a detour host
+  \end{itemize}
+}
+
 \begin{frame}
@@
 
-  \textbf{Problem:} Internet routing inefficiencies can prevent TCP connections
-  from achieving their full potential throughput.
+  \textbf{Problem:} Unmodified applications cannot use detour routing to
+  circumvent Internet routing inefficiencies.
 
@@
 
-  \textbf{Solution:} A system which can use and aggregate alternative routes for
-  TCP connections.
+  \textbf{Solution:} An OS-level detour routing system that leverages Multipath
+  TCP (MPTCP).
 
@@
   \begin{itemize}
-  \item A mechanism for adding alternative routes to Multipath TCP (MPTCP)
-    sessions among single-homed hosts.
-  \item An evaluation of this mechanism on emulated and real-world networks.
+  \item A method for performing detour routing with unmodified applications
+  \item A prototype implementation in the Linux kernel
+  \item An evaluation of this mechanism on emulated networks and the Internet
   \end{itemize}
@@
     \begin{itemize}
-    \item Putting together, problem: routing inefficiencies can prevent TCP
-      connections from achieving throughput.
-    \item Furthermore, while detour approaches are plentiful, and even a few
-      which use multiple paths, as well as application-level strategies, there
-      are no strategies which are usable by unmodified applications
+    \item Putting together, problem: unmodified applications can't take
+      advantage of detour routing solutions to Internet routing inefficiencies.
     \item A solution could be a system which uses and aggregates alternative
@@
     \vfill
-\begin{BVerbatim}
-+-----------------------------------+
-| ver(1) | op(1)  | reserved (2)    |
-+-----------------------------------+
-|           rip (4 bytes)           |
-+-----------------------------------+
-| rpt (2 bytes)   | dpt (2 bytes)   |
-+-----------------|-----------------+
-\end{BVerbatim}
+    \begin{bytefield}[bitwidth=0.6em]{32}
+      \bitheader{0,7,8,15,16,23,24,31}\\
+      \bitbox{8}{\texttt{ver}} & \bitbox{8}{\texttt{op}} & \bitbox{16}{\texttt{reserved}} \\
+      \wordbox{1}{\texttt{rip}} \\
+      \bitbox{16}{\texttt{rpt}} & \bitbox{16}{\texttt{dpt}} \\
+    \end{bytefield}
     \vfill

### 2017-06-04 src/aws_experiment.py
aws: i'm rollin in credits, let's go m4.large

@@ -88,7 +88,7 @@ CREATE_INSTANCES_PARAMS = {
     'MinCount': 1,
     'MaxCount': 1,
     'UserData': open(SETUP_SCRIPT).read(),
-    'InstanceType': 't2.micro',
+    'InstanceType': 'm4.large',
     'TagSpecifications': [
         {
             'ResourceType': 'instance',

### 2017-06-04 paper/presentation.tex
paper: git rid of out of date comments
presentation: updated related work from new changes

@@
       routes for MPTCP connections.
-    \item Contribution 1 - a mechanism for adding new routes to MPTCP
-      connections which are single-homed, with no modification to applications.
-    \item Contribution 2 - experimental validation that the mechanism works,
-      even at higher throughputs
+    \item Read contributions
     \end{itemize}
@@
 \SectionHead{}{
+
   \textbf{Transition:} Now in order to appreciate the place of this thesis,
-  let's examine some related work. The topic includes detour routing, overlay
-  routing, source routing, and multipath transport.
+  let's examine some related work. While there is a lot of related work, we will
+  restrict this discussion to three major points: overlay networks (which
+  operate at the network level), one very similar transport-layer approach, and
+  some application level approaches.
+
 }
@@
       compatibility concerns - routed segments across different paths. Worked on
-      single-homed devices due to RON}
+      single-homed devices due to RON. Unmodified applications may not use.}
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Other Related Items}
+  \frametitle{Application Layer}
 
-  \note[item]{Outside of overlay networks, several related pieces of work, both
-    in research and in the standards bodies.}
+  \note[item]{Outside of network-layer approaches, lots of app level approaches
+    to overlays, detours, and bandwidth aggregation.}
+  \note[item]{Since none of the net-layer approaches can be applied to
+    unmodified apps, apps designed around overlays/detours are the only ones
+    which can be used.}
 
   \begin{itemize}
-  \item Source routing (IPv4) and segment routing (IPv6)
-    \note[item]{Notable - it is possible to specify intermediaries in both IPv4
-      and IPv6. For security reasons these aren't respected. Standards exist
-      mainly for use in internal networks.}
-  \item Binder: aggregating bandwidth of several Internet gateways
-    \mycite{binder}
-    \note[item]{One use of MPTCP that is very similar was called Binder. A rural
-      town was away from any ISP. Had several potential gateways far away, but
-      low bandwidth and not good reliability. Established connections across
-      each gateway, and created an OpenVPN connection to an Internet-side
-      server, with subflow across each.}
-  \item SOSR: Scalable One-hop Source Routing \mycite{gummadi2004improving}
-    \note[item]{Work inspired by RON,
-      but trying to simplify. This work simply tunneled connections across a
-      single hop, using a NAT approach similar to ours. Found they got most of
-      RON's benefit without the full overlay system.}
-  \item IETF MPTCP proxying draft
-    \note[item]{Similar because allows single-homed devices to use MPTCP.
-      However this approach expects non MPTCP server, and is ``transparent''
-      proxy.}
+    \item Gnutella\mycite{adar2000free}
+      \note[item]{\textbf{Gnutella} was a file sharing network. It allowed
+        queries and responses to be forwarded from peer to peer as an overlay.
+        But actual file exchanges were single path.}
+      \begin{itemize}
+      \item Requests forwarded via overlay network
+      \item Content exchanged via single path
+      \end{itemize}
+    \item BitTorrent\mycite{cohen2003incentives}
+      \note[item]{\textbf{BitTorrent} is a protocol for file sharing which
+        allows content to be exchanged along the overlay. Each peer shares with
+        many other peers. Somewhat similar to detour routing - gets around
+        routing inefficiencies by simply using more paths}
+      \begin{itemize}
+      \item Pieces of content exchanged between many pairs of peers
+      \item Multiple paths simulate detour routing
+      \end{itemize}
+    \item HTTP Range Requests\mycite{kaspar2010enhancing}
+      \note[item]{\textbf{HTTP Range Requests} allow an app to request byte
+        ranges of a file. Kaspar et al demonstrate a mechanism for downloading
+        or streaming a file using range requests on multiple network interfaces}
+      \begin{itemize}
+      \item Range requests allow requesting byte ranges of a file
+      \item Request from different network interfaces or to different endpoints
+        to create alternative paths
+      \end{itemize}
   \end{itemize}

### 2017-06-04 data/2017-05-21-cpu-limit/plots.py
paper: begin table for data

@@ -33,15 +33,18 @@ SCEN = {
 
 
 def make_single_plot(base):
-    print(base)
     data = []
     labels = []
+    text = META[base]
     for s, label in SCEN.items():
-        data.append(o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000)
+        arr = o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000
+        data.append(arr)
+        text += ' & %.2f (%.2f)' % (arr.mean(), arr.std())
         labels.append(label)
 
+    print(text + r' \\')
+
     fig, ax = plt.subplots()
-    print(list(map(len, data)))
     ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
     ax.set_title('Throughput Comparison: ' + META[base])
@@ -49,6 +52,10 @@ def make_single_plot(base):
 
 
 def make_all_plots():
+    title = ''
+    for v in SCEN.values():
+        title += r' & \textbf{%s}' % v
+    print(title + r'\\')
     for base in META.keys():
         make_single_plot(base)
 

### 2017-06-04 data/param_validation/plots.py
paper: include data from parameter validation into table

@@ -37,15 +37,18 @@ SCEN = {
 
 
 def make_single_plot(base):
-    print(base)
     data = []
     labels = []
+    text = META[base]
     for s, label in SCEN.items():
-        data.append(o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000)
+        arr = o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000
+        data.append(arr)
+        text += ' & %.2f (%.2f)' % (arr.mean(), arr.std())
         labels.append(label)
 
+    print(text + r' \\')
+
     fig, ax = plt.subplots()
-    print(list(map(len, data)))
     ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
     ax.set_title('Throughput Comparison: ' + META[base])
@@ -53,6 +56,10 @@ def make_single_plot(base):
 
 
 def make_all_plots():
+    title = ''
+    for v in SCEN.values():
+        title += r' & \textbf{%s}' % v
+    print(title + r'\\')
     for base in META.keys():
         make_single_plot(base)
 

### 2017-06-04 paper/presentation.tex
presentation: revamp evaluation commentary and points

@@
 \begin{frame}
-  \frametitle{Types of Experiment}
+  \frametitle{Types of Experiments}
 
@@
     detour routing can improve path characteristics
-  \item These experiments do not duplicate results
-  \item Answer the following:
+  \item We simply attempt show mechanism works as expected
+  \item Answer the following
     \begin{itemize}
-    \item Given a network where an alternative route has better properties, can
-      this mechanism effectively use it?
-    \item Given a network where there is an alternative route and the potential
-      for aggregation, can this mechanism aggregate throughput?
-    \item Can this mechanism run without modification on the Internet?
+    \item Can we achieve throughput of best available path?
+    \item When bandwidth aggregation is possible, can we aggregate path
+      bandwidth?
+    \item What overheads exist in this mechanism?
+    \item Can this mechanism be used across the Internet at higher throughput?
     \end{itemize}
@@
     \item \textbf{Normal:} no loss
-    \item \textbf{Lossy:} 1\% packet loss on Link 2
-    \item \textbf{Delayed:} 100ms delay on Link 2
+    \item \textbf{Lossy:} 1\% packet loss on Link 2 \note[item]{For lossy, we
+        tried 0.5\%, 1\%, 2\%, and 5\%. Both 1 and 2 were reasonable - causing
+        decreased throughput but not too bad. 0.5 had nearly no effect. 5 had
+        too much of an effect. Can show graphs if interested.}
+    \item \textbf{Delayed:} 100ms delay on Link 2 \note[item]{For latency, we
+        tried 20 and 50 ms. 100 only just started to show throughput drop,
+        likely due to slow start.} \note[item]{Large differences in latency
+        affect MPTCP more - needs huge buffers to achieve full potential.}
     \end{itemize}
-  \item Total of 6 scenarios
+  \item Workload: send as much data as possible from client to server
   \end{itemize}
@@
   \end{itemize}
-  \note[item]{Don't forget to say we use iperf to measure throughput}
+  \note[item]{Don't forget to say we use iperf to measure throughput, and try to
+  send lots of data.}
 \end{frame}
@@
   \item This is a comparison of throughput on Symmetric, Normal. takeaways:
-  \item (1) MPTCP is more variable than TCP
-  \item (2) The VPN and NAT approaches don't perform as well when there is
-    nothing to gain. This seems to be a result of their overhead and maybe
-    congestion control.
+  \item (1) MPTCP has about 140 kbps overhead compared to standard TCP, or about
+    1.5\%
+  \item Incidentally, segment size 1460, 28 byte overhead of DSS, puts overhead
+    at 1.9\%, so not terribly far off.
+  \item (2) NAT has no observable overhead in terms of throughput
+  \item (3) VPN has observable overhead. 630kbps overhead, or about 6.6\%
+  \item Incidintally, segment size 1337 compared to 1432 is about 6.6\%, so spot
+    on!
+  \item (4) When VPN used with MPTCP, overhead falls to somewhere between the
+    1.5\% and 6.6\%, since data is split across a 1.5\% overhead and 6.6\%
+    overhead subflow
   \end{itemize}
@@
   \begin{itemize}
-  \item MPTCP trials are more variable
-  \item Mechanism introduces overhead when there is no potential for gain
+  \item MPTCP has 140kbps, or about 1.5\% overhead
+  \item VPN approach has overhead of about 6.6\%
+  \item Mechanism performs well even when no aggregation benefit possible
   \end{itemize}
@@
   \item When a small loss rate is added, the story begins to change.
-  \item Suddenly, MPTCP NAT and VPN are performing very comparably to vanilla
-    TCP over those same paths
-  \item MPTCP 1 subflow takes a huge performance hit
+  \item MPTCP NAT and VPN are performing very comparably to vanilla TCP over
+    those same paths
+  \item MPTCP 1 subflow takes a bigger performance hit than TCP, likely due to
+    its more conservative congestion control
+  \item MPTCP NAT and VPN have more variance because ACKS are transmitted on
+    both paths and get lost
   \end{itemize}
@@
   \begin{itemize}
-  \item MPTCP trials are more variable
-  \item Mechanism introduces overhead when there is no potential for gain
+  \item MPTCP has 140kbps, or about 1.5\% overhead
+  \item VPN approach has overhead of about 6.6\%
+  \item Mechanism performs well even when no aggregation benefit possible
   \item Mechanism performs similarly to TCP over best path when the default path
@@
     again.
-  \item The mechanism is capable of aggregating nearly all bandwidth
+  \item The mechanism achieves 98.5\% aggregation for NAT and 98.1\% aggregation
+    for VPN
   \end{itemize}
@@
   \begin{itemize}
-  \item With delayed link, there are some funky interactions with slow slow
-    start, but MPTCP still outperforms the best paths
+  \item With delayed link, MPTCP has more variance.
+  \item Similar to lossy link, ACKS can come on high latency link, blocking
+    receive window advance
   \end{itemize}

### 2017-06-04 data/aws4/plots.py
aws: some new funky data - rerunning vanilla to be sure

@@ -1,0 +1,61 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved
+SCEN = {
+    'server.ctrl.mptcp': '1 Subflow',
+    'server.nat.mptcp': 'NAT',
+    'server.vpn.mptcp': 'VPN',
+    'server.ctrl.vanilla': 'TCP',
+    'server.nat.vanilla': 'TCP (NAT)',
+    'server.vpn.vanilla': 'TCP (VPN)',
+}
+
+
+def make_throughput_plot():
+    data = []
+    labels = []
+    for s, label in SCEN.items():
+        data.append(o.get_numpy_array('%s.json' % s) / 1000000)
+        labels.append(label)
+
+    fig, ax = plt.subplots()
+    print(list(map(len, data)))
+    ax.boxplot(data, 0, '', labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: AWS')
+    fig.savefig('aws.pdf')
+
+
+def make_time_series_grid():
+    rows = ['mptcp', 'vanilla']
+    cols = ['ctrl', 'nat', 'vpn']
+    fig, axarr = plt.subplots(len(rows), len(cols), sharex=True, sharey=True)
+    for ridx, row in enumerate(rows):
+        for cidx, col in enumerate(cols):
+            data = o.get_time_series_arrays('server.%s.%s.json' % (col, row))
+            for series in data:
+                axarr[ridx][cidx].plot(series[:10])
+    axarr[0][0].set_ylabel('MPTCP')
+    axarr[1][0].set_ylabel('TCP')
+    axarr[1][0].set_xlabel('Control')
+    axarr[1][1].set_xlabel('NAT')
+    axarr[1][2].set_xlabel('VPN')
+    fig.savefig('timegrid-aws.pdf')
+
+
+if __name__ == '__main__':
+    if len(sys.argv) > 1 and sys.argv[1] == 'time':
+        make_time_series_grid()
+    else:
+        make_throughput_plot()

### 2017-06-05 paper/presentation.tex
paper,presentation: update Topology and AWS figures
presentation: future work
paper: reorder conclusion/fw. this is the last commit of presentation!

@@
   \item We use tunneling mechanisms to put a subflow across a detour host
+  \item Deployment scenario: friends with high bandwidth form a collective,
+    offer each other detouring services.
   \end{itemize}
@@
     \item Packets encapsulated via OpenVPN protocol
-    \item Detour still performs NAT
     \end{itemize}
@@
   \item Each link has 5ms delay.
+  \item Core links red, access links blue
   \end{itemize}
@@
   \item Performed at much higer level, but didn't show similar improvements
+  \item OpenVPN cannot sustain above 60Mbps in our setup
   \end{itemize}
@@
 \begin{frame}
-  \frametitle{Future Work}
-
-  \begin{itemize}
-  \item Detour Collective
-  \item Dynamic subflow addition and removal
-  \item Data scheduling
-  \end{itemize}
-\end{frame}
-
-\begin{frame}
   \frametitle{Summary}
@@
     single-homed devices.
-  \item Like MPTCP, this system works outside of experimental networks
+  \item Like MPTCP, this system works with unmodified applications
   \item System is capable of achieving similar performance to the best available
@@
 
-\appendix % slide numbering includes nothing after here
+\begin{frame}
+  \frametitle{Future Work}
+
+  \begin{itemize}
+  \item Deployment scenarios
+  \item Dynamic subflow addition and removal
+  \item Data scheduling
+  \item 0-RTT NAT establishment
+  \end{itemize}
+\end{frame}
+
+\appendix % slide numbering resets here
 

### 2017-06-09 setup.sh
lets update packages for great good

@@ -3,6 +3,7 @@
 
 sudo dpkg -i tmp/*.deb
 
+sude apt update -y
 sudo apt install -y python3-pip libnl-genl-3-dev openvpn iperf3
 sudo pip3 install psutil
 

### 2018-02-18 etc/aws_exp_setup.sh
Add all sorts of key pairs.

@@ -36,6 +36,8 @@ sudo -u ubuntu make -C ~ubuntu/src all
 
 echo Adding ssh keys...
 echo "ssh-rsa AAAAB3NzaC1yc2EAAAADAQABAAABAQDEupr15Lbq8jFBca/tXhydvD2nKQ/vWnxXzOndG7TqKGy53mtlgR6VzaV4WR3blrLdDG223a64VlQaj00xm+8e/eAn+IRH3RpV2ZvFT+BdZmej8E4y0sSlkiCw2sjtyKiGu0Pyk2HKbA8njG+aVsJ5YSQVwBBP1nEqKoMItvXzZwQJQ09SCeklT5zugcJz+frGFifjz1I6bs7Hb9vNa7hNi79JjJ1rNoeANkBmunW20+M7hMOuNjGWl47XGDlew7tn3vc2Eiiohf0cyrueBPq63Y7tMEfhmyCEPfakkjg5idt3soRGKCBqyoVuYyEIOJSTnuGBkThZteGHLlpKvMVz stephen@greed_2013-09-26" >> ~ubuntu/.ssh/authorized_keys
+echo "ssh-rsa AAAAB3NzaC1yc2EAAAADAQABAAABAQDE29+4kwdzCHhh/deeYYD8WcJchfD4IAq6m/E3XqDmS/4dgg7QzZk26gCdHQGcbUYRbYhtSnOG4ItkLBbM0u72YKyZ8VTtSeFFbvwBhgQqFUHgnxkyZCuUodJTTIwQBcE6feFbm0/StWJnpHis8Uvgjh/aunFYLJCZu9vIfgyLWPVIMpwItH2uMflTe9tmbp5KQyegtede4IkiUsThh5CslO+PnwVLhUfMvBc82Rtv866AM8UZXYqhYIgqwPuGC6r1mUfiaBS1VEwJYNfrvlAO3eOzWdVCYThMRhjvG72T0S/2HpkNHnpIx6BMG3lGpNxXIvogIRxQjB/KBSDTTeRT stephen@pride" >> ~ubuntu/.ssh/authorized_keys
[…]

### 2018-02-18 etc/aws_exp_setup_vanilla.sh

@@ -30,6 +30,8 @@ sudo -u ubuntu make -C ~ubuntu/src all
 
 echo Adding ssh keys...
 echo "ssh-rsa AAAAB3NzaC1yc2EAAAADAQABAAABAQDEupr15Lbq8jFBca/tXhydvD2nKQ/vWnxXzOndG7TqKGy53mtlgR6VzaV4WR3blrLdDG223a64VlQaj00xm+8e/eAn+IRH3RpV2ZvFT+BdZmej8E4y0sSlkiCw2sjtyKiGu0Pyk2HKbA8njG+aVsJ5YSQVwBBP1nEqKoMItvXzZwQJQ09SCeklT5zugcJz+frGFifjz1I6bs7Hb9vNa7hNi79JjJ1rNoeANkBmunW20+M7hMOuNjGWl47XGDlew7tn3vc2Eiiohf0cyrueBPq63Y7tMEfhmyCEPfakkjg5idt3soRGKCBqyoVuYyEIOJSTnuGBkThZteGHLlpKvMVz stephen@greed_2013-09-26" >> ~ubuntu/.ssh/authorized_keys
+echo "ssh-rsa AAAAB3NzaC1yc2EAAAADAQABAAABAQDE29+4kwdzCHhh/deeYYD8WcJchfD4IAq6m/E3XqDmS/4dgg7QzZk26gCdHQGcbUYRbYhtSnOG4ItkLBbM0u72YKyZ8VTtSeFFbvwBhgQqFUHgnxkyZCuUodJTTIwQBcE6feFbm0/StWJnpHis8Uvgjh/aunFYLJCZu9vIfgyLWPVIMpwItH2uMflTe9tmbp5KQyegtede4IkiUsThh5CslO+PnwVLhUfMvBc82Rtv866AM8UZXYqhYIgqwPuGC6r1mUfiaBS1VEwJYNfrvlAO3eOzWdVCYThMRhjvG72T0S/2HpkNHnpIx6BMG3lGpNxXIvogIRxQjB/KBSDTTeRT stephen@pride" >> ~ubuntu/.ssh/authorized_keys
[…]

### 2018-02-18 data/2018-02-14-EXT/plots.py
Plots and table for extended parameter range.

@@ -1,0 +1,94 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Produce plots.
+"""
+
+from os.path import dirname, abspath
+import sys
+sys.path.insert(0, dirname(dirname(abspath(__file__))))
+import output as o
+
+import matplotlib.pyplot as plt
+
+# order preserved (thanks python 3.6...?)
+META = {
+    'lossy05': 'Core-limited, 0.5% Loss',
+    'lossy2': 'Core-limited, 2% Loss',
+    'lossy5': 'Core-limited, 5% Loss',
+    'delayed20': 'Core-limited, 20ms Delay',
+    'delayed50': 'Core-limited, 50ms Delay',
+    'sym_lossy05': 'Symmetric, 0.5% Loss',
+    'sym_lossy2': 'Symmetric, 2% Loss',
+    'sym_lossy5': 'Symmetric, 5% Loss',
+    'sym_delayed20': 'Symmetric, 20ms Delay',
+    'sym_delayed50': 'Symmetric, 50ms Delay',
+}
+
+# order preserved
+SCEN = {
+    'control.mptcp': '1 Subflow',
+    'nat.mptcp': 'NAT',
+    'vpn.mptcp': 'VPN',
+    'control.vanilla': 'TCP',
+    'nat.vanilla': 'TCP (NAT)',
+    'vpn.vanilla': 'TCP (VPN)',
+}
+
+
+def make_single_plot(base):
+    data = []
+    labels = []
+    text = META[base]
+    for s, label in SCEN.items():
+        arr = o.get_numpy_array('%s.%s.json' % (base, s)) / 1000000
+        data.append(arr)
+        text += ' & %.2f (%.2f)' % (arr.mean(), arr.std())
+        labels.append(label)
+
+    print(text + r' \\')
+
+    fig, ax = plt.subplots()
+    ax.boxplot(data, labels=labels)
+    ax.set_ylabel('Throughput (Mbps)')
+    ax.set_title('Throughput Comparison: ' + META[base])
+    fig.savefig('%s.pdf' % base.replace('.', '-').replace('_', '-'))
+
+
+def make_all_plots():
+    title = ''
+    for v in SCEN.values():
+        title += r' & \textbf{%s}' % v
+    print(title + r'\\')
+    for base in META.keys():
+        make_single_plot(base)
+
+
+def make_time_series_grid(base):
+    print(base)
+    rows = ['mptcp', 'vanilla']
[…]

### 2018-02-26 data/2018-02-18-AWS/plots.py
AWS plots now produce numbers

@@ -25,9 +25,13 @@ SCEN = {
 def make_throughput_plot():
     data = []
     labels = []
+    text = 'AWS'
     for s, label in SCEN.items():
-        data.append(o.get_numpy_array('%s.json' % s) / 1000000)
+        arr = o.get_numpy_array('%s.json' % s) / 1000000
+        data.append(arr)
+        text += ' & %.2f (%.2f)' % (arr.mean(), arr.std())
         labels.append(label)
+    print(text + r' \\')
 
     fig, ax = plt.subplots()
     print(list(map(len, data)))

### 2018-02-26 data/2018-02-13-MAIN/plots.py
Updated plots to not have titles, add timegrid + table

@@ -47,7 +47,7 @@ def make_single_plot(base):
     fig, ax = plt.subplots()
     ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
-    ax.set_title('Throughput Comparison: ' + META[base])
+    #ax.set_title('Throughput Comparison: ' + META[base])
     fig.savefig('%s.pdf' % base.replace('.', '-').replace('_', '-'))
 
 

### 2018-02-26 data/2018-02-14-EXT/plots.py

@@ -51,7 +51,7 @@ def make_single_plot(base):
     fig, ax = plt.subplots()
     ax.boxplot(data, labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
-    ax.set_title('Throughput Comparison: ' + META[base])
+    #ax.set_title('Throughput Comparison: ' + META[base])
     fig.savefig('%s.pdf' % base.replace('.', '-').replace('_', '-'))
 
 

### 2018-02-26 data/2018-02-18-AWS/plots.py

@@ -37,7 +37,7 @@ def make_throughput_plot():
     print(list(map(len, data)))
     ax.boxplot(data, 0, '', labels=labels)
     ax.set_ylabel('Throughput (Mbps)')
-    ax.set_title('Throughput Comparison: AWS')
+    #ax.set_title('Throughput Comparison: AWS')
     fig.savefig('aws.pdf')
 
 
